### 2024-09-10 causalDT-manuscript/notebooks/paper_figures.Rmd
update sims and paper figures

---
title: "Causal Distillation Trees"
output: html_document
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE
)

KEEP_METHODS <- c(
  "Distilled Causal Forest",
  "Distilled Rboost",
  "Distilled Rlasso",
  "Virtual Twins",
  "Causal Tree (pruned)",
  "Causal Tree (unpruned)",
  "Linear Regression",
  "Lasso"
)
COLORS <- c(
  # "Distilled Causal Forest" = "#3153A5",
  "Distilled Causal Forest" = "#1f5d8f",
  # "Distilled Causal Forest" = "#5d8cae",
  # "Distilled Rboost" = "#94B1c9",
  # "Distilled Rboost" = "#678096",
  ## "Distilled Rboost" = "#79b2dd",
  "Distilled Rboost" = "#6aafe4",
  # "Distilled Rboost" = "#6DA7D3",
  # "Distilled Rboost" = "#6DA7D3",
  # "Distilled Rlasso" = "#ACC2D0",
  "Distilled Rlasso" = "#9ab6c8",
  # "Distilled Rlasso" = "#B8CEDD",
  "Virtual Twins" = "#768c50",
  # "Virtual Twins" = "#6F9940",
  "Causal Tree (pruned)" = "#953a2f",
  "Causal Tree (unpruned)" = "#DB9685",
  # "Linear Regression" = "#DAA520",
  "Linear Regression" = "#b2954a",
  "Lasso" = "#D2BE90"
  # "Lasso" = "#e4d2a4"
)

FACET_LABS <- list(
  "0.2" = "Weak",
  "0.6" = "Moderate",
  "1" = "High"
)
signal_labeller <- function(variable, value) {
  if (variable == "tau_heritability") {
    return(FACET_LABS[as.character(value)])
  } else {
    return(value)
  }
}

FIG_DIR <- here::here("results", "figures")
if (!dir.exists(FIG_DIR)) {
  dir.create(FIG_DIR, recursive = TRUE)
}
```

# Simulations {.tabset .tabset-vmodern}

```{r}
EXP_NAME <- "Test"
source(here::here(file.path("meals", "setup.R")))
source(here::here(file.path("meals", "shared_dgps.R")))
source(here::here(file.path("meals", "shared_methods.R")))
source(here::here(file.path("meals", "shared_evaluators.R")))
source(here::here(file.path("meals", "shared_visualizers.R")))
source(here::here(file.path("meals", "shared_experiments.R")))
experiment <- experiment |>
  add_dgp(gaussian_X_unbiased_Z_and) |>
  add_vary_across(
    .dgp = gaussian_X_unbiased_Z_and$name,
    tau_heritability = c(0.2, 0.4, 0.6, 0.8, 1)
  )
```

```{r results = "asis"}
sim_names <- c(
  "AND DGP",
  "Additive DGP",
  "OR DGP",
  "AND with Linear Covariates DGP",
  "Additive with Linear Covariates DGP",
  "OR with Linear Covariates DGP"
)
sim_results_dir <- here::here(
  file.path("results", "Unbiased W")
)

plt_subgroup_ls <- list()
plt_subgroup2_ls <- list()
plt_thresholds_ls <- list()
plt_stability_ls <- list()
plt_freq_ls <- list()
for (sim_name in sim_names) {
  cat(sprintf("\n\n## %s {.tabset .tabset-pills .tabset-square}\n\n", sim_name))

  # fit_results <- readRDS(
  #   file.path(sim_results_dir, sim_name, "Varying tau_heritability", "fit_results.rds")
  # ) |>
  #   dplyr::filter(
  #     .method_name %in% !!KEEP_METHODS
  #   )
  # eval_results <- evaluate_experiment(experiment, fit_results)
  # viz_results <- visualize_experiment(experiment, fit_results, eval_results)
  # 
  # plt_selection <- viz_results$`Subgroup Feature Selection Errors Plot` +
  #   ggplot2::scale_color_manual(
  #     breaks = KEEP_METHODS,
  #     values = COLORS
  #   ) +
  #   ggplot2::scale_fill_manual(
  #     breaks = KEEP_METHODS,
  #     values = COLORS
  #   )
  # 
  # plt_cate <- viz_results$`Subgroup CATE Errors Plot` +
  #   ggplot2::scale_color_manual(
  #     breaks = KEEP_METHODS,
  #     values = COLORS
  #   ) +
  #   ggplot2::scale_fill_manual(
  #     breaks = KEEP_METHODS,
  #     values = COLORS
  #   ) +
  #   ggplot2::labs(
  #     y = "CATE RMSE"
  #   ) +
  #   vthemes::theme_vmodern(
  #     axis_title_size = 16,
  #     axis_text_size = 12,
  #     legend_title_size = 16,
  #     legend_text_size = 12
  #   ) +
  #   ggplot2::theme(
  #     strip.background = ggplot2::element_blank(),
  #     strip.placement = "outside"
  #     # strip.background = ggplot2::element_rect(
  #     #   color = "transparent", fill = "transparent"
  #     # ),
  #     # strip.text = ggplot2::element_text(
  #     #   color = "black", face = "bold", size = 16
  #     # )
  #   ) +
  #   ggplot2::guides(
  #     color = "none", fill = "none"
  #   )
  # 
  # pplt <- patchwork::wrap_plots(
  #   plt_selection, plt_cate, nrow = 1, guides = "collect",
  #   widths = c(3.5, 1)
  # ) &
  #   ggplot2::labs(x = NULL)
  # plt_subgroup_ls[[sim_name]] <- pplt
  # plt <- patchwork::wrap_elements(pplt) +
  #   ggplot2::labs(
  #     tag = expression(
  #       bold(paste("Proportion of Variance Explained in ", tau))
  #     )
  #   ) +
  #   ggplot2::theme(
  #     plot.tag = ggplot2::element_text(
  #       size = 16
  #     ),
  #     plot.tag.position = c(0.45, 0),
  #     plot.margin = ggplot2::unit(c(-0.1, 0, .2, -0.1), units = "in")
  #   )
  # plt_subgroup2_ls[[sim_name]] <- plt

  fit_results <- readRDS(
    file.path(sim_results_dir, sim_name, "Varying tau_heritability", "fit_results.rds")
  ) |>
    dplyr::filter(
      .method_name %in% !!KEEP_METHODS,
      tau_heritability %in% c(0.2, 0.6, 1)
    )
  eval_results <- evaluate_experiment(experiment, fit_results)
  viz_results <- visualize_experiment(experiment, fit_results, eval_results)

  plt_thresholds <- viz_results$`Subgroup Thresholds Distribution Plot` +
    ggplot2::scale_color_manual(
      breaks = KEEP_METHODS,
      values = COLORS,
      labels = stringr::str_remove(KEEP_METHODS, " \\(pruned\\)")
    ) +
    ggplot2::scale_fill_manual(
      breaks = KEEP_METHODS,
      values = COLORS,
      labels = stringr::str_remove(KEEP_METHODS, " \\(pruned\\)")
    ) +
    ggplot2::facet_grid(.var ~ tau_heritability, labeller = signal_labeller)
  plt_thresholds_ls[[sim_name]] <- plt_thresholds

  plt_stability <- viz_results$`Subgroup Stability Diagnostics Plot`[[1]] +
    ggplot2::scale_color_manual(
      breaks = KEEP_METHODS,
      values = COLORS,
      labels = stringr::str_remove(KEEP_METHODS, " \\(pruned\\)")
    ) +
    ggplot2::scale_fill_manual(
      breaks = KEEP_METHODS,
      values = COLORS,
      labels = stringr::str_remove(KEEP_METHODS, " \\(pruned\\)")
    ) +
    ggplot2::facet_grid(~ tau_heritability, labeller = signal_labeller) +
    ggplot2::coord_cartesian(xlim = c(1, 4)) +
    ggplot2::scale_x_continuous(
      breaks = 1:4
    ) +
    ggplot2::labs(
      x = "Tree Depth",
      y = "Mean Jaccard Stability",
      color = "Method",
      fill = "Method"
    ) +
    vthemes::theme_vmodern(size_preset = "medium")
  plt_stability_ls[[sim_name]] <- plt_stability

  plt_nsplits <- viz_results$`Number of Splits Plot` +
    ggplot2::facet_grid(tau_heritability ~ ., labeller = signal_labeller) +
    viridis::scale_fill_viridis(
      begin = 0.15, end = 0.95, option = "magma", limits = c(0., 7)
    )
  plt_ntrees <- viz_results$`Number of Trees Plot` +
    ggplot2::facet_grid(tau_heritability ~ ., labeller = signal_labeller) +
    viridis::scale_fill_viridis(
      begin = 0.15, end = 0.95, option = "magma", limits = c(0, 100)
    )
  plt <- patchwork::wrap_plots(
    plt_ntrees, plt_nsplits, nrow = 1
  ) +
    patchwork::plot_layout(
      axis_titles = "collect_y"
    ) &
    ggplot2::labs(
      x = "Variable"
    )

  plt_freq_ls[[sim_name]] <- plt
}
```

```{r}
plt_subgroup <- purrr::map2(
  plt_subgroup_ls[1:3], 1:3,
  function(plt, idx) {
    sim_name <- names(plt_subgroup_ls[idx])
    tag <- LETTERS[idx]
    if (sim_name != names(plt_subgroup_ls[2])) {
      plt <- plt &
        ggplot2::guides(
          color = "none", fill = "none"
        )
      hjust <- 0.78
    } else {
      hjust <- 0.82
    }
    plt[[1]] <- plt[[1]] +
      ggplot2::labs(
        tag = sprintf("(%s) %s", tag, sim_name)
      ) +
      ggplot2::theme(
        plot.tag = ggplot2::element_text(
          size = 18, face = "italic", angle = 90, hjust = 0.5, vjust = 1
        ),
        plot.tag.position = c(-.03, 0.48),
        plot.margin = ggplot2::unit(c(-0.1, 0, 0, 0.5), units = "in")
      )
    plt <- plt &
      ggplot2::theme(
        panel.background = ggplot2::element_rect(
          fill = "grey98"
        ),
        panel.grid.major = ggplot2::element_line(
          color = "grey98"
        )
      )
      # ggplot2::theme(
      #   plot.title = ggplot2::element_text(
      #     size = 16, face = "bold", hjust = hjust
      #   ),
      #   plot.title.position = "plot"
      # )
    return(plt)
  }
) |>
  patchwork::wrap_plots(
    ncol = 1
  ) |>
  patchwork::wrap_elements() +
  ggplot2::labs(
    tag = expression(
      bold(paste("Proportion of Variance Explained in ", tau))
    )
  ) +
  ggplot2::theme(
    plot.tag = ggplot2::element_text(
      size = 16
    ),
    plot.tag.position = c(0.45, 0),
    plot.margin = ggplot2::unit(c(-0.1, 0, .2, -0.1), units = "in")
  )
ggplot2::ggsave(
  plt_subgroup, 
  filename = file.path(FIG_DIR, "sim_subgroup_errors.pdf"),
  height = 9, 
  width = 14
)

plt_subgroup_cov <- purrr::map2(
  plt_subgroup_ls[4:6], 4:6,
  function(plt, idx) {
    sim_name_orig <- names(plt_subgroup_ls[idx])
    sim_name <- stringr::str_replace(
      sim_name_orig, "with", "DGP with\n"
    ) |>
      stringr::str_remove(" DGP$")
    tag <- LETTERS[idx - 3]
    if (sim_name_orig != names(plt_subgroup_ls[5])) {
      plt <- plt &
        ggplot2::guides(
          color = "none", fill = "none"
        )
      hjust <- 0.78
    } else {
      hjust <- 0.82
    }
    plt[[1]] <- plt[[1]] +
      ggplot2::labs(
        tag = sprintf("(%s) %s", tag, sim_name)
      ) +
      ggplot2::theme(
        plot.tag = ggplot2::element_text(
          size = 18, face = "italic", angle = 90, hjust = 0.5, vjust = 1
        ),
        plot.tag.position = c(-.06, 0.48),
        plot.margin = ggplot2::unit(c(-0.1, 0, 0, 0.75), units = "in")
      )
    plt <- plt &
      ggplot2::theme(
        panel.background = ggplot2::element_rect(
          fill = "grey98"
        ),
        panel.grid.major = ggplot2::element_line(
          color = "grey98"
        )
      )
      # ggplot2::theme(
      #   plot.title = ggplot2::element_text(
      #     size = 16, face = "bold", hjust = hjust
      #   ),
      #   plot.title.position = "plot"
      # )
    return(plt)
  }
) |>
  patchwork::wrap_plots(
    ncol = 1
  ) |>
  patchwork::wrap_elements() +
  ggplot2::labs(
    tag = expression(
      bold(paste("Proportion of Variance Explained in ", tau))
    )
  ) +
  ggplot2::theme(
    plot.tag = ggplot2::element_text(
      size = 16
    ),
    plot.tag.position = c(0.45, 0),
    plot.margin = ggplot2::unit(c(-0.1, 0, .2, -0.1), units = "in")
  )
ggplot2::ggsave(
  plt_subgroup_cov, 
  filename = file.path(FIG_DIR, "sim_subgroup_errors_cov.pdf"),
  height = 9, 
  width = 14
)
```

```{r}
plt_thresholds <- purrr::map2(
  plt_thresholds_ls[1:3], 1:3,
  function(plt, idx) {
    sim_name <- names(plt_thresholds_ls[idx])
    tag <- LETTERS[idx]
    plt <- plt +
      ggplot2::labs(
        title = sprintf("(%s) %s", tag, sim_name)
      ) +
      ggplot2::theme(
        plot.title = ggplot2::element_text(
          size = 18, face = "italic", hjust = 0.5
        ),
        panel.background = ggplot2::element_rect(
          fill = "grey98"
        ),
        panel.grid.major = ggplot2::element_line(
          color = "grey98"
        )
      )
    return(plt)
  }
) |>
  patchwork::wrap_plots(
    ncol = 1, guides = "collect"
  ) +
  patchwork::plot_layout(
    axes = "collect"
  )
ggplot2::ggsave(
  plt_thresholds, 
  filename = file.path(FIG_DIR, "sim_thresholds.pdf"),
  height = 10, 
  width = 11
)

plt_thresholds_cov <- purrr::map2(
  plt_thresholds_ls[4:6], 4:6,
  function(plt, idx) {
    sim_name <- stringr::str_replace(
      names(plt_thresholds_ls[idx]), "with", "DGP with\n"
    ) |>
      stringr::str_remove(" DGP$")
    tag <- LETTERS[idx - 3]
    plt <- plt +
      ggplot2::labs(
        title = sprintf("(%s) %s", tag, sim_name)
      ) +
      ggplot2::theme(
        plot.title = ggplot2::element_text(
          size = 18, face = "italic", hjust = 0.5
        ),
        panel.background = ggplot2::element_rect(
          fill = "grey98"
        ),
        panel.grid.major = ggplot2::element_line(
          color = "grey98"
        )
      )
    return(plt)
  }
) |>
  patchwork::wrap_plots(
    ncol = 1, guides = "collect"
  ) +
  patchwork::plot_layout(
    axes = "collect"
  )
ggplot2::ggsave(
  plt_thresholds_cov, 
  filename = file.path(FIG_DIR, "sim_thresholds_cov.pdf"),
  height = 10, 
  width = 11
)
```

```{r}
# plt_ntrees_ls <- purrr::map(
#   plt_freq_ls[1:3], ~ .x[[1]]
# )
# plt_nsplits_ls <- purrr::map(
#   plt_freq_ls[1:3], ~ .x[[2]]
# )
# plt_ntrees <- patchwork::wrap_plots(
#   plt_ntrees_ls, nrow = 1, guides = "collect"
# ) +
#   patchwork::plot_layout(
#     axes = "collect"
#   )

plt_freq <- purrr::map2(
  plt_freq_ls[1:3], 1:3,
  function(plt, idx) {
    sim_name <- names(plt_freq_ls[idx])
    tag <- LETTERS[idx]
    plt <- patchwork::wrap_elements(plt) +
      ggplot2::labs(
        tag = sprintf("(%s) %s", tag, sim_name)
      ) +
      ggplot2::theme(
        plot.tag = ggplot2::element_text(
          size = 18, face = "italic", hjust = 0.5
        ),
        plot.tag.position = c(0.47, 1.02),
        plot.margin = ggplot2::unit(c(0.3, 0, 0, 0.5), units = "in")
      )
    return(plt)
  }
) |>
  patchwork::wrap_plots(
    ncol = 1
  )
ggplot2::ggsave(
  plt_freq, 
  filename = file.path(FIG_DIR, "sim_freq.pdf"),
  height = 16, 
  width = 14
)

plt_freq_cov <- purrr::map2(
  plt_freq_ls[4:6], 4:6,
  function(plt, idx) {
    sim_name <- stringr::str_replace(
      names(plt_freq_ls[idx]), "with", "DGP with"
    ) |>
      stringr::str_remove(" DGP$")
    tag <- LETTERS[idx - 3]
    plt <- patchwork::wrap_elements(plt) +
      ggplot2::labs(
        tag = sprintf("(%s) %s", tag, sim_name)
      ) +
      ggplot2::theme(
        plot.tag = ggplot2::element_text(
          size = 18, face = "italic", hjust = 0.5
        ),
        plot.tag.position = c(0.47, 1.02),
        plot.margin = ggplot2::unit(c(0.3, 0, 0, 0.5), units = "in")
      )
    return(plt)
  }
) |>
  patchwork::wrap_plots(
    ncol = 1
  )
ggplot2::ggsave(
  plt_freq_cov, 
  filename = file.path(FIG_DIR, "sim_freq_cov.pdf"),
  height = 16, 
  width = 14
)
```

```{r}
plt_stability <- purrr::map2(
  plt_stability_ls, 1:length(plt_stability_ls),
  function(plt, idx) {
    plt_names <- names(plt_stability_ls)
    sim_name <- dplyr::case_when(
      stringr::str_detect(plt_names, "with") ~ 
        stringr::str_replace(
          plt_names, "with", "DGP with\n"
        ) |> 
        stringr::str_remove(" DGP$"),
      TRUE ~ plt_names
    )[idx]
    tag <- LETTERS[idx]
    plt_df <- plt$data |>
      dplyr::mutate(
        sim_name = sprintf("(%s) %s", tag, sim_name)
      )
    return(plt_df)
  }
) |>
  purrr::list_rbind() |> 
  dplyr::filter(
    depth <= 4
  )
  ggplot2::ggplot() +
  ggplot2::geom_line(
    ggplot2::aes(
      x = depth, y = mean_jaccard, color = .method_name
    )
  ) +
  ggplot2::geom_ribbon(
    ggplot2::aes(
      x = depth,
      ymin = mean_jaccard - se_jaccard,
      ymax = mean_jaccard + se_jaccard,
      fill = .method_name
    ),
    alpha = 0.2
  ) +
  ggplot2::facet_grid(sim_name ~ tau_heritability, labeller = signal_labeller) +
  ggplot2::scale_color_manual(
    breaks = KEEP_METHODS,
    values = COLORS,
    labels = stringr::str_remove(KEEP_METHODS, " \\(pruned\\)")
  ) +
  ggplot2::scale_fill_manual(
    breaks = KEEP_METHODS,
    values = COLORS,
    labels = stringr::str_remove(KEEP_METHODS, " \\(pruned\\)")
  ) +
  ggplot2::coord_cartesian(xlim = c(1, 4)) +
  ggplot2::scale_x_continuous(
    breaks = 1:4
  ) +
  ggplot2::labs(
    x = "Tree Depth",
    y = "Mean Jaccard Stability",
    color = "Method",
    fill = "Method"
  ) +
  vthemes::theme_vmodern(
    size_preset = "medium", grid_color = "grey98"
  )
ggplot2::ggsave(
  plt_stability, 
  filename = file.path(FIG_DIR, "sim_stability.pdf"),
  height = 12, 
  width = 11
)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.