### 2024-12-05 causalDT-manuscript/notebooks/paper_figures.Rmd
add nreps_crossfit simulations + bug fixes

---
title: "Causal Distillation Trees"
output: html_document
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE
)

COLORS <- c(
  "Distilled Causal Forest" = "#1f5d8f",
  "Distilled Rboost" = "#6aafe4",
  "Virtual Twins" = "#768c50",
  "Causal Tree" = "#953a2f",
  "Linear Regression" = "#b2954a",
  "Lasso" = "#D2BE90"
)
CROSSFIT_COLORS <- c(
  "#981643", "#ED724F", "#F8B06E", "#FCE096", "#8FB989", "#4287B7", "#5E519B"
)
RULEFIT_COLORS <- c(
  "CART" = "#1f5d8f",
  "Rulefit (both, max depth = 3)" = "#FF9300",
  "Rulefit (both, max depth = 2)" = "#E7298A",
  "Rulefit (rules only, max depth = 3)" = "#66A61E",
  "Rulefit (rules only, max depth = 2)" = "#7570B3"
)
LINETYPES <- c(
  "Pruned" = "solid",
  "Unpruned" = "dashed"
)

FACET_LABS <- list(
  "0.2" = "Weak (0.2)",
  "0.6" = "Moderate (0.6)",
  "1" = "Strong (1)"
)
signal_labeller <- function(variable, value) {
  if (variable == "tau_heritability") {
    return(FACET_LABS[as.character(value)])
  } else if (variable == "metric") {
    return(METRIC_LABS[as.character(value)])
  } else {
    return(value)
  }
}

add_pruned_column <- function(data) {
  data |> 
    dplyr::mutate(
      .is_pruned = dplyr::case_when(
        stringr::str_detect(.method_name, "unpruned") ~ "Unpruned",
        TRUE ~ "Pruned"
      ),
      .method_name = stringr::str_remove(.method_name, " \\(unpruned\\)")
    )
}

rename_rulefit_methods <- function(data) {
  data <- data |> 
    dplyr::mutate(
      .method_name = dplyr::case_when(
        .method_name == "Distilled Causal Forest" ~ "CART",
        .method_name == "Distilled Causal Forest (rulefit v1)" ~ "Rulefit (both, max depth = 3)",
        .method_name == "Distilled Causal Forest (rulefit v2)" ~ "Rulefit (both, max depth = 2)",
        .method_name == "Distilled Causal Forest (rulefit v3)" ~ "Rulefit (rules only, max depth = 3)",
        .method_name == "Distilled Causal Forest (rulefit v4)" ~ "Rulefit (rules only, max depth = 2)",
      )
    )
  return(data)
}

FIG_DIR <- here::here("results", "figures")
if (!dir.exists(FIG_DIR)) {
  dir.create(FIG_DIR, recursive = TRUE)
}
```

# Simulations {.tabset .tabset-vmodern}

```{r}
EXP_NAME <- "Test"
source(here::here(file.path("meals", "setup.R")))
source(here::here(file.path("meals", "shared_dgps.R")))
source(here::here(file.path("meals", "shared_methods.R")))
source(here::here(file.path("meals", "shared_evaluators.R")))
source(here::here(file.path("meals", "shared_visualizers.R")))
source(here::here(file.path("meals", "shared_experiments.R")))
experiment <- experiment |>
  add_dgp(gaussian_X_unbiased_Z_and) |>
  add_vary_across(
    .dgp = gaussian_X_unbiased_Z_and$name,
    tau_heritability = c(0.2, 0.4, 0.6, 0.8, 1)
  )

example_experiment <- example_experiment |>
  add_dgp(gaussian_X_unbiased_Z_and) |>
  add_vary_across(
    .dgp = gaussian_X_unbiased_Z_and$name,
    n = c(100, 500, 1000, 2000)
  )

crossfit_experiment <- crossfit_experiment |>
  add_dgp(gaussian_X_unbiased_Z_and) |>
  add_vary_across(
    .dgp = gaussian_X_unbiased_Z_and$name,
    tau_heritability = c(0.2, 0.4, 0.6, 0.8, 1)
  )

rulefit_experiment <- rulefit_experiment |> 
  add_dgp(gaussian_X_unbiased_Z_and) |>
  add_vary_across(
    .dgp = gaussian_X_unbiased_Z_and$name,
    tau_heritability = c(0.2, 0.4, 0.6, 0.8, 1)
  )
```

## Toy Example {.tabset .tabset-pills .tabset-pills-square}

```{r results = "asis"}
sim_results_dir <- here::here(
  file.path("results", "Main Simulations")
)
sim_names <- c(
  "AND DGP",
  "Additive DGP",
  "OR DGP",
  "AND with Linear Covariates DGP",
  "Additive with Linear Covariates DGP",
  "OR with Linear Covariates DGP"
)

keep_methods_ls <- list(
  "causal_tree_only" = c("Causal Tree"),
  "all_methods" = c("Distilled Causal Forest", "Causal Tree")
)

for (keep_methods_idx in 1:length(keep_methods_ls)) {
  keep_methods <- keep_methods_ls[[keep_methods_idx]]
  keep_methods_name <- names(keep_methods_ls)[keep_methods_idx]
  for (sim_name in sim_names) {
    fit_results <- readRDS(
      file.path(
        sim_results_dir, sim_name, "Varying n", "fit_results.rds"
      )
    )
    eval_results <- readRDS(
      file.path(
        sim_results_dir, sim_name, "Varying n", "eval_results.rds"
      )
    )
    viz_results <- visualize_experiment(
      example_experiment, fit_results, eval_results
    )

    plt <- viz_results$`Number of Subgroups Plot`
    plt$data <- plt$data |> 
      dplyr::filter(
        .method_name %in% !!keep_methods
      )
    plt <- plt +
      ggplot2::scale_color_manual(
        breaks = keep_methods,
        values = COLORS
      ) +
      ggplot2::scale_fill_manual(
        breaks = keep_methods,
        values = COLORS
      ) +
      ggplot2::labs(
        x = "Sample Size"
      ) +
      ggplot2::theme(
        panel.background = ggplot2::element_rect(fill = "grey98"),
        panel.grid.major = ggplot2::element_line(color = "grey98")
      )
    ggplot2::ggsave(
      plt, 
      filename = file.path(
        FIG_DIR, 
        sprintf("sim_example_nsubgroup_%s.pdf", keep_methods_name)
      ),
      height = 3.1,
      width = 11.5
    )
  }
}
```

## Main Simulations {.tabset .tabset-pills .tabset-pills-square}

```{r results = "asis"}
sim_results_dir <- here::here(
  file.path("results", "Main Simulations")
)
SIM_NAMES <- c(
  "AND DGP" = "AND DGP",
  "Additive DGP" = "Additive DGP",
  "OR DGP" = "OR DGP",
  "AND with Linear Covariates DGP" = "AND DGP with Linear Covariates",
  "Additive with Linear Covariates DGP" = "Additive DGP with Linear Covariates",
  "OR with Linear Covariates DGP" = "OR DGP with Linear Covariates"
)

# load in results
eval_results_ls <- list()
viz_results_ls <- list()
for (sim_name in names(SIM_NAMES)) {
  fit_results <- readRDS(
    file.path(
      sim_results_dir, sim_name, "Varying tau_heritability", "fit_results.rds"
    )
  )
  eval_results <- readRDS(
    file.path(
      sim_results_dir, sim_name, "Varying tau_heritability", "eval_results.rds"
    )
  )
  viz_results <- visualize_experiment(experiment, fit_results, eval_results)
  eval_results_ls[[SIM_NAMES[[sim_name]]]] <- eval_results
  viz_results_ls[[SIM_NAMES[[sim_name]]]] <- viz_results
}
```

### Selected Subgroup Features + Estimated Subgroup CATEs

```{r}
KEEP_METHODS <- c(
  "Distilled Causal Forest",
  "Distilled Rboost",
  "Virtual Twins",
  "Causal Tree",
  "Linear Regression",
  "Lasso"
)

## without covariates
sim_names <- c(
  "AND DGP",
  "Additive DGP",
  "OR DGP"
)
plt <- purrr::map(
  sim_names,
  function(sim_name) {
    plt_name <- stringr::str_replace(sim_name, "with ", "with\n")
    plt <- viz_results_ls[[sim_name]]$`Subgroup Feature Selection Errors Plot`
    plt$data <- plt$data |>
      dplyr::filter(
        .method_name %in% KEEP_METHODS
      )
    plt_selection <- plt +
      ggplot2::scale_color_manual(
        breaks = KEEP_METHODS,
        values = COLORS
      ) +
      ggplot2::scale_fill_manual(
        breaks = KEEP_METHODS,
        values = COLORS
      )

    plt <- viz_results_ls[[sim_name]]$`Subgroup CATE Errors Plot`
    plt$data <- plt$data |>
      dplyr::filter(
        .method_name %in% KEEP_METHODS
      )
    plt_cate <- plt +
      ggplot2::scale_color_manual(
        breaks = KEEP_METHODS,
        values = COLORS
      ) +
      ggplot2::scale_fill_manual(
        breaks = KEEP_METHODS,
        values = COLORS
      ) +
      ggplot2::labs(
        y = "CATE RMSE"
      ) +
      ggplot2::theme(
        strip.background = ggplot2::element_blank(),
        strip.placement = "outside",
        axis.title.y = ggplot2::element_text(face = "bold", size = 16),
        axis.text = ggplot2::element_text(size = 12),
        legend.title = ggplot2::element_text(face = "bold", size = 16),
        legend.text = ggplot2::element_text(size = 12)
      )
    plt <- patchwork::wrap_plots(
      plt_selection, plt_cate, nrow = 1, guides = "collect", widths = c(3.5, 1)
    ) &
      ggplot2::labs(x = NULL)

    if (sim_name != sim_names[2]) {
      plt <- plt &
        ggplot2::guides(
          color = "none", fill = "none"
        )
      hjust <- 0.78
    } else {
      hjust <- 0.82
    }
    if (sim_name == sim_names[1]) {
      plt[[1]] <- plt[[1]] +
        ggplot2::labs(
          title = "(A) Selected Subgroup Features"
        )
      plt[[2]] <- plt[[2]] +
        ggplot2::labs(
          title = "(B) Estimated Subgroup CATEs"
        )
    }
    plt[[1]] <- plt[[1]] +
      ggplot2::labs(
        tag = plt_name
      ) +
      ggplot2::theme(
        plot.tag = ggplot2::element_text(
          size = 18, face = "italic", angle = 90, hjust = 0.5, vjust = 1
        ),
        plot.tag.position = c(-.03, 0.48),
        plot.margin = ggplot2::unit(c(0, 0.35, 0, 0.5), units = "in")
      )
    plt <- plt &
      ggplot2::theme(
        panel.background = ggplot2::element_rect(fill = "grey98"),
        panel.grid.major = ggplot2::element_line(color = "grey98"),
        plot.title = ggplot2::element_text(
          size = 18, face = "italic", hjust = 0.5, vjust = -1
        )
      )
    return(plt)
  }
) |>
  patchwork::wrap_plots(
    ncol = 1
  ) |>
  patchwork::wrap_elements() +
  ggplot2::labs(
    tag = expression(
      bold(paste("Proportion of Variance Explained in ", tau, " by Covariates"))
    )
  ) +
  ggplot2::theme(
    plot.tag = ggplot2::element_text(
      size = 16
    ),
    plot.tag.position = c(0.5, 0),
    plot.margin = ggplot2::unit(c(-0.1, 0, .2, -0.1), units = "in")
  )
ggplot2::ggsave(
  plt, 
  filename = file.path(FIG_DIR, "sim_subgroup_errors.pdf"),
  height = 9, 
  width = 14
)

## with covariates
sim_names <- c(
  "AND DGP with Linear Covariates",
  "Additive DGP with Linear Covariates",
  "OR DGP with Linear Covariates"
)
plt <- purrr::map(
  sim_names,
  function(sim_name) {
    plt_name <- stringr::str_replace(sim_name, "with ", "with\n")
    plt <- viz_results_ls[[sim_name]]$`Subgroup Feature Selection Errors Plot`
    plt$data <- plt$data |>
      dplyr::filter(
        .method_name %in% KEEP_METHODS
      )
    plt_selection <- plt +
      ggplot2::scale_color_manual(
        breaks = KEEP_METHODS,
        values = COLORS
      ) +
      ggplot2::scale_fill_manual(
        breaks = KEEP_METHODS,
        values = COLORS
      )

    plt <- viz_results_ls[[sim_name]]$`Subgroup CATE Errors Plot`
    plt$data <- plt$data |>
      dplyr::filter(
        .method_name %in% KEEP_METHODS
      )
    plt_cate <- plt +
      ggplot2::scale_color_manual(
        breaks = KEEP_METHODS,
        values = COLORS
      ) +
      ggplot2::scale_fill_manual(
        breaks = KEEP_METHODS,
        values = COLORS
      ) +
      ggplot2::labs(
        y = "CATE RMSE"
      ) +
      ggplot2::theme(
        strip.background = ggplot2::element_blank(),
        strip.placement = "outside",
        axis.title.y = ggplot2::element_text(face = "bold", size = 16),
        axis.text = ggplot2::element_text(size = 12),
        legend.title = ggplot2::element_text(face = "bold", size = 16),
        legend.text = ggplot2::element_text(size = 12)
      )
    plt <- patchwork::wrap_plots(
      plt_selection, plt_cate, nrow = 1, guides = "collect", widths = c(3.5, 1)
    ) &
      ggplot2::labs(x = NULL)

    if (sim_name != sim_names[2]) {
      plt <- plt &
        ggplot2::guides(
          color = "none", fill = "none"
        )
      hjust <- 0.78
    } else {
      hjust <- 0.82
    }
    if (sim_name == sim_names[1]) {
      plt[[1]] <- plt[[1]] +
        ggplot2::labs(
          title = "(A) Selected Subgroup Features"
        )
      plt[[2]] <- plt[[2]] +
        ggplot2::labs(
          title = "(B) Estimated Subgroup CATEs"
        )
    }
    plt[[1]] <- plt[[1]] +
      ggplot2::labs(
        tag = plt_name
      ) +
      ggplot2::theme(
        plot.tag = ggplot2::element_text(
          size = 18, face = "italic", angle = 90, hjust = 0.5, vjust = 1
        ),
        plot.tag.position = c(-.06, 0.48),
        plot.margin = ggplot2::unit(c(0, 0.35, 0, 0.75), units = "in")
      )
    plt <- plt &
      ggplot2::theme(
        panel.background = ggplot2::element_rect(fill = "grey98"),
        panel.grid.major = ggplot2::element_line(color = "grey98"),
        plot.title = ggplot2::element_text(
          size = 18, face = "italic", hjust = 0.5, vjust = -1
        )
      )
    return(plt)
  }
) |>
  patchwork::wrap_plots(
    ncol = 1
  ) |>
  patchwork::wrap_elements() +
  ggplot2::labs(
    tag = expression(
      bold(paste("Proportion of Variance Explained in ", tau, " by Covariates"))
    )
  ) +
  ggplot2::theme(
    plot.tag = ggplot2::element_text(
      size = 16
    ),
    plot.tag.position = c(0.45, 0),
    plot.margin = ggplot2::unit(c(-0.1, 0, .2, -0.1), units = "in")
  )
ggplot2::ggsave(
  plt, 
  filename = file.path(FIG_DIR, "sim_subgroup_errors_cov.pdf"),
  height = 9, 
  width = 14
)
```

```{r}
KEEP_METHODS <- c(
  "Distilled Causal Forest",
  "Distilled Causal Forest (unpruned)",
  "Distilled Rboost",
  "Distilled Rboost (unpruned)",
  "Causal Tree",
  "Causal Tree (unpruned)"
)

## without covariates
sim_names <- c(
  "AND DGP",
  "Additive DGP",
  "OR DGP"
)
plt <- purrr::map(
  sim_names,
  function(sim_name) {
    plt_name <- stringr::str_replace(sim_name, "with ", "with\n")
    plt <- viz_results_ls[[sim_name]]$`Subgroup Feature Selection Errors Plot`
    plt$data <- plt$data |>
      dplyr::filter(
        .method_name %in% KEEP_METHODS
      ) |> 
      add_pruned_column()
    plt$layers[[1]]$mapping <- ggplot2::aes(
      x = tau_heritability, 
      ymin = mean_feature_selection_err - se_feature_selection_err,
      ymax = mean_feature_selection_err + se_feature_selection_err,
      fill = .method_name,
      group = interaction(.method_name, .is_pruned)
    )
    plt$layers[[2]]$mapping <- ggplot2::aes(
      x = tau_heritability, 
      y = mean_feature_selection_err,
      color = .method_name,
      linetype = .is_pruned, 
      group = interaction(.method_name, .is_pruned)
    )

    plt_selection <- plt +
      ggplot2::scale_color_manual(
        breaks = KEEP_METHODS,
        values = COLORS
      ) +
      ggplot2::scale_fill_manual(
        breaks = KEEP_METHODS,
        values = COLORS
      ) +
      ggplot2::scale_linetype_manual(
        values = LINETYPES
      )

    plt <- viz_results_ls[[sim_name]]$`Subgroup CATE Errors Plot`
    plt$data <- plt$data |>
      dplyr::filter(
        .method_name %in% KEEP_METHODS
      ) |> 
      add_pruned_column()
    plt$layers[[1]]$mapping <- ggplot2::aes(
      x = tau_heritability, 
      ymin = mean_cate_err - se_cate_err,
      ymax = mean_cate_err + se_cate_err,
      fill = .method_name,
      group = interaction(.method_name, .is_pruned)
    )
    plt$layers[[2]]$mapping <- ggplot2::aes(
      x = tau_heritability, 
      y = mean_cate_err,
      color = .method_name,
      linetype = .is_pruned, 
      group = interaction(.method_name, .is_pruned)
    )

    plt_cate <- plt +
      ggplot2::scale_color_manual(
        breaks = KEEP_METHODS,
        values = COLORS
      ) +
      ggplot2::scale_fill_manual(
        breaks = KEEP_METHODS,
        values = COLORS
      ) +
      ggplot2::scale_linetype_manual(
        values = LINETYPES
      ) +
      ggplot2::labs(
        y = "CATE RMSE"
      ) +
      ggplot2::theme(
        strip.background = ggplot2::element_blank(),
        strip.placement = "outside",
        axis.title.y = ggplot2::element_text(face = "bold", size = 16),
        axis.text = ggplot2::element_text(size = 12),
        legend.title = ggplot2::element_text(face = "bold", size = 16),
        legend.text = ggplot2::element_text(size = 12)
      )
    plt <- patchwork::wrap_plots(
      plt_selection, plt_cate, nrow = 1, guides = "collect", widths = c(3.5, 1)
    ) &
      ggplot2::labs(x = NULL, linetype = "Pruning Mode")

    if (sim_name != sim_names[2]) {
      plt <- plt &
        ggplot2::guides(
          color = "none", fill = "none", linetype = "none"
        )
      hjust <- 0.78
    } else {
      hjust <- 0.82
    }
    if (sim_name == sim_names[1]) {
      plt[[1]] <- plt[[1]] +
        ggplot2::labs(
          title = "(A) Selected Subgroup Features"
        )
      plt[[2]] <- plt[[2]] +
        ggplot2::labs(
          title = "(B) Estimated Subgroup CATEs"
        )
    }
    plt[[1]] <- plt[[1]] +
      ggplot2::labs(
        tag = plt_name
      ) +
      ggplot2::theme(
        plot.tag = ggplot2::element_text(
          size = 18, face = "italic", angle = 90, hjust = 0.5, vjust = 1
        ),
        plot.tag.position = c(-.03, 0.48),
        plot.margin = ggplot2::unit(c(0, 0.35, 0, 0.5), units = "in")
      )
    plt <- plt &
      ggplot2::theme(
        panel.background = ggplot2::element_rect(fill = "grey98"),
        panel.grid.major = ggplot2::element_line(color = "grey98"),
        plot.title = ggplot2::element_text(
          size = 18, face = "italic", hjust = 0.5, vjust = -1
        ),
        legend.key.width = ggplot2::unit(0.9, "cm")
      )
    return(plt)
  }
) |>
  patchwork::wrap_plots(
    ncol = 1
  ) |>
  patchwork::wrap_elements() +
  ggplot2::labs(
    tag = expression(
      bold(paste("Proportion of Variance Explained in ", tau, " by Covariates"))
    )
  ) +
  ggplot2::theme(
    plot.tag = ggplot2::element_text(
      size = 16
    ),
    plot.tag.position = c(0.5, 0),
    plot.margin = ggplot2::unit(c(-0.1, 0, .2, -0.1), units = "in")
  )
ggplot2::ggsave(
  plt, 
  filename = file.path(FIG_DIR, "sim_subgroup_errors_with_pruning.pdf"),
  height = 9, 
  width = 14
)

## with covariates
sim_names <- c(
  "AND DGP with Linear Covariates",
  "Additive DGP with Linear Covariates",
  "OR DGP with Linear Covariates"
)
plt <- purrr::map(
  sim_names,
  function(sim_name) {
    plt_name <- stringr::str_replace(sim_name, "with ", "with\n")
    plt <- viz_results_ls[[sim_name]]$`Subgroup Feature Selection Errors Plot`
    plt$data <- plt$data |>
      dplyr::filter(
        .method_name %in% KEEP_METHODS
      ) |> 
      add_pruned_column()
    plt$layers[[1]]$mapping <- ggplot2::aes(
      x = tau_heritability, 
      ymin = mean_feature_selection_err - se_feature_selection_err,
      ymax = mean_feature_selection_err + se_feature_selection_err,
      fill = .method_name,
      group = interaction(.method_name, .is_pruned)
    )
    plt$layers[[2]]$mapping <- ggplot2::aes(
      x = tau_heritability, 
      y = mean_feature_selection_err,
      color = .method_name,
      linetype = .is_pruned, 
      group = interaction(.method_name, .is_pruned)
    )

    plt_selection <- plt +
      ggplot2::scale_color_manual(
        breaks = KEEP_METHODS,
        values = COLORS
      ) +
      ggplot2::scale_fill_manual(
        breaks = KEEP_METHODS,
        values = COLORS
      ) +
      ggplot2::scale_linetype_manual(
        values = LINETYPES
      )

    plt <- viz_results_ls[[sim_name]]$`Subgroup CATE Errors Plot`
    plt$data <- plt$data |>
      dplyr::filter(
        .method_name %in% KEEP_METHODS
      ) |> 
      add_pruned_column()
    plt$layers[[1]]$mapping <- ggplot2::aes(
      x = tau_heritability, 
      ymin = mean_cate_err - se_cate_err,
      ymax = mean_cate_err + se_cate_err,
      fill = .method_name,
      group = interaction(.method_name, .is_pruned)
    )
    plt$layers[[2]]$mapping <- ggplot2::aes(
      x = tau_heritability, 
      y = mean_cate_err,
      color = .method_name,
      linetype = .is_pruned, 
      group = interaction(.method_name, .is_pruned)
    )

    plt_cate <- plt +
      ggplot2::scale_color_manual(
        breaks = KEEP_METHODS,
        values = COLORS
      ) +
      ggplot2::scale_fill_manual(
        breaks = KEEP_METHODS,
        values = COLORS
      ) +
      ggplot2::scale_linetype_manual(
        values = LINETYPES
      ) +
      ggplot2::labs(
        y = "CATE RMSE"
      ) +
      ggplot2::theme(
        strip.background = ggplot2::element_blank(),
        strip.placement = "outside",
        axis.title.y = ggplot2::element_text(face = "bold", size = 16),
        axis.text = ggplot2::element_text(size = 12),
        legend.title = ggplot2::element_text(face = "bold", size = 16),
        legend.text = ggplot2::element_text(size = 12)
      )
    plt <- patchwork::wrap_plots(
      plt_selection, plt_cate, nrow = 1, guides = "collect", widths = c(3.5, 1)
    ) &
      ggplot2::labs(x = NULL, linetype = "Pruning Mode")

    if (sim_name != sim_names[2]) {
      plt <- plt &
        ggplot2::guides(
          color = "none", fill = "none", linetype = "none"
        )
      hjust <- 0.78
    } else {
      hjust <- 0.82
    }
    if (sim_name == sim_names[1]) {
      plt[[1]] <- plt[[1]] +
        ggplot2::labs(
          title = "(A) Selected Subgroup Features"
        )
      plt[[2]] <- plt[[2]] +
        ggplot2::labs(
          title = "(B) Estimated Subgroup CATEs"
        )
    }
    plt[[1]] <- plt[[1]] +
      ggplot2::labs(
        tag = plt_name
      ) +
      ggplot2::theme(
        plot.tag = ggplot2::element_text(
          size = 18, face = "italic", angle = 90, hjust = 0.5, vjust = 1
        ),
        plot.tag.position = c(-.06, 0.48),
        plot.margin = ggplot2::unit(c(0, 0.35, 0, 0.75), units = "in")
      )
    plt <- plt &
      ggplot2::theme(
        panel.background = ggplot2::element_rect(fill = "grey98"),
        panel.grid.major = ggplot2::element_line(color = "grey98"),
        plot.title = ggplot2::element_text(
          size = 18, face = "italic", hjust = 0.5, vjust = -1
        ),
        legend.key.width = ggplot2::unit(0.9, "cm")
      )
    return(plt)
  }
) |>
  patchwork::wrap_plots(
    ncol = 1
  ) |>
  patchwork::wrap_elements() +
  ggplot2::labs(
    tag = expression(
      bold(paste("Proportion of Variance Explained in ", tau, " by Covariates"))
    )
  ) +
  ggplot2::theme(
    plot.tag = ggplot2::element_text(
      size = 16
    ),
    plot.tag.position = c(0.45, 0),
    plot.margin = ggplot2::unit(c(-0.1, 0, .2, -0.1), units = "in")
  )
ggplot2::ggsave(
  plt, 
  filename = file.path(FIG_DIR, "sim_subgroup_errors_cov_with_pruning.pdf"),
  height = 9, 
  width = 14
)
```

### Selected Subgroup Features + Estimated Subgroup CATEs (max depth = 2)

```{r}
KEEP_METHODS <- c(
  "Distilled Causal Forest",
  "Distilled Rboost",
  "Virtual Twins",
  "Causal Tree",
  "Linear Regression",
  "Lasso"
)

## without covariates
sim_names <- c(
  "AND DGP",
  "Additive DGP",
  "OR DGP"
)
plt <- purrr::map(
  sim_names,
  function(sim_name) {
    plt_name <- stringr::str_replace(sim_name, "with ", "with\n")
    plt <- viz_results_ls[[sim_name]]$`Subgroup Feature Selection Errors Plot (max depth = 2)`
    plt$data <- plt$data |>
      dplyr::filter(
        .method_name %in% KEEP_METHODS
      )
    plt_selection <- plt +
      ggplot2::scale_color_manual(
        breaks = KEEP_METHODS,
        values = COLORS
      ) +
      ggplot2::scale_fill_manual(
        breaks = KEEP_METHODS,
        values = COLORS
      )

    plt <- viz_results_ls[[sim_name]]$`Subgroup CATE Errors Plot (max depth = 2)`
    plt$data <- plt$data |>
      dplyr::filter(
        .method_name %in% KEEP_METHODS
      )
    plt_cate <- plt +
      ggplot2::scale_color_manual(
        breaks = KEEP_METHODS,
        values = COLORS
      ) +
      ggplot2::scale_fill_manual(
        breaks = KEEP_METHODS,
        values = COLORS
      ) +
      ggplot2::labs(
        y = "CATE RMSE"
      ) +
      ggplot2::theme(
        strip.background = ggplot2::element_blank(),
        strip.placement = "outside",
        axis.title.y = ggplot2::element_text(face = "bold", size = 16),
        axis.text = ggplot2::element_text(size = 12),
        legend.title = ggplot2::element_text(face = "bold", size = 16),
        legend.text = ggplot2::element_text(size = 12)
      )
    plt <- patchwork::wrap_plots(
      plt_selection, plt_cate, nrow = 1, guides = "collect", widths = c(3.5, 1)
    ) &
      ggplot2::labs(x = NULL)

    if (sim_name != sim_names[2]) {
      plt <- plt &
        ggplot2::guides(
          color = "none", fill = "none"
        )
      hjust <- 0.78
    } else {
      hjust <- 0.82
    }
    if (sim_name == sim_names[1]) {
      plt[[1]] <- plt[[1]] +
        ggplot2::labs(
          title = "(A) Selected Subgroup Features\n(max depth = 2)"
        )
      plt[[2]] <- plt[[2]] +
        ggplot2::labs(
          title = "(B) Estimated Subgroup CATEs\n(max depth = 2)"
        )
    }
    plt[[1]] <- plt[[1]] +
      ggplot2::labs(
        tag = plt_name
      ) +
      ggplot2::theme(
        plot.tag = ggplot2::element_text(
          size = 18, face = "italic", angle = 90, hjust = 0.5, vjust = 1
        ),
        plot.tag.position = c(-.03, 0.48),
        plot.margin = ggplot2::unit(c(0, 0.35, 0, 0.5), units = "in")
      )
    plt <- plt &
      ggplot2::theme(
        panel.background = ggplot2::element_rect(fill = "grey98"),
        panel.grid.major = ggplot2::element_line(color = "grey98"),
        plot.title = ggplot2::element_text(
          size = 18, face = "italic", hjust = 0.5, vjust = -1
        )
      )
    return(plt)
  }
) |>
  patchwork::wrap_plots(
    ncol = 1
  ) |>
  patchwork::wrap_elements() +
  ggplot2::labs(
    tag = expression(
      bold(paste("Proportion of Variance Explained in ", tau, " by Covariates"))
    )
  ) +
  ggplot2::theme(
    plot.tag = ggplot2::element_text(
      size = 16
    ),
    plot.tag.position = c(0.5, 0),
    plot.margin = ggplot2::unit(c(-0.1, 0, .2, -0.1), units = "in")
  )
ggplot2::ggsave(
  plt, 
  filename = file.path(FIG_DIR, "sim_subgroup_errors_max_depth2.pdf"),
  height = 9.2, 
  width = 14
)

## with covariates
sim_names <- c(
  "AND DGP with Linear Covariates",
  "Additive DGP with Linear Covariates",
  "OR DGP with Linear Covariates"
)
plt <- purrr::map(
  sim_names,
  function(sim_name) {
    plt_name <- stringr::str_replace(sim_name, "with ", "with\n")
    plt <- viz_results_ls[[sim_name]]$`Subgroup Feature Selection Errors Plot (max depth = 2)`
    plt$data <- plt$data |>
      dplyr::filter(
        .method_name %in% KEEP_METHODS
      )
    plt_selection <- plt +
      ggplot2::scale_color_manual(
        breaks = KEEP_METHODS,
        values = COLORS
      ) +
      ggplot2::scale_fill_manual(
        breaks = KEEP_METHODS,
        values = COLORS
      )

    plt <- viz_results_ls[[sim_name]]$`Subgroup CATE Errors Plot (max depth = 2)`
    plt$data <- plt$data |>
      dplyr::filter(
        .method_name %in% KEEP_METHODS
      )
    plt_cate <- plt +
      ggplot2::scale_color_manual(
        breaks = KEEP_METHODS,
        values = COLORS
      ) +
      ggplot2::scale_fill_manual(
        breaks = KEEP_METHODS,
        values = COLORS
      ) +
      ggplot2::labs(
        y = "CATE RMSE"
      ) +
      ggplot2::theme(
        strip.background = ggplot2::element_blank(),
        strip.placement = "outside",
        axis.title.y = ggplot2::element_text(face = "bold", size = 16),
        axis.text = ggplot2::element_text(size = 12),
        legend.title = ggplot2::element_text(face = "bold", size = 16),
        legend.text = ggplot2::element_text(size = 12)
      )
    plt <- patchwork::wrap_plots(
      plt_selection, plt_cate, nrow = 1, guides = "collect", widths = c(3.5, 1)
    ) &
      ggplot2::labs(x = NULL)

    if (sim_name != sim_names[2]) {
      plt <- plt &
        ggplot2::guides(
          color = "none", fill = "none"
        )
      hjust <- 0.78
    } else {
      hjust <- 0.82
    }
    if (sim_name == sim_names[1]) {
      plt[[1]] <- plt[[1]] +
        ggplot2::labs(
          title = "(A) Selected Subgroup Features\n(max depth = 2)"
        )
      plt[[2]] <- plt[[2]] +
        ggplot2::labs(
          title = "(B) Estimated Subgroup CATEs\n(max depth = 2)"
        )
    }
    plt[[1]] <- plt[[1]] +
      ggplot2::labs(
        tag = plt_name
      ) +
      ggplot2::theme(
        plot.tag = ggplot2::element_text(
          size = 18, face = "italic", angle = 90, hjust = 0.5, vjust = 1
        ),
        plot.tag.position = c(-.06, 0.48),
        plot.margin = ggplot2::unit(c(0, 0.35, 0, 0.75), units = "in")
      )
    plt <- plt &
      ggplot2::theme(
        panel.background = ggplot2::element_rect(fill = "grey98"),
        panel.grid.major = ggplot2::element_line(color = "grey98"),
        plot.title = ggplot2::element_text(
          size = 18, face = "italic", hjust = 0.5, vjust = -1
        )
      )
    return(plt)
  }
) |>
  patchwork::wrap_plots(
    ncol = 1
  ) |>
  patchwork::wrap_elements() +
  ggplot2::labs(
    tag = expression(
      bold(paste("Proportion of Variance Explained in ", tau, " by Covariates"))
    )
  ) +
  ggplot2::theme(
    plot.tag = ggplot2::element_text(
      size = 16
    ),
    plot.tag.position = c(0.45, 0),
    plot.margin = ggplot2::unit(c(-0.1, 0, .2, -0.1), units = "in")
  )
ggplot2::ggsave(
  plt, 
  filename = file.path(FIG_DIR, "sim_subgroup_errors_cov_max_depth2.pdf"),
  height = 9.2, 
  width = 14
)
```

### Thresholds Distances

```{r}
KEEP_METHODS <- c(
  "Distilled Causal Forest",
  "Distilled Rboost",
  "Virtual Twins",
  "Causal Tree",
  "Linear Regression",
  "Lasso"
)

sim_names <- c(
  "AND DGP",
  "Additive DGP",
  "OR DGP",
  "AND DGP with Linear Covariates",
  "Additive DGP with Linear Covariates",
  "OR DGP with Linear Covariates"
)
plt <- purrr::imap(
  sim_names,
  function(sim_name, idx) {
    tag <- LETTERS[idx]
    plt <- viz_results_ls[[sim_name]]$`Threshold Distances Plot`
    plt$data <- plt$data |>
      dplyr::filter(
        .method_name %in% KEEP_METHODS
      )
    plt <- plt +
      ggplot2::scale_color_manual(
        breaks = KEEP_METHODS,
        values = COLORS
      ) +
      ggplot2::scale_fill_manual(
        breaks = KEEP_METHODS,
        values = COLORS
      ) +
      ggplot2::labs(
        title = sprintf("(%s) %s", tag, sim_name),
        y = "Average  | Estimated Threshold - True Threshold |",
        x = expression(
          bold(
            paste("Proportion of Variance Explained in ", tau, " by Covariates")
          )
        )
      ) +
      ggplot2::theme(
        panel.background = ggplot2::element_rect(fill = "grey98"),
        panel.grid.major = ggplot2::element_line(color = "grey98"),
        plot.title = ggplot2::element_text(
          size = 18, face = "italic", hjust = 0.5
        ),
        axis.title = ggplot2::element_text(face = "bold", size = 16),
        axis.text = ggplot2::element_text(size = 12),
        legend.text = ggplot2::element_text(size = 12),
        legend.title = ggplot2::element_text(size = 16),
        strip.text = ggplot2::element_text(size = 14)
      )
    if (sim_name != sim_names[5]) {
      plt <- plt +
        ggplot2::guides(
          color = "none", fill = "none"
        )
    }
    if (stringr::str_detect(sim_name, "Linear")) {
      plt <- plt + 
        ggplot2::theme(
          axis.title.y = ggplot2::element_blank(),
          plot.margin = ggplot2::unit(c(0, 0, 0.1, 0.2), units = "in")
        )
    } else {
      plt <- plt +
        ggplot2::theme(
          plot.margin = ggplot2::unit(c(0, 0.2, 0.1, 0), units = "in")
        )
    }
    if (!stringr::str_detect(sim_name, "OR")) {
      plt <- plt + 
        ggplot2::theme(
          axis.title.x = ggplot2::element_blank()
        )
    }
    return(plt)
  }
) |>
  patchwork::wrap_plots(
    ncol = 2, byrow = FALSE
  ) +
  patchwork::plot_layout(
    axis_titles = "collect_y"
  )
ggplot2::ggsave(
  plt, 
  filename = file.path(
    FIG_DIR, sprintf("sim_threshold_distances.pdf")
  ),
  height = 9, 
  width = 14
)
```

```{r}
KEEP_METHODS <- c(
  "Distilled Causal Forest",
  "Distilled Causal Forest (unpruned)",
  "Distilled Rboost",
  "Distilled Rboost (unpruned)",
  "Causal Tree",
  "Causal Tree (unpruned)"
)

sim_names <- c(
  "AND DGP",
  "Additive DGP",
  "OR DGP",
  "AND DGP with Linear Covariates",
  "Additive DGP with Linear Covariates",
  "OR DGP with Linear Covariates"
)
plt <- purrr::imap(
  sim_names,
  function(sim_name, idx) {
    tag <- LETTERS[idx]
    plt <- viz_results_ls[[sim_name]]$`Threshold Distances Plot`
    plt$data <- plt$data |>
      dplyr::filter(
        .method_name %in% KEEP_METHODS
      ) |> 
      add_pruned_column()
    plt$layers[[1]]$mapping <- ggplot2::aes(
      x = tau_heritability, 
      ymin = mean_threshold_dist - se_threshold_dist,
      ymax = mean_threshold_dist + se_threshold_dist,
      fill = .method_name,
      group = interaction(.method_name, .is_pruned)
    )
    plt$layers[[2]]$mapping <- ggplot2::aes(
      x = tau_heritability, 
      y = mean_threshold_dist,
      color = .method_name,
      linetype = .is_pruned, 
      group = interaction(.method_name, .is_pruned)
    )
    plt <- plt +
      ggplot2::scale_color_manual(
        breaks = KEEP_METHODS,
        values = COLORS
      ) +
      ggplot2::scale_fill_manual(
        breaks = KEEP_METHODS,
        values = COLORS
      ) +
      ggplot2::scale_linetype_manual(
        values = LINETYPES
      ) +
      ggplot2::labs(
        title = sprintf("(%s) %s", tag, sim_name),
        y = "Average  | Estimated Threshold - True Threshold |",
        x = expression(
          bold(
            paste("Proportion of Variance Explained in ", tau, " by Covariates")
          )
        ),
        linetype = "Pruning Mode"
      ) +
      ggplot2::theme(
        panel.background = ggplot2::element_rect(fill = "grey98"),
        panel.grid.major = ggplot2::element_line(color = "grey98"),
        plot.title = ggplot2::element_text(
          size = 18, face = "italic", hjust = 0.5
        ),
        axis.title = ggplot2::element_text(face = "bold", size = 16),
        axis.text = ggplot2::element_text(size = 12),
        legend.text = ggplot2::element_text(size = 12),
        legend.title = ggplot2::element_text(size = 16),
        strip.text = ggplot2::element_text(size = 14)
      )
    if (sim_name != sim_names[5]) {
      plt <- plt +
        ggplot2::guides(
          color = "none", fill = "none", linetype = "none"
        )
    }
    if (stringr::str_detect(sim_name, "Linear")) {
      plt <- plt + 
        ggplot2::theme(
          axis.title.y = ggplot2::element_blank(),
          plot.margin = ggplot2::unit(c(0, 0, 0.1, 0.2), units = "in")
        )
    } else {
      plt <- plt +
        ggplot2::theme(
          plot.margin = ggplot2::unit(c(0, 0.2, 0.1, 0), units = "in")
        )
    }
    if (!stringr::str_detect(sim_name, "OR")) {
      plt <- plt + 
        ggplot2::theme(
          axis.title.x = ggplot2::element_blank()
        )
    }
    return(plt)
  }
) |>
  patchwork::wrap_plots(
    ncol = 2, byrow = FALSE
  ) +
  patchwork::plot_layout(
    axis_titles = "collect_y"
  )
ggplot2::ggsave(
  plt, 
  filename = file.path(
    FIG_DIR, sprintf("sim_threshold_distances_with_pruning.pdf")
  ),
  height = 9, 
  width = 14
)
```

### Thresholds Distances (max depth = 2)

```{r}
KEEP_METHODS <- c(
  "Distilled Causal Forest",
  "Distilled Rboost",
  "Virtual Twins",
  "Causal Tree",
  "Linear Regression",
  "Lasso"
)

sim_names <- c(
  "AND DGP",
  "Additive DGP",
  "OR DGP",
  "AND DGP with Linear Covariates",
  "Additive DGP with Linear Covariates",
  "OR DGP with Linear Covariates"
)
plt <- purrr::imap(
  sim_names,
  function(sim_name, idx) {
    tag <- LETTERS[idx]
    plt <- viz_results_ls[[sim_name]]$`Threshold Distances Plot (max depth = 2)`
    plt$data <- plt$data |>
      dplyr::filter(
        .method_name %in% KEEP_METHODS
      )
    plt <- plt +
      ggplot2::scale_color_manual(
        breaks = KEEP_METHODS,
        values = COLORS
      ) +
      ggplot2::scale_fill_manual(
        breaks = KEEP_METHODS,
        values = COLORS
      ) +
      ggplot2::labs(
        title = sprintf("(%s) %s\n", tag, sim_name),
        y = "Average  | Estimated Threshold - True Threshold |\n(max depth = 2)",
        x = expression(
          bold(
            paste("Proportion of Variance Explained in ", tau, " by Covariates")
          )
        )
      ) +
      ggplot2::theme(
        panel.background = ggplot2::element_rect(fill = "grey98"),
        panel.grid.major = ggplot2::element_line(color = "grey98"),
        plot.title = ggplot2::element_text(
          size = 18, face = "italic", hjust = 0.5
        ),
        axis.title = ggplot2::element_text(face = "bold", size = 16),
        axis.text = ggplot2::element_text(size = 12),
        legend.text = ggplot2::element_text(size = 12),
        legend.title = ggplot2::element_text(size = 16),
        strip.text = ggplot2::element_text(size = 14)
      )
    if (sim_name != sim_names[5]) {
      plt <- plt +
        ggplot2::guides(
          color = "none", fill = "none"
        )
    }
    if (stringr::str_detect(sim_name, "Linear")) {
      plt <- plt + 
        ggplot2::theme(
          axis.title.y = ggplot2::element_blank(),
          plot.margin = ggplot2::unit(c(0, 0, 0.1, 0.2), units = "in")
        )
    } else {
      plt <- plt +
        ggplot2::theme(
          plot.margin = ggplot2::unit(c(0, 0.2, 0.1, 0), units = "in")
        )
    }
    if (!stringr::str_detect(sim_name, "OR")) {
      plt <- plt + 
        ggplot2::theme(
          axis.title.x = ggplot2::element_blank()
        )
    }
    return(plt)
  }
) |>
  patchwork::wrap_plots(
    ncol = 2, byrow = FALSE
  ) +
  patchwork::plot_layout(
    axis_titles = "collect_y"
  )
ggplot2::ggsave(
  plt, 
  filename = file.path(
    FIG_DIR, sprintf("sim_threshold_distances_max_depth2.pdf")
  ),
  height = 9, 
  width = 14.2
)
```

### Threshold Distributions

```{r}
KEEP_METHODS <- c(
  "Distilled Causal Forest",
  "Distilled Rboost",
  "Virtual Twins",
  "Causal Tree",
  "Linear Regression",
  "Lasso"
)
KEEP_HERITABILITY <- c(0.2, 0.6, 1)

# without covariates
sim_names <- c(
  "AND DGP",
  "Additive DGP",
  "OR DGP"
)
plt <- purrr::imap(
  sim_names,
  function(sim_name, idx) {
    tag <- LETTERS[idx]
    plt <- viz_results_ls[[sim_name]]$`Subgroup Thresholds Distribution Plot`
    plt$data <- plt$data |>
      dplyr::filter(
        .method_name %in% KEEP_METHODS,
        tau_heritability %in% KEEP_HERITABILITY
      )
    plt <- plt +
      ggplot2::scale_color_manual(
        breaks = KEEP_METHODS,
        values = COLORS
      ) +
      ggplot2::scale_fill_manual(
        breaks = KEEP_METHODS,
        values = COLORS
      ) +
      ggplot2::facet_grid(.var ~ tau_heritability, labeller = signal_labeller) +
      ggplot2::labs(
        title = sprintf("(%s) %s", tag, sim_name)
      ) +
      ggplot2::theme(
        panel.background = ggplot2::element_rect(fill = "grey98"),
        panel.grid.major = ggplot2::element_line(color = "grey98"),
        plot.title = ggplot2::element_text(
          size = 18, face = "italic", hjust = 0.5
        )
      )
    return(plt)
  }
) |>
  patchwork::wrap_plots(
    ncol = 1, guides = "collect"
  ) +
  patchwork::plot_layout(
    axes = "collect"
  )
ggplot2::ggsave(
  plt, 
  filename = file.path(FIG_DIR, "sim_thresholds.pdf"),
  height = 10, 
  width = 11
)

# with covariates
sim_names <- c(
  "AND DGP with Linear Covariates",
  "Additive DGP with Linear Covariates",
  "OR DGP with Linear Covariates"
)
plt <- purrr::imap(
  sim_names,
  function(sim_name, idx) {
    tag <- LETTERS[idx]
    plt <- viz_results_ls[[sim_name]]$`Subgroup Thresholds Distribution Plot`
    plt$data <- plt$data |>
      dplyr::filter(
        .method_name %in% KEEP_METHODS,
        tau_heritability %in% KEEP_HERITABILITY
      )
    plt <- plt +
      ggplot2::scale_color_manual(
        breaks = KEEP_METHODS,
        values = COLORS
      ) +
      ggplot2::scale_fill_manual(
        breaks = KEEP_METHODS,
        values = COLORS
      ) +
      ggplot2::facet_grid(.var ~ tau_heritability, labeller = signal_labeller) +
      ggplot2::labs(
        title = sprintf("(%s) %s", tag, sim_name)
      ) +
      ggplot2::theme(
        panel.background = ggplot2::element_rect(fill = "grey98"),
        panel.grid.major = ggplot2::element_line(color = "grey98"),
        plot.title = ggplot2::element_text(
          size = 18, face = "italic", hjust = 0.5
        )
      )
    return(plt)
  }
) |>
  patchwork::wrap_plots(
    ncol = 1, guides = "collect"
  ) +
  patchwork::plot_layout(
    axes = "collect"
  )
ggplot2::ggsave(
  plt, 
  filename = file.path(FIG_DIR, "sim_thresholds_cov.pdf"),
  height = 10, 
  width = 11
)
```

### Threshold Distributions (max depth = 2)

```{r}
KEEP_METHODS <- c(
  "Distilled Causal Forest",
  "Distilled Rboost",
  "Virtual Twins",
  "Causal Tree",
  "Linear Regression",
  "Lasso"
)
KEEP_HERITABILITY <- c(0.2, 0.6, 1)

# without covariates
sim_names <- c(
  "AND DGP",
  "Additive DGP",
  "OR DGP"
)
plt <- purrr::imap(
  sim_names,
  function(sim_name, idx) {
    tag <- LETTERS[idx]
    plt <- viz_results_ls[[sim_name]]$`Subgroup Thresholds Distribution Plot (max depth = 2)`
    plt$data <- plt$data |>
      dplyr::filter(
        .method_name %in% KEEP_METHODS,
        tau_heritability %in% KEEP_HERITABILITY
      )
    plt <- plt +
      ggplot2::scale_color_manual(
        breaks = KEEP_METHODS,
        values = COLORS
      ) +
      ggplot2::scale_fill_manual(
        breaks = KEEP_METHODS,
        values = COLORS
      ) +
      ggplot2::facet_grid(.var ~ tau_heritability, labeller = signal_labeller) +
      ggplot2::labs(
        y = "Threshold (max depth = 2)",
        title = sprintf("(%s) %s", tag, sim_name)
      ) +
      ggplot2::theme(
        panel.background = ggplot2::element_rect(fill = "grey98"),
        panel.grid.major = ggplot2::element_line(color = "grey98"),
        plot.title = ggplot2::element_text(
          size = 18, face = "italic", hjust = 0.5
        )
      )
    return(plt)
  }
) |>
  patchwork::wrap_plots(
    ncol = 1, guides = "collect"
  ) +
  patchwork::plot_layout(
    axes = "collect"
  )
ggplot2::ggsave(
  plt, 
  filename = file.path(FIG_DIR, "sim_thresholds_max_depth2.pdf"),
  height = 10, 
  width = 11
)

# with covariates
sim_names <- c(
  "AND DGP with Linear Covariates",
  "Additive DGP with Linear Covariates",
  "OR DGP with Linear Covariates"
)
plt <- purrr::imap(
  sim_names,
  function(sim_name, idx) {
    tag <- LETTERS[idx]
    plt <- viz_results_ls[[sim_name]]$`Subgroup Thresholds Distribution Plot (max depth = 2)`
    plt$data <- plt$data |>
      dplyr::filter(
        .method_name %in% KEEP_METHODS,
        tau_heritability %in% KEEP_HERITABILITY
      )
    plt <- plt +
      ggplot2::scale_color_manual(
        breaks = KEEP_METHODS,
        values = COLORS
      ) +
      ggplot2::scale_fill_manual(
        breaks = KEEP_METHODS,
        values = COLORS
      ) +
      ggplot2::facet_grid(.var ~ tau_heritability, labeller = signal_labeller) +
      ggplot2::labs(
        y = "Threshold (max depth = 2)",
        title = sprintf("(%s) %s", tag, sim_name)
      ) +
      ggplot2::theme(
        panel.background = ggplot2::element_rect(fill = "grey98"),
        panel.grid.major = ggplot2::element_line(color = "grey98"),
        plot.title = ggplot2::element_text(
          size = 18, face = "italic", hjust = 0.5
        )
      )
    return(plt)
  }
) |>
  patchwork::wrap_plots(
    ncol = 1, guides = "collect"
  ) +
  patchwork::plot_layout(
    axes = "collect"
  )
ggplot2::ggsave(
  plt, 
  filename = file.path(FIG_DIR, "sim_thresholds_cov_max_depth2.pdf"),
  height = 10, 
  width = 11
)
```

### Subgroup Stability

```{r}
KEEP_METHODS <- c(
  "Distilled Causal Forest",
  "Distilled Rboost",
  "Causal Tree"
)
KEEP_HERITABILITY <- c(0.2, 0.6, 1)

sim_names <- c(
  "AND DGP",
  "Additive DGP",
  "OR DGP",
  "AND DGP with Linear Covariates",
  "Additive DGP with Linear Covariates",
  "OR DGP with Linear Covariates"
)

plt <- purrr::map(
  sim_names,
  function(sim_name) {
    f1_df <- viz_results_ls[[sim_name]]$`Subgroup Feature Selection Errors Plot`$data |> 
      dplyr::filter(
        .method_name %in% KEEP_METHODS,
        tau_heritability %in% KEEP_HERITABILITY,
        .metric == "F1"
      ) |> 
      dplyr::mutate(
        .value = mean_feature_selection_err
      )
    cate_df <- viz_results_ls[[sim_name]]$`Subgroup CATE Errors Plot`$data |>
      dplyr::filter(
        .method_name %in% KEEP_METHODS,
        tau_heritability %in% KEEP_HERITABILITY,
        .metric == "rmse"
      ) |> 
      dplyr::mutate(
        .metric = "CATE RMSE",
        .value = mean_cate_err
      )

    plt_err <- dplyr::bind_rows(f1_df, cate_df) |> 
      dplyr::mutate(
        .metric = forcats::fct_inorder(.metric)
      ) |> 
      ggplot2::ggplot() +
      ggplot2::aes(
        x = .method_name,
        y = .value, 
        fill = .method_name
      ) +
      ggplot2::geom_bar(
        stat = "identity", size = 1, width = 0.75
      ) +
      ggplot2::scale_color_manual(
        breaks = KEEP_METHODS,
        values = COLORS
      ) +
      ggplot2::scale_fill_manual(
        breaks = KEEP_METHODS,
        values = COLORS
      ) +
      ggplot2::facet_grid(
        .metric ~ tau_heritability, labeller = signal_labeller, 
        scales = "free_y", switch = "y"
      ) +
      ggplot2::scale_x_discrete(
        breaks = KEEP_METHODS,
        labels = stringr::str_replace_all(KEEP_METHODS, " ", "\n")
      ) +
      ggplot2::labs(y = "", x = "Method") +
      ggplot2::guides(color = "none", fill = "none") +
      vthemes::theme_vmodern(size_preset = "medium", grid_color = "grey98") +
      ggplot2::theme(
        strip.background.y = ggplot2::element_blank(),
        strip.text.y = ggplot2::element_text(
          color = "black", size = 16
        ),
        strip.placement = "outside",
        axis.title.y = ggplot2::element_blank(),
        axis.text.y = ggplot2::element_text(size = 9)
      )

    plt <- viz_results_ls[[sim_name]]$`Subgroup Stability Diagnostics Plot`[[1]]
    plt$data <- plt$data |> 
      dplyr::filter(
        .method_name %in% KEEP_METHODS,
        tau_heritability %in% KEEP_HERITABILITY
      ) |> 
      dplyr::mutate(
        .metric = "Jaccard SSI"
      )
    plt_jaccard <- plt +
      ggplot2::geom_line(size = 0.75) +
      ggplot2::scale_color_manual(
        breaks = KEEP_METHODS,
        values = COLORS
      ) +
      ggplot2::scale_fill_manual(
        breaks = KEEP_METHODS,
        values = COLORS
      ) +
      ggplot2::facet_grid(
        ~ tau_heritability, labeller = signal_labeller, switch = "y"
      ) +
      ggplot2::coord_cartesian(xlim = c(1, 4)) +
      ggplot2::scale_x_continuous(
        breaks = 1:4
      ) +
      ggplot2::scale_y_continuous(
        limits = c(0, 1), breaks = seq(0, 1, by = 0.2)
      ) +
      ggplot2::labs(
        x = "Tree Depth",
        y = "Jaccard SSI",
        color = "Method",
        fill = "Method"
      ) +
      vthemes::theme_vmodern(size_preset = "medium") +
      ggplot2::theme(
        strip.background.y = ggplot2::element_blank(),
        strip.text.y = ggplot2::element_text(
          color = "black", size = 16
        ),
        strip.placement = "outside",
        axis.title.y = ggplot2::element_blank(),
        axis.text.y = ggplot2::element_text(size = 9)
      )

    plt <- patchwork::wrap_plots(
      plt_jaccard, plt_err,
      ncol = 1, heights = c(1.2, 2), guides = "collect"
    ) +
      patchwork::plot_layout(
        axis_titles = "collect_x"
      ) +
      patchwork::plot_annotation(
        title = sim_name, 
        tag_levels = "A", 
        tag_prefix = "(",
        tag_suffix = ")",
        theme = ggplot2::theme(
          plot.title = ggplot2::element_text(
            size = 18, face = "italic", hjust = 0.42
          )
        )
      ) &
      ggplot2::theme(
        plot.tag = ggplot2::element_text(size = 16)#,
        # legend.position = "bottom"  # for slides
      )
    # ggplot2::ggsave(
    #   plt, 
    #   filename = file.path(
    #     FIG_DIR, 
    #     sprintf("sim_stability_%s_slides.pdf", sim_name)
    #   ),
    #   width = 11.5,
    #   height = 9
    # )
    ggplot2::ggsave(
      plt, 
      filename = file.path(
        FIG_DIR,
        sprintf("sim_stability_%s.pdf", sim_name)
      ),
      width = 10.5,
      height = 7.5
    )
  }
)
```

```{r}
KEEP_METHODS <- c(
  "Distilled Causal Forest",
  "Distilled Rboost",
  "Causal Tree"
)
KEEP_HERITABILITY <- c(0.2, 0.6, 1)

sim_names <- c(
  "AND DGP",
  "Additive DGP",
  "OR DGP",
  "AND DGP with Linear Covariates",
  "Additive DGP with Linear Covariates",
  "OR DGP with Linear Covariates"
)

plt <- purrr::map(
  sim_names,
  function(sim_name) {
    plt_df <- viz_results_ls[[sim_name]]$`Subgroup Feature Selection Errors Plot`$data |> 
      dplyr::filter(
        .method_name %in% KEEP_METHODS,
        tau_heritability %in% KEEP_HERITABILITY,
        .metric == "F1"
      )
    plt_f1 <- plt_df |> 
      ggplot2::ggplot() +
      ggplot2::aes(
        x = .method_name,
        y = mean_feature_selection_err,
        fill = .method_name
      ) +
      ggplot2::geom_bar(stat = "identity", size = 1) +
      ggplot2::scale_color_manual(
        breaks = KEEP_METHODS,
        values = COLORS
      ) +
      ggplot2::scale_fill_manual(
        breaks = KEEP_METHODS,
        values = COLORS
      ) +
      ggplot2::facet_grid(
        tau_heritability ~ ., labeller = signal_labeller, scales = "free_y"
      ) +
      ggplot2::labs(y = "F1", x = "Method") +
      ggplot2::guides(color = "none", fill = "none") +
      vthemes::theme_vmodern(size_preset = "medium", grid_color = "grey98") +
      ggplot2::theme(
        strip.background = ggplot2::element_blank(),
        strip.text = ggplot2::element_blank(),
        axis.text.x = ggplot2::element_text(angle = 90, hjust = 1, vjust = 0.5)
      )

    plt_df <- viz_results_ls[[sim_name]]$`Subgroup CATE Errors Plot`$data |>
      dplyr::filter(
        .method_name %in% KEEP_METHODS,
        tau_heritability %in% KEEP_HERITABILITY,
        .metric == "rmse"
      )
    plt_cate <- plt_df |> 
      ggplot2::ggplot() +
      ggplot2::aes(
        x = .method_name,
        y = mean_cate_err,
        fill = .method_name
      ) +
      ggplot2::geom_bar(stat = "identity", size = 1) +
      ggplot2::scale_color_manual(
        breaks = KEEP_METHODS,
        values = COLORS
      ) +
      ggplot2::scale_fill_manual(
        breaks = KEEP_METHODS,
        values = COLORS
      ) +
      ggplot2::facet_grid(
        tau_heritability ~ ., labeller = signal_labeller, scales = "free_y"
      ) +
      ggplot2::labs(y = "CATE RMSE", x = "Method") +
      ggplot2::guides(color = "none", fill = "none") +
      vthemes::theme_vmodern(size_preset = "medium", grid_color = "grey98") +
      ggplot2::theme(
        strip.background = ggplot2::element_blank(),
        strip.text = ggplot2::element_blank(),
        axis.text.x = ggplot2::element_text(angle = 90, hjust = 1, vjust = 0.5)
      )

    plt <- viz_results_ls[[sim_name]]$`Subgroup Stability Diagnostics Plot`[[1]]
    plt$data <- plt$data |> 
      dplyr::filter(
        .method_name %in% KEEP_METHODS,
        tau_heritability %in% KEEP_HERITABILITY
      )
    plt_jaccard <- plt +
      ggplot2::geom_line(size = 0.75) +
      ggplot2::scale_color_manual(
        breaks = KEEP_METHODS,
        values = COLORS
      ) +
      ggplot2::scale_fill_manual(
        breaks = KEEP_METHODS,
        values = COLORS
      ) +
      ggplot2::facet_grid(tau_heritability ~ ., labeller = signal_labeller) +
      ggplot2::coord_cartesian(xlim = c(1, 4)) +
      ggplot2::scale_x_continuous(
        breaks = 1:4
      ) +
      ggplot2::scale_y_continuous(
        limits = c(0, 1), breaks = seq(0, 1, by = 0.2)
      ) +
      ggplot2::labs(
        x = "Tree Depth",
        y = "Jaccard SSI",
        color = "Method",
        fill = "Method"
      ) +
      vthemes::theme_vmodern(size_preset = "medium")

    plt <- viz_results_ls[[sim_name]]$`Subgroup Stability Diagnostics Plot`[[2]]
    plt$data <- plt$data |> 
      dplyr::filter(
        .method_name %in% KEEP_METHODS,
        tau_heritability %in% KEEP_HERITABILITY
      )
    if (stringr::str_detect(sim_name, "Covariates")) {
      FT_COLORS <- c(
        "#d3831b", "#eeb062",# "#e19e49"
        "#CCCCCC", "#B5B5B5", "#9E9E9E", "#888888",
        "#717171", "#5A5A5A", "#434343", "#2D2D2D"
      )
      # FT_COLORS <- c(
      #   "#d07d13", "#eeb062",
      #   # "#FFa59B", "#FFD2CC",
      #   "#B34363", "#cb8c9d",
      #   "#9E9E9E", "#888888",
      #   "#717171", "#5A5A5A", "#434343", "#2D2D2D"
      # )
    } else {
      FT_COLORS <- c(
        "#d3831b", "#eeb062",# "#e19e49"
        "#CCCCCC", "#B5B5B5", "#9E9E9E", "#888888",
        "#717171", "#5A5A5A", "#434343", "#2D2D2D"
      )
    }
    plt_stability_feature <- plt +
      ggplot2::scale_fill_manual(values = FT_COLORS) +
      ggplot2::facet_grid(tau_heritability ~ depth, labeller = signal_labeller) +
      ggplot2::labs(
        x = "Method",
        y = "Proportion of Splits",
        fill = "Variable"
      ) +
      vthemes::theme_vmodern(size_preset = "medium") +
      ggplot2::theme(
        axis.text.x = ggplot2::element_text(angle = 90, hjust = 1, vjust = 0.5)
      )

    plt <- patchwork::wrap_plots(
      plt_f1, plt_cate, plt_jaccard, plt_stability_feature,
      nrow = 1, widths = c(0.5, 0.5, 1.25, 3.25), guides = "collect"
    ) |> 
      patchwork::wrap_elements() +
      ggplot2::labs(
        title = "   (A) Subgroup Estimation       (B) Jaccard Stability                              (C) Feature Stability",
        tag = sim_name
      ) +
      ggplot2::theme(
        plot.tag = ggplot2::element_text(
          size = 20, face = "bold", hjust = 0.5
        ),
        plot.tag.position = c(0.47, 1.02),
        plot.margin = ggplot2::unit(c(0.32, 0, 0, 0.2), units = "in"),
        plot.title = ggplot2::element_text(
          size = 18, face = "italic", hjust = 0, vjust = -0.3
        )
      )

    ggplot2::ggsave(
      plt, 
      filename = file.path(
        FIG_DIR, 
        sprintf("sim_stability_feature_%s.pdf", sim_name)
      ),
      height = 8, 
      width = 14
    )
  }
)
```

### Number of Feature Splits Heatmap

```{r}
KEEP_METHODS <- c(
  "Distilled Causal Forest",
  "Distilled Causal Forest (unpruned)",
  "Distilled Rboost",
  "Distilled Rboost (unpruned)",
  "Virtual Twins",
  "Causal Tree",
  "Causal Tree (unpruned)",
  "Linear Regression",
  "Lasso"
)
KEEP_HERITABILITY <- c(0.2, 0.6, 1)

# without covariates
sim_names <- c(
  "AND DGP",
  "Additive DGP",
  "OR DGP"
)
plt <- purrr::imap(
  sim_names,
  function(sim_name, idx) {
    tag <- LETTERS[idx]
    plt <- viz_results_ls[[sim_name]]$`Number of Splits Plot`
    plt$data <- plt$data |>
      dplyr::filter(
        .method_name %in% KEEP_METHODS,
        tau_heritability %in% KEEP_HERITABILITY
      )
    plt_nsplits <- plt +
      ggplot2::facet_grid(tau_heritability ~ ., labeller = signal_labeller) +
      viridis::scale_fill_viridis(
        begin = 0.1, end = 1, option = "magma", limits = c(0., 4.5)
      )

    plt <- viz_results_ls[[sim_name]]$`Number of Trees Plot`
    plt$data <- plt$data |>
      dplyr::filter(
        .method_name %in% KEEP_METHODS,
        tau_heritability %in% KEEP_HERITABILITY
      )
    plt_ntrees <- plt +
      ggplot2::facet_grid(tau_heritability ~ ., labeller = signal_labeller) +
      viridis::scale_fill_viridis(
        begin = 0.1, end = 1, option = "magma", limits = c(0, 100)
      )

    plt <- patchwork::wrap_plots(plt_ntrees, plt_nsplits, nrow = 1) +
      patchwork::plot_layout(
        axis_titles = "collect_y"
      ) &
      ggplot2::labs(x = "Variable")
    plt <- patchwork::wrap_elements(plt) +
      ggplot2::labs(
        tag = sprintf("(%s) %s", tag, sim_name)
      ) +
      ggplot2::theme(
        plot.tag = ggplot2::element_text(
          size = 18, face = "italic", hjust = 0.5
        ),
        plot.tag.position = c(0.47, 1.02),
        plot.margin = ggplot2::unit(c(0.3, 0, 0, 0.5), units = "in")
      )
    return(plt)
  }
) |>
  patchwork::wrap_plots(
    ncol = 1
  )
ggplot2::ggsave(
  plt, 
  filename = file.path(FIG_DIR, "sim_freq.pdf"),
  height = 17.5, 
  width = 15
)

# with covariates
sim_names <- c(
  "AND DGP with Linear Covariates",
  "Additive DGP with Linear Covariates",
  "OR DGP with Linear Covariates"
)
plt <- purrr::imap(
  sim_names,
  function(sim_name, idx) {
    tag <- LETTERS[idx]
    plt <- viz_results_ls[[sim_name]]$`Number of Splits Plot`
    plt$data <- plt$data |>
      dplyr::filter(
        .method_name %in% KEEP_METHODS,
        tau_heritability %in% KEEP_HERITABILITY
      )
    plt_nsplits <- plt +
      ggplot2::facet_grid(tau_heritability ~ ., labeller = signal_labeller) +
      viridis::scale_fill_viridis(
        begin = 0.1, end = 1, option = "magma", limits = c(0., 4.5)
      )

    plt <- viz_results_ls[[sim_name]]$`Number of Trees Plot`
    plt$data <- plt$data |>
      dplyr::filter(
        .method_name %in% KEEP_METHODS,
        tau_heritability %in% KEEP_HERITABILITY
      )
    plt_ntrees <- plt +
      ggplot2::facet_grid(tau_heritability ~ ., labeller = signal_labeller) +
      viridis::scale_fill_viridis(
        begin = 0.1, end = 1, option = "magma", limits = c(0, 100)
      )

    plt <- patchwork::wrap_plots(plt_ntrees, plt_nsplits, nrow = 1) +
      patchwork::plot_layout(
        axis_titles = "collect_y"
      ) &
      ggplot2::labs(x = "Variable")
    plt <- patchwork::wrap_elements(plt) +
      ggplot2::labs(
        tag = sprintf("(%s) %s", tag, sim_name)
      ) +
      ggplot2::theme(
        plot.tag = ggplot2::element_text(
          size = 18, face = "italic", hjust = 0.5
        ),
        plot.tag.position = c(0.47, 1.02),
        plot.margin = ggplot2::unit(c(0.3, 0, 0, 0.5), units = "in")
      )
    return(plt)
  }
) |>
  patchwork::wrap_plots(
    ncol = 1
  )
ggplot2::ggsave(
  plt, 
  filename = file.path(FIG_DIR, "sim_freq_cov.pdf"),
  height = 17.5, 
  width = 15
)
```

### Number of Feature Splits Heatmap (max depth = 2)

```{r}
KEEP_METHODS <- c(
  "Distilled Causal Forest",
  "Distilled Rboost",
  "Virtual Twins",
  "Causal Tree",
  "Linear Regression",
  "Lasso"
)
KEEP_HERITABILITY <- c(0.2, 0.6, 1)

# without covariates
sim_names <- c(
  "AND DGP",
  "Additive DGP",
  "OR DGP"
)
plt <- purrr::imap(
  sim_names,
  function(sim_name, idx) {
    tag <- LETTERS[idx]
    plt <- viz_results_ls[[sim_name]]$`Number of Splits Plot (max depth = 2)`
    plt$data <- plt$data |>
      dplyr::filter(
        .method_name %in% KEEP_METHODS,
        tau_heritability %in% KEEP_HERITABILITY
      )
    plt_nsplits <- plt +
      ggplot2::facet_grid(tau_heritability ~ ., labeller = signal_labeller) +
      viridis::scale_fill_viridis(
        begin = 0.1, end = 1, option = "magma", limits = c(0., 4.5)
      ) +
      ggplot2::labs(fill = "Average #\nSplits per Tree\n(max depth = 2)")

    plt <- viz_results_ls[[sim_name]]$`Number of Trees Plot (max depth = 2)`
    plt$data <- plt$data |>
      dplyr::filter(
        .method_name %in% KEEP_METHODS,
        tau_heritability %in% KEEP_HERITABILITY
      )
    plt_ntrees <- plt +
      ggplot2::facet_grid(tau_heritability ~ ., labeller = signal_labeller) +
      viridis::scale_fill_viridis(
        begin = 0.1, end = 1, option = "magma", limits = c(0, 100)
      ) +
      ggplot2::labs(fill = "# Trees\n(max depth = 2)")

    plt <- patchwork::wrap_plots(plt_ntrees, plt_nsplits, nrow = 1) +
      patchwork::plot_layout(
        axis_titles = "collect_y"
      ) &
      ggplot2::labs(x = "Variable")
    plt <- patchwork::wrap_elements(plt) +
      ggplot2::labs(
        tag = sprintf("(%s) %s", tag, sim_name)
      ) +
      ggplot2::theme(
        plot.tag = ggplot2::element_text(
          size = 18, face = "italic", hjust = 0.5
        ),
        plot.tag.position = c(0.47, 1.02),
        plot.margin = ggplot2::unit(c(0.3, 0, 0, 0.5), units = "in")
      )
    return(plt)
  }
) |>
  patchwork::wrap_plots(
    ncol = 1
  )
ggplot2::ggsave(
  plt, 
  filename = file.path(FIG_DIR, "sim_freq_max_depth2.pdf"),
  height = 17.5, 
  width = 15
)

# with covariates
sim_names <- c(
  "AND DGP with Linear Covariates",
  "Additive DGP with Linear Covariates",
  "OR DGP with Linear Covariates"
)
plt <- purrr::imap(
  sim_names,
  function(sim_name, idx) {
    tag <- LETTERS[idx]
    plt <- viz_results_ls[[sim_name]]$`Number of Splits Plot (max depth = 2)`
    plt$data <- plt$data |>
      dplyr::filter(
        .method_name %in% KEEP_METHODS,
        tau_heritability %in% KEEP_HERITABILITY
      )
    plt_nsplits <- plt +
      ggplot2::facet_grid(tau_heritability ~ ., labeller = signal_labeller) +
      viridis::scale_fill_viridis(
        begin = 0.1, end = 1, option = "magma", limits = c(0., 4.5)
      ) +
      ggplot2::labs(fill = "Average #\nSplits per Tree\n(max depth = 2)")

    plt <- viz_results_ls[[sim_name]]$`Number of Trees Plot (max depth = 2)`
    plt$data <- plt$data |>
      dplyr::filter(
        .method_name %in% KEEP_METHODS,
        tau_heritability %in% KEEP_HERITABILITY
      )
    plt_ntrees <- plt +
      ggplot2::facet_grid(tau_heritability ~ ., labeller = signal_labeller) +
      viridis::scale_fill_viridis(
        begin = 0.1, end = 1, option = "magma", limits = c(0, 100)
      ) +
      ggplot2::labs(fill = "# Trees\n(max depth = 2)")

    plt <- patchwork::wrap_plots(plt_ntrees, plt_nsplits, nrow = 1) +
      patchwork::plot_layout(
        axis_titles = "collect_y"
      ) &
      ggplot2::labs(x = "Variable")
    plt <- patchwork::wrap_elements(plt) +
      ggplot2::labs(
        tag = sprintf("(%s) %s", tag, sim_name)
      ) +
      ggplot2::theme(
        plot.tag = ggplot2::element_text(
          size = 18, face = "italic", hjust = 0.5
        ),
        plot.tag.position = c(0.47, 1.02),
        plot.margin = ggplot2::unit(c(0.3, 0, 0, 0.5), units = "in")
      )
    return(plt)
  }
) |>
  patchwork::wrap_plots(
    ncol = 1
  )
ggplot2::ggsave(
  plt, 
  filename = file.path(FIG_DIR, "sim_freq_cov_max_depth2.pdf"),
  height = 17.5, 
  width = 15
)
```

## Varying \# Repeated Crossfits Simulations {.tabset .tabset-pills .tabset-pills-square}

```{r results = "asis"}
sim_results_dir <- here::here(
  file.path("results", "Varying nrep crossfits")
)
SIM_NAMES <- c(
  "AND DGP" = "AND DGP",
  "Additive DGP" = "Additive DGP",
  "OR DGP" = "OR DGP"
)

# load in results
eval_results_ls <- list()
viz_results_ls <- list()
for (sim_name in names(SIM_NAMES)) {
  fit_results <- readRDS(
    file.path(
      sim_results_dir, 
      paste0(sim_name, "-Distilled Causal Forest (crossfit)-Distilled Rboost (crossfit)"), 
      "Varying tau_heritability-nreps_crossfit", "fit_results.rds"
    )
  )
  eval_results <- readRDS(
    file.path(
      sim_results_dir, 
      paste0(sim_name, "-Distilled Causal Forest (crossfit)-Distilled Rboost (crossfit)"), 
      "Varying tau_heritability-nreps_crossfit", "eval_results.rds"
    )
  )
  eval_results <- purrr::map(
    eval_results,
    ~ .x |>
      dplyr::filter(
        !stringr::str_detect(.method_name, "unpruned"),
        !stringr::str_detect(.method_name, "Causal Forest")
      ) |>
      dplyr::mutate(
        .method_name = stringr::str_remove_all(.method_name, " \\(.*\\)"),
        nreps_crossfit = tidyr::replace_na(nreps_crossfit, 0) |>
          as.factor() |>
          forcats::fct_inseq()
      )
  )

  viz_results <- visualize_experiment(
    crossfit_experiment, fit_results, eval_results
  )
  eval_results_ls[[SIM_NAMES[[sim_name]]]] <- eval_results
  viz_results_ls[[SIM_NAMES[[sim_name]]]] <- viz_results
}
```

### Selected Subgroup Features + Estimated Subgroup CATEs

```{r}
## without covariates
sim_names <- c(
  "AND DGP",
  "Additive DGP",
  "OR DGP"
)
plt <- purrr::map(
  sim_names,
  function(sim_name) {
    plt_name <- stringr::str_replace(sim_name, "with ", "with\n")
    plt <- viz_results_ls[[sim_name]]$`Subgroup Feature Selection Errors Plot`
    plt_selection <- plt +
      ggplot2::scale_color_manual(
        values = CROSSFIT_COLORS
      ) +
      ggplot2::scale_fill_manual(
        values = CROSSFIT_COLORS
      ) +
      ggplot2::labs(
        color = "# of Repeated\nCrossfits",
        fill = "# of Repeated\nCrossfits"
      )

    plt <- viz_results_ls[[sim_name]]$`Subgroup CATE Errors Plot`
    plt_cate <- plt +
      ggplot2::scale_color_manual(
        values = CROSSFIT_COLORS
      ) +
      ggplot2::scale_fill_manual(
        values = CROSSFIT_COLORS
      ) +
      ggplot2::labs(
        y = "CATE RMSE",
        color = "# of Repeated\nCrossfits",
        fill = "# of Repeated\nCrossfits"
      ) +
      ggplot2::theme(
        strip.background = ggplot2::element_blank(),
        strip.placement = "outside",
        axis.title.y = ggplot2::element_text(face = "bold", size = 16),
        axis.text = ggplot2::element_text(size = 12),
        legend.title = ggplot2::element_text(face = "bold", size = 16),
        legend.text = ggplot2::element_text(size = 12)
      )
    plt <- patchwork::wrap_plots(
      plt_selection, plt_cate, nrow = 1, guides = "collect", widths = c(3.5, 1)
    ) &
      ggplot2::labs(x = NULL)

    if (sim_name != sim_names[2]) {
      plt <- plt &
        ggplot2::guides(
          color = "none", fill = "none"
        )
      hjust <- 0.78
    } else {
      hjust <- 0.82
    }
    if (sim_name == sim_names[1]) {
      plt[[1]] <- plt[[1]] +
        ggplot2::labs(
          title = "(A) Selected Subgroup Features"
        )
      plt[[2]] <- plt[[2]] +
        ggplot2::labs(
          title = "(B) Estimated Subgroup CATEs"
        )
    }
    plt[[1]] <- plt[[1]] +
      ggplot2::labs(
        tag = plt_name
      ) +
      ggplot2::theme(
        plot.tag = ggplot2::element_text(
          size = 18, face = "italic", angle = 90, hjust = 0.5, vjust = 1
        ),
        plot.tag.position = c(-.03, 0.48),
        plot.margin = ggplot2::unit(c(0, 0.35, 0, 0.5), units = "in")
      )
    plt <- plt &
      ggplot2::theme(
        panel.background = ggplot2::element_rect(fill = "grey98"),
        panel.grid.major = ggplot2::element_line(color = "grey98"),
        plot.title = ggplot2::element_text(
          size = 18, face = "italic", hjust = 0.5, vjust = -1
        )
      )
    return(plt)
  }
) |>
  patchwork::wrap_plots(
    ncol = 1
  ) |>
  patchwork::wrap_elements() +
  ggplot2::labs(
    tag = expression(
      bold(paste("Proportion of Variance Explained in ", tau, " by Covariates"))
    )
  ) +
  ggplot2::theme(
    plot.tag = ggplot2::element_text(
      size = 16
    ),
    plot.tag.position = c(0.5, 0),
    plot.margin = ggplot2::unit(c(-0.1, 0, .2, -0.1), units = "in")
  )
ggplot2::ggsave(
  plt, 
  filename = file.path(FIG_DIR, "sim_subgroup_errors_nreps_crossfit.pdf"),
  height = 9, 
  width = 14
)
```

### Thresholds Distances

```{r}
sim_names <- c(
  "AND DGP",
  "Additive DGP",
  "OR DGP"
)
plt <- purrr::imap(
  sim_names,
  function(sim_name, idx) {
    tag <- LETTERS[idx]
    plt <- viz_results_ls[[sim_name]]$`Threshold Distances Plot`
    plt <- plt +
      ggplot2::scale_color_manual(
        values = CROSSFIT_COLORS
      ) +
      ggplot2::scale_fill_manual(
        values = CROSSFIT_COLORS
      ) +
      ggplot2::labs(
        title = sprintf("(%s) %s", tag, sim_name),
        y = "Average  | Estimated Threshold - True Threshold |",
        x = expression(
          bold(
            paste("Proportion of Variance Explained in ", tau, " by Covariates")
          )
        ),
        color = "# of Repeated\nCrossfits",
        fill = "# of Repeated\nCrossfits"
      ) +
      ggplot2::theme(
        panel.background = ggplot2::element_rect(fill = "grey98"),
        panel.grid.major = ggplot2::element_line(color = "grey98"),
        plot.title = ggplot2::element_text(
          size = 18, face = "italic", hjust = 0.5
        ),
        axis.title = ggplot2::element_text(face = "bold", size = 16),
        axis.text = ggplot2::element_text(size = 12),
        legend.text = ggplot2::element_text(size = 12),
        legend.title = ggplot2::element_text(size = 16),
        strip.text = ggplot2::element_text(size = 14)
      )
    if (sim_name != sim_names[2]) {
      plt <- plt +
        ggplot2::guides(
          color = "none", fill = "none"
        )
    }
    if (stringr::str_detect(sim_name, "Linear")) {
      plt <- plt + 
        ggplot2::theme(
          axis.title.y = ggplot2::element_blank(),
          plot.margin = ggplot2::unit(c(0, 0, 0.1, 0.2), units = "in")
        )
    } else {
      plt <- plt +
        ggplot2::theme(
          plot.margin = ggplot2::unit(c(0, 0.2, 0.1, 0), units = "in")
        )
    }
    if (!stringr::str_detect(sim_name, "OR")) {
      plt <- plt + 
        ggplot2::theme(
          axis.title.x = ggplot2::element_blank()
        )
    }
    return(plt)
  }
) |>
  patchwork::wrap_plots(
    ncol = 1, byrow = FALSE
  ) +
  patchwork::plot_layout(
    axis_titles = "collect_y"
  )
ggplot2::ggsave(
  plt, 
  filename = file.path(
    FIG_DIR, sprintf("sim_threshold_distances_nreps_crossfit.pdf")
  ),
  height = 9, 
  width = 8 #14
)
```

### Threshold Distributions

```{r}
KEEP_HERITABILITY <- c(0.2, 0.6, 1)

# without covariates
sim_names <- c(
  "AND DGP",
  "Additive DGP",
  "OR DGP"
)
plt <- purrr::imap(
  sim_names,
  function(sim_name, idx) {
    tag <- LETTERS[idx]
    plt <- viz_results_ls[[sim_name]]$`Subgroup Thresholds Distribution Plot`
    plt$data <- plt$data |>
      dplyr::filter(
        tau_heritability %in% KEEP_HERITABILITY
      )
    plt$layers[[1]]$mapping <- ggplot2::aes(
      x = nreps_crossfit, 
      y = raw_threshold,
      fill = nreps_crossfit
    )
    plt$layers[[2]]$mapping <- ggplot2::aes(
      x = nreps_crossfit, 
      y = raw_threshold,
      color = nreps_crossfit
    )
    plt <- plt +
      ggplot2::scale_color_manual(
        values = CROSSFIT_COLORS
      ) +
      ggplot2::scale_fill_manual(
        values = CROSSFIT_COLORS
      ) +
      ggplot2::facet_grid(.var ~ tau_heritability, labeller = signal_labeller) +
      ggplot2::labs(
        title = sprintf("(%s) %s", tag, sim_name),
        x = "# of Repeated Crossfits",
        color = "# of Repeated\nCrossfits",
        fill = "# of Repeated\nCrossfits"
      ) +
      ggplot2::theme(
        axis.ticks.x = ggplot2::element_line(),
        axis.title.x = ggplot2::element_text(size = 14, face = "bold"),
        axis.text.x = ggplot2::element_text(size = 10),
        panel.background = ggplot2::element_rect(fill = "grey98"),
        panel.grid.major = ggplot2::element_line(color = "grey98"),
        plot.title = ggplot2::element_text(
          size = 18, face = "italic", hjust = 0.5
        )
      ) +
      ggplot2::guides(color = "none", fill = "none")
    return(plt)
  }
) |>
  patchwork::wrap_plots(
    ncol = 1, guides = "collect"
  ) +
  patchwork::plot_layout(
    axes = "collect"
  )
ggplot2::ggsave(
  plt, 
  filename = file.path(FIG_DIR, "sim_thresholds_nreps_crossfit.pdf"),
  height = 10, 
  width = 10
)
```

## Rulefit Simulations

```{r results = "asis"}
sim_results_dir <- here::here(
  file.path("results", "Rulefit Simulations")
)
SIM_NAMES <- c(
  "AND DGP" = "AND DGP",
  "Additive DGP" = "Additive DGP",
  "OR DGP" = "OR DGP",
  "AND with Linear Covariates DGP" = "AND DGP with Linear Covariates",
  "Additive with Linear Covariates DGP" = "Additive DGP with Linear Covariates",
  "OR with Linear Covariates DGP" = "OR DGP with Linear Covariates"
)

# load in results
eval_results_ls <- list()
viz_results_ls <- list()
for (sim_name in names(SIM_NAMES)) {
  # fit_results <- readRDS(
  #   file.path(
  #     sim_results_dir, sim_name, "Varying tau_heritability", "fit_results.rds"
  #   )
  # )
  eval_results <- readRDS(
    file.path(
      sim_results_dir, sim_name, "Varying tau_heritability", "eval_results.rds"
    )
  )
  viz_results <- visualize_experiment(
    rulefit_experiment, NULL, eval_results
  )
  eval_results_ls[[SIM_NAMES[[sim_name]]]] <- eval_results
  viz_results_ls[[SIM_NAMES[[sim_name]]]] <- viz_results
}
```

### Selected Subgroup Features + Estimated Subgroup CATEs

```{r}
KEEP_METHODS <- c(
  "Distilled Causal Forest",
  "Distilled Causal Forest (rulefit v1)",
  "Distilled Causal Forest (rulefit v2)",
  "Distilled Causal Forest (rulefit v3)",
  "Distilled Causal Forest (rulefit v4)"
)

## without covariates
sim_names <- c(
  "AND DGP",
  "Additive DGP",
  "OR DGP"
)
plt <- purrr::map(
  sim_names,
  function(sim_name) {
    plt_name <- stringr::str_replace(sim_name, "with ", "with\n")
    plt <- viz_results_ls[[sim_name]]$`Subgroup Feature Selection Errors Plot`
    plt$data <- plt$data |>
      dplyr::filter(
        .method_name %in% KEEP_METHODS
      ) |> 
      rename_rulefit_methods()
    plt_selection <- plt +
      ggplot2::scale_color_manual(
        values = RULEFIT_COLORS
      ) +
      ggplot2::scale_fill_manual(
        values = RULEFIT_COLORS
      )

    plt <- viz_results_ls[[sim_name]]$`Subgroup CATE Errors Plot`
    plt$data <- plt$data |>
      dplyr::filter(
        .method_name %in% KEEP_METHODS
      ) |> 
      rename_rulefit_methods()
    plt_cate <- plt +
      ggplot2::scale_color_manual(
        values = RULEFIT_COLORS
      ) +
      ggplot2::scale_fill_manual(
        values = RULEFIT_COLORS
      ) +
      ggplot2::labs(
        y = "CATE RMSE"
      ) +
      ggplot2::theme(
        strip.background = ggplot2::element_blank(),
        strip.placement = "outside",
        axis.title.y = ggplot2::element_text(face = "bold", size = 16),
        axis.text = ggplot2::element_text(size = 12),
        legend.title = ggplot2::element_text(face = "bold", size = 16),
        legend.text = ggplot2::element_text(size = 12)
      )
    plt <- patchwork::wrap_plots(
      plt_selection, plt_cate, nrow = 1, guides = "collect", widths = c(3.5, 1)
    ) &
      ggplot2::labs(x = NULL)

    if (sim_name != sim_names[2]) {
      plt <- plt &
        ggplot2::guides(
          color = "none", fill = "none"
        )
      hjust <- 0.78
    } else {
      hjust <- 0.82
    }
    if (sim_name == sim_names[1]) {
      plt[[1]] <- plt[[1]] +
        ggplot2::labs(
          title = "(A) Selected Subgroup Features"
        )
      plt[[2]] <- plt[[2]] +
        ggplot2::labs(
          title = "(B) Estimated Subgroup CATEs"
        )
    }
    plt[[1]] <- plt[[1]] +
      ggplot2::labs(
        tag = plt_name
      ) +
      ggplot2::theme(
        plot.tag = ggplot2::element_text(
          size = 18, face = "italic", angle = 90, hjust = 0.5, vjust = 1
        ),
        plot.tag.position = c(-.03, 0.48),
        plot.margin = ggplot2::unit(c(0, 0.35, 0, 0.5), units = "in")
      )
    plt <- plt &
      ggplot2::theme(
        panel.background = ggplot2::element_rect(fill = "grey98"),
        panel.grid.major = ggplot2::element_line(color = "grey98"),
        plot.title = ggplot2::element_text(
          size = 18, face = "italic", hjust = 0.5, vjust = -1
        )
      )
    return(plt)
  }
) |>
  patchwork::wrap_plots(
    ncol = 1
  ) |>
  patchwork::wrap_elements() +
  ggplot2::labs(
    tag = expression(
      bold(paste("Proportion of Variance Explained in ", tau, " by Covariates"))
    )
  ) +
  ggplot2::theme(
    plot.tag = ggplot2::element_text(
      size = 16
    ),
    plot.tag.position = c(0.5, 0),
    plot.margin = ggplot2::unit(c(-0.1, 0, .2, -0.1), units = "in")
  )
ggplot2::ggsave(
  plt, 
  filename = file.path(FIG_DIR, "sim_subgroup_errors_rulefit.pdf"),
  height = 9, 
  width = 14
)

## with covariates
sim_names <- c(
  "AND DGP with Linear Covariates",
  "Additive DGP with Linear Covariates",
  "OR DGP with Linear Covariates"
)
plt <- purrr::map(
  sim_names,
  function(sim_name) {
    plt_name <- stringr::str_replace(sim_name, "with ", "with\n")
    plt <- viz_results_ls[[sim_name]]$`Subgroup Feature Selection Errors Plot`
    plt$data <- plt$data |>
      dplyr::filter(
        .method_name %in% KEEP_METHODS
      ) |> 
      rename_rulefit_methods()
    plt_selection <- plt +
      ggplot2::scale_color_manual(
        values = RULEFIT_COLORS
      ) +
      ggplot2::scale_fill_manual(
        values = RULEFIT_COLORS
      )

    plt <- viz_results_ls[[sim_name]]$`Subgroup CATE Errors Plot`
    plt$data <- plt$data |>
      dplyr::filter(
        .method_name %in% KEEP_METHODS
      ) |> 
      rename_rulefit_methods()
    plt_cate <- plt +
      ggplot2::scale_color_manual(
        values = RULEFIT_COLORS
      ) +
      ggplot2::scale_fill_manual(
        values = RULEFIT_COLORS
      ) +
      ggplot2::labs(
        y = "CATE RMSE"
      ) +
      ggplot2::theme(
        strip.background = ggplot2::element_blank(),
        strip.placement = "outside",
        axis.title.y = ggplot2::element_text(face = "bold", size = 16),
        axis.text = ggplot2::element_text(size = 12),
        legend.title = ggplot2::element_text(face = "bold", size = 16),
        legend.text = ggplot2::element_text(size = 12)
      )
    plt <- patchwork::wrap_plots(
      plt_selection, plt_cate, nrow = 1, guides = "collect", widths = c(3.5, 1)
    ) &
      ggplot2::labs(x = NULL)

    if (sim_name != sim_names[2]) {
      plt <- plt &
        ggplot2::guides(
          color = "none", fill = "none"
        )
      hjust <- 0.78
    } else {
      hjust <- 0.82
    }
    if (sim_name == sim_names[1]) {
      plt[[1]] <- plt[[1]] +
        ggplot2::labs(
          title = "(A) Selected Subgroup Features"
        )
      plt[[2]] <- plt[[2]] +
        ggplot2::labs(
          title = "(B) Estimated Subgroup CATEs"
        )
    }
    plt[[1]] <- plt[[1]] +
      ggplot2::labs(
        tag = plt_name
      ) +
      ggplot2::theme(
        plot.tag = ggplot2::element_text(
          size = 18, face = "italic", angle = 90, hjust = 0.5, vjust = 1
        ),
        plot.tag.position = c(-.06, 0.48),
        plot.margin = ggplot2::unit(c(0, 0.35, 0, 0.75), units = "in")
      )
    plt <- plt &
      ggplot2::theme(
        panel.background = ggplot2::element_rect(fill = "grey98"),
        panel.grid.major = ggplot2::element_line(color = "grey98"),
        plot.title = ggplot2::element_text(
          size = 18, face = "italic", hjust = 0.5, vjust = -1
        )
      )
    return(plt)
  }
) |>
  patchwork::wrap_plots(
    ncol = 1
  ) |>
  patchwork::wrap_elements() +
  ggplot2::labs(
    tag = expression(
      bold(paste("Proportion of Variance Explained in ", tau, " by Covariates"))
    )
  ) +
  ggplot2::theme(
    plot.tag = ggplot2::element_text(
      size = 16
    ),
    plot.tag.position = c(0.45, 0),
    plot.margin = ggplot2::unit(c(-0.1, 0, .2, -0.1), units = "in")
  )
ggplot2::ggsave(
  plt, 
  filename = file.path(FIG_DIR, "sim_subgroup_errors_cov_rulefit.pdf"),
  height = 9, 
  width = 14
)
```

### Thresholds Distances

```{r}
KEEP_METHODS <- c(
  "Distilled Causal Forest",
  "Distilled Causal Forest (rulefit v1)",
  "Distilled Causal Forest (rulefit v2)",
  "Distilled Causal Forest (rulefit v3)",
  "Distilled Causal Forest (rulefit v4)"
)

sim_names <- c(
  "AND DGP",
  "Additive DGP",
  "OR DGP",
  "AND DGP with Linear Covariates",
  "Additive DGP with Linear Covariates",
  "OR DGP with Linear Covariates"
)
plt <- purrr::imap(
  sim_names,
  function(sim_name, idx) {
    tag <- LETTERS[idx]
    plt <- viz_results_ls[[sim_name]]$`Threshold Distances Plot`
    plt$data <- plt$data |>
      dplyr::filter(
        .method_name %in% KEEP_METHODS
      ) |> 
      rename_rulefit_methods()
    plt <- plt +
      ggplot2::scale_color_manual(
        values = RULEFIT_COLORS
      ) +
      ggplot2::scale_fill_manual(
        values = RULEFIT_COLORS
      ) +
      ggplot2::labs(
        title = sprintf("(%s) %s", tag, sim_name),
        y = "Average  | Estimated Threshold - True Threshold |",
        x = expression(
          bold(
            paste("Proportion of Variance Explained in ", tau, " by Covariates")
          )
        )
      ) +
      ggplot2::theme(
        panel.background = ggplot2::element_rect(fill = "grey98"),
        panel.grid.major = ggplot2::element_line(color = "grey98"),
        plot.title = ggplot2::element_text(
          size = 18, face = "italic", hjust = 0.5
        ),
        axis.title = ggplot2::element_text(face = "bold", size = 16),
        axis.text = ggplot2::element_text(size = 12),
        legend.text = ggplot2::element_text(size = 12),
        legend.title = ggplot2::element_text(size = 16),
        strip.text = ggplot2::element_text(size = 14)
      )
    if (sim_name != sim_names[5]) {
      plt <- plt +
        ggplot2::guides(
          color = "none", fill = "none"
        )
    }
    if (stringr::str_detect(sim_name, "Linear")) {
      plt <- plt + 
        ggplot2::theme(
          axis.title.y = ggplot2::element_blank(),
          plot.margin = ggplot2::unit(c(0, 0, 0.1, 0.2), units = "in")
        )
    } else {
      plt <- plt +
        ggplot2::theme(
          plot.margin = ggplot2::unit(c(0, 0.2, 0.1, 0), units = "in")
        )
    }
    if (!stringr::str_detect(sim_name, "OR")) {
      plt <- plt + 
        ggplot2::theme(
          axis.title.x = ggplot2::element_blank()
        )
    }
    return(plt)
  }
) |>
  patchwork::wrap_plots(
    ncol = 2, byrow = FALSE
  ) +
  patchwork::plot_layout(
    axis_titles = "collect_y"
  )
ggplot2::ggsave(
  plt, 
  filename = file.path(
    FIG_DIR, sprintf("sim_threshold_distances_rulefit.pdf")
  ),
  height = 9, 
  width = 14
)
```

### Threshold Distributions

```{r}
KEEP_METHODS <- c(
  "Distilled Causal Forest",
  "Distilled Causal Forest (rulefit v1)",
  "Distilled Causal Forest (rulefit v2)",
  "Distilled Causal Forest (rulefit v3)",
  "Distilled Causal Forest (rulefit v4)"
)
KEEP_HERITABILITY <- c(0.2, 0.6, 1)

# without covariates
sim_names <- c(
  "AND DGP",
  "Additive DGP",
  "OR DGP"
)
plt <- purrr::imap(
  sim_names,
  function(sim_name, idx) {
    tag <- LETTERS[idx]
    plt <- viz_results_ls[[sim_name]]$`Subgroup Thresholds Distribution Plot`
    plt$data <- plt$data |>
      dplyr::filter(
        .method_name %in% KEEP_METHODS,
        tau_heritability %in% KEEP_HERITABILITY
      ) |> 
      rename_rulefit_methods()
    plt <- plt +
      ggplot2::scale_color_manual(
        values = RULEFIT_COLORS
      ) +
      ggplot2::scale_fill_manual(
        values = RULEFIT_COLORS
      ) +
      ggplot2::facet_grid(.var ~ tau_heritability, labeller = signal_labeller) +
      ggplot2::labs(
        title = sprintf("(%s) %s", tag, sim_name)
      ) +
      ggplot2::theme(
        panel.background = ggplot2::element_rect(fill = "grey98"),
        panel.grid.major = ggplot2::element_line(color = "grey98"),
        plot.title = ggplot2::element_text(
          size = 18, face = "italic", hjust = 0.5
        )
      )
    return(plt)
  }
) |>
  patchwork::wrap_plots(
    ncol = 1, guides = "collect"
  ) +
  patchwork::plot_layout(
    axes = "collect"
  )
ggplot2::ggsave(
  plt, 
  filename = file.path(FIG_DIR, "sim_thresholds_rulefit.pdf"),
  height = 10, 
  width = 11
)

# with covariates
sim_names <- c(
  "AND DGP with Linear Covariates",
  "Additive DGP with Linear Covariates",
  "OR DGP with Linear Covariates"
)
plt <- purrr::imap(
  sim_names,
  function(sim_name, idx) {
    tag <- LETTERS[idx]
    plt <- viz_results_ls[[sim_name]]$`Subgroup Thresholds Distribution Plot`
    plt$data <- plt$data |>
      dplyr::filter(
        .method_name %in% KEEP_METHODS,
        tau_heritability %in% KEEP_HERITABILITY
      ) |> 
      rename_rulefit_methods()
    plt <- plt +
      ggplot2::scale_color_manual(
        values = RULEFIT_COLORS
      ) +
      ggplot2::scale_fill_manual(
        values = RULEFIT_COLORS
      ) +
      ggplot2::facet_grid(.var ~ tau_heritability, labeller = signal_labeller) +
      ggplot2::labs(
        title = sprintf("(%s) %s", tag, sim_name)
      ) +
      ggplot2::theme(
        panel.background = ggplot2::element_rect(fill = "grey98"),
        panel.grid.major = ggplot2::element_line(color = "grey98"),
        plot.title = ggplot2::element_text(
          size = 18, face = "italic", hjust = 0.5
        )
      )
    return(plt)
  }
) |>
  patchwork::wrap_plots(
    ncol = 1, guides = "collect"
  ) +
  patchwork::plot_layout(
    axes = "collect"
  )
ggplot2::ggsave(
  plt, 
  filename = file.path(FIG_DIR, "sim_thresholds_cov_rulefit.pdf"),
  height = 10, 
  width = 11
)
```

### Number of Feature Splits Heatmap

```{r}
KEEP_METHODS <- c(
  "Distilled Causal Forest",
  "Distilled Causal Forest (rulefit v1)",
  "Distilled Causal Forest (rulefit v2)",
  "Distilled Causal Forest (rulefit v3)",
  "Distilled Causal Forest (rulefit v4)"
)
KEEP_HERITABILITY <- c(0.2, 0.6, 1)

# without covariates
sim_names <- c(
  "AND DGP",
  "Additive DGP",
  "OR DGP"
)
plt <- purrr::imap(
  sim_names,
  function(sim_name, idx) {
    tag <- LETTERS[idx]
    plt <- viz_results_ls[[sim_name]]$`Number of Splits Plot`
    plt$data <- plt$data |>
      dplyr::filter(
        .method_name %in% KEEP_METHODS,
        tau_heritability %in% KEEP_HERITABILITY
      ) |> 
      rename_rulefit_methods()
    plt_nsplits <- plt +
      ggplot2::facet_grid(tau_heritability ~ ., labeller = signal_labeller) +
      viridis::scale_fill_viridis(
        begin = 0.1, end = 1, option = "magma"#, limits = c(0., 4.5)
      )

    plt <- viz_results_ls[[sim_name]]$`Number of Trees Plot`
    plt$data <- plt$data |>
      dplyr::filter(
        .method_name %in% KEEP_METHODS,
        tau_heritability %in% KEEP_HERITABILITY
      ) |> 
      rename_rulefit_methods()
    plt_ntrees <- plt +
      ggplot2::facet_grid(tau_heritability ~ ., labeller = signal_labeller) +
      viridis::scale_fill_viridis(
        begin = 0.1, end = 1, option = "magma", limits = c(0, 100)
      )

    plt <- patchwork::wrap_plots(plt_ntrees, plt_nsplits, nrow = 1) +
      patchwork::plot_layout(
        axis_titles = "collect_y"
      ) &
      ggplot2::labs(x = "Variable")
    plt <- patchwork::wrap_elements(plt) +
      ggplot2::labs(
        tag = sprintf("(%s) %s", tag, sim_name)
      ) +
      ggplot2::theme(
        plot.tag = ggplot2::element_text(
          size = 18, face = "italic", hjust = 0.5
        ),
        plot.tag.position = c(0.47, 1.02),
        plot.margin = ggplot2::unit(c(0.3, 0, 0, 0.5), units = "in")
      )
    return(plt)
  }
) |>
  patchwork::wrap_plots(
    ncol = 1
  )
ggplot2::ggsave(
  plt, 
  filename = file.path(FIG_DIR, "sim_freq_rulefit.pdf"),
  height = 17.5, 
  width = 15
)

# with covariates
sim_names <- c(
  "AND DGP with Linear Covariates",
  "Additive DGP with Linear Covariates",
  "OR DGP with Linear Covariates"
)
plt <- purrr::imap(
  sim_names,
  function(sim_name, idx) {
    tag <- LETTERS[idx]
    plt <- viz_results_ls[[sim_name]]$`Number of Splits Plot`
    plt$data <- plt$data |>
      dplyr::filter(
        .method_name %in% KEEP_METHODS,
        tau_heritability %in% KEEP_HERITABILITY
      ) |> 
      rename_rulefit_methods()
    plt_nsplits <- plt +
      ggplot2::facet_grid(tau_heritability ~ ., labeller = signal_labeller) +
      viridis::scale_fill_viridis(
        begin = 0.1, end = 1, option = "magma"#, limits = c(0., 4.5)
      )

    plt <- viz_results_ls[[sim_name]]$`Number of Trees Plot`
    plt$data <- plt$data |>
      dplyr::filter(
        .method_name %in% KEEP_METHODS,
        tau_heritability %in% KEEP_HERITABILITY
      ) |> 
      rename_rulefit_methods()
    plt_ntrees <- plt +
      ggplot2::facet_grid(tau_heritability ~ ., labeller = signal_labeller) +
      viridis::scale_fill_viridis(
        begin = 0.1, end = 1, option = "magma", limits = c(0, 100)
      )

    plt <- patchwork::wrap_plots(plt_ntrees, plt_nsplits, nrow = 1) +
      patchwork::plot_layout(
        axis_titles = "collect_y"
      ) &
      ggplot2::labs(x = "Variable")
    plt <- patchwork::wrap_elements(plt) +
      ggplot2::labs(
        tag = sprintf("(%s) %s", tag, sim_name)
      ) +
      ggplot2::theme(
        plot.tag = ggplot2::element_text(
          size = 18, face = "italic", hjust = 0.5
        ),
        plot.tag.position = c(0.47, 1.02),
        plot.margin = ggplot2::unit(c(0.3, 0, 0, 0.5), units = "in")
      )
    return(plt)
  }
) |>
  patchwork::wrap_plots(
    ncol = 1
  )
ggplot2::ggsave(
  plt, 
  filename = file.path(FIG_DIR, "sim_freq_cov_rulefit.pdf"),
  height = 17.5, 
  width = 15
)
```