### 2025-02-13 causalDT-manuscript/README.Rmd

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Causal Distillation Trees

[Causal Distillation Trees](https://arxiv.org/abs/2502.07275) (CDT) is a novel machine learning method for estimating interpretable subgroups in causal inference. CDT allows researchers to fit *any* machine learning model of their choice to estimate the individual-level treatment effect, and then leverages a simple, second-stage tree-based model to "distill" the estimated treatment effect into meaningful subgroups. As a result, CDT inherits the improvements in predictive performance from black-box machine learning models while preserving the interpretability of a simple decision tree.

![](../causalDT/man/figures/cdt_diagram.png)

Briefly, CDT is a two-stage learner that first fits a teacher model (e.g., a black-box metalearner) to estimate individual-level treatment effects and secondly fits a student model (e.g., a decision tree) to predict the estimated individual-level treatment effects, in effect distilling the estimated individual-level treatment effects and producing interpretable subgroups. This two-stage learner is learned using the training data. Finally, using the estimated subgroups, the subgroup average treatment effects are honestly estimated with a held-out estimation set.

For more details, check out [Huang, M., Tang, T. M., Kenney, A. M. "Distilling heterogeneous treatment effects: Stable subgroup estimation in causal inference." (2025).](https://arxiv.org/abs/2502.07275)

## Project Structure

This directory contains all of the code necessary to reproduce the analysis and figures in [Huang et al. (2025)](https://arxiv.org/abs/2502.07275).

Moreover, to facilitate reproducibility of this work, we leveraged [`renv`](https://rstudio.github.io/renv/articles/renv.html) to create a reproducible environment and [`simChef`](https://yu-group.github.io/simChef/) to create a reproducible and transparent simulation workflow. Thus, to install the necessary dependencies and reproduce the simulation, the following steps are recommended:

1. Clone this repository.
2. Open the `causalDT-manuscript.Rproj` in RStudio.
3. Run the following code in the R console:

```r
# install.packages("renv")
renv::restore()
```

4. Run each of the simulation driver scripts in the `meals/` directory.
5. Render all of the results into a single summary `simChef` documentation by running the `meals/00_render_docs.R` script.

These steps will generate all of the figures and results from the mansucript. The rendered `simChef` documentation can be found in the `results/` directory (see [here](results/Results.html)).

## Citation

```r
@article{huang2025distilling,
  title={Distilling heterogeneous treatment effects: Stable subgroup estimation in causal inference}, 
  author={Melody Huang and Tiffany M. Tang and Ana M. Kenney},
  year={2025},
  eprint={2502.07275},
  archivePrefix={arXiv},
  primaryClass={stat.ME},
  url={https://arxiv.org/abs/2502.07275}, 
}
```