### 2025-02-13 causalDT-manuscript/results/Results.Rmd
add results

---
author: 'Melody Huang, Tiffany M. Tang, Ana M. Kenney'
date: '`r format(Sys.time(), "%B %d, %Y")`'
title: "Causal Distillation Trees: Supplemental Simulation Results"
output: vthemes::vmodern
params:
  sim_path:
    input: text
    label: 'Path to Simulation Experiment Folder:'
    value: /Users/ttang4/Documents/Research/Causal Subgroups/causalDT/causalDT-manuscript/results
---

<script src="/Users/ttang4/Library/Caches/org.R-project.R/R/renv/cache/v5/macos/R-4.4/aarch64-apple-darwin20/simChef/0.1.0.9000/35207a8200887c29dd6eb0e5f5baa7f9/simChef/rmd/js/simchefNavClass.js"></script>

<link rel="stylesheet" href="/Users/ttang4/Library/Caches/org.R-project.R/R/renv/cache/v5/macos/R-4.4/aarch64-apple-darwin20/simChef/0.1.0.9000/35207a8200887c29dd6eb0e5f5baa7f9/simChef/rmd/css/simchef.css">

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  cache = FALSE,
  fig.align = "center",
  fig.pos = "H",
  fig.height = 12,
  fig.width = 10
)

options(
  width = 10000,
  knitr.kable.NA = 'NA'
)

library(simChef)

local({
  hook_output <- knitr::knit_hooks$get("output")
  knitr::knit_hooks$set(
    # scrollable text output
    output = function(x, options) {
      if (!is.null(options$max.height)) {
        options$attr.output <- c(
          options$attr.output,
          sprintf('style="max-height: %s;"', options$max.height)
        )
      }
      hook_output(x, options)
    },
    # add panel around figures
    add.panel = function(before, options, envir) {
      if (before) {
        HTML("<div class='panel panel-default padded-panel'>")
      } else {
        HTML("</div>")
      }
    }
  )
})
```

```{css}
h5 {
  font-size: 16pt;
}
```

```{r}
#' Get all experiments under a given directory name.
#'
#' @param dir_name Name of directory.
#' @return List of named experiments.
get_descendants <- function(dir_name) {
  experiments <- list()
  for (d in list.dirs(dir_name)) {
    if (file.exists(file.path(d, "experiment.rds"))) {
      if (identical(d, params$sim_path)) {
        exp_name <- "Base"
      } else {
        exp_name <- stringr::str_replace_all(
          stringr::str_remove(d, paste0(params$sim_path, "/")),
          "/", " - "
        )
      }
      experiments[[exp_name]] <- readRDS(file.path(d, "experiment.rds"))
    }
  }
  return(experiments)
}

#' Reads in file if it exists and returns NULL if the file does not exist
#' 
#' @description If filetype is ".rds", the file is read in using readRDS(). 
#'   Otherwise, the file is read in using data.table::fread().
#'
#' @param filename name of file (with file extension) to try reading in
#' @param filetype file extension
#' @param experiment experiment object
#' @param experiment_save_dir directory where experiment results are saved
#' @param field_name one of "evaluator" or "visualizer"
#' @return output of experiment results if the file exists and NULL otherwise
get_results <- function(filename, filetype = ".rds",
                        experiment, experiment_save_dir = NULL,
                        field_name = c("evaluator", "visualizer")) {
  field_name <- match.arg(field_name)
  if (field_name == "evaluator") {
    save_in_bulk <- experiment$get_save_in_bulk()[["eval"]]
  } else if (field_name == "visualizer") {
    save_in_bulk <- experiment$get_save_in_bulk()[["viz"]]
  }
  results <- NULL
  if (save_in_bulk) {
    if (file.exists(filename)) {
      if (filetype == ".rds") {
        results <- readRDS(filename)
      } else {
        results <- data.table::fread(results)
      }
    }
  } else {
    if (is.null(experiment_save_dir)) {
      stop("experiment_save_dir must be provided if save_in_bulk is FALSE")
    }
    if (field_name == "evaluator") {
      obj_names <- names(experiment$get_evaluators())
      obj_dirname <- file.path(experiment_save_dir, "eval_results")
    } else if (field_name == "visualizer") {
      obj_names <- names(experiment$get_visualizers())
      obj_dirname <- file.path(experiment_save_dir, "viz_results")
    }
    if (length(obj_names) == 0) {
      return(NULL)
    }
    names(obj_names) <- obj_names
    results <- purrr::map(
      obj_names, 
      function(obj_name) {
        obj_fname <- file.path(obj_dirname, sprintf("%s%s", obj_name, filetype))
        if (file.exists(obj_fname)) {
          if (filetype == ".rds") {
            results <- readRDS(obj_fname)
          } else {
            results <- data.table::fread(obj_fname)
          }
        } else {
          results <- NULL
        }
        return(results)
      } 
    ) |> 
      purrr::compact()
  }

  return(results)
}

#' Reads in list of image result files
#' 
#' @param filenames vector of filenames (without file extension) to read in
#' @param filetype file extension
#' @return list of image results, wrapped as knitr::include_graphics() objects
get_image_results <- function(filenames, filetype) {
  purrr::map(
    filenames,
    function(f) {
      if (file.exists(sprintf("%s%s", f, filetype))) {
        return(knitr::include_graphics(sprintf("%s%s", f, filetype)))
      } else {
        return(NULL)
      }
    }
  ) |> 
    setNames(basename(filenames)) |> 
    purrr::compact()
}

#' Get results from experiment
#' 
#' @param dir_name name of directory
#' @param show_eval logical; whether or not to show evaluators
#' @param show_viz logical; whether or not to show visualizers
#' @param eval_cache file extension for cached evaluator results to read in.
#'   Typically ".rds" or "none", but can be any file extension where
#'   evaluator results are stored as `eval_results.ext` and can be read in
#'   using `data.table::fread(eval_results.ext)`. If "none", evaluator results
#'   are computed using the experiment via `evaluate_experiment()`.
#' @param viz_cache file extension for cached visualizer results to read in.
#'   Typically ".rds" or "none", but can be any (image) file extension 
#'   (e.g., "png", "jpg") where the visualizer results have been previously 
#'   stored as separate `{visualizer_name}.ext` images (e.g., using
#'   `export_visualizers()`). If "none", visualizer results are computed using
#'   the experiment via `visualize_experiment()`.
get_exp_results <- function(dir_name, 
                            show_eval = TRUE, show_viz = TRUE,
                            eval_cache = ".rds", viz_cache = ".rds") {

  exp_fname <- file.path(dir_name, "experiment.rds")
  eval_fname <- file.path(dir_name, sprintf("eval_results%s", eval_cache))
  viz_fname <- file.path(dir_name, sprintf("viz_results%s", viz_cache))

  if (file.exists(exp_fname)) {
    exp <- readRDS(exp_fname)
  } else {
    results <- list(
      exp = NULL,
      eval_results = NULL,
      viz_results = NULL
    )
    return(results)
  }
  fit_results <- NULL
  eval_results <- NULL
  viz_results <- NULL
  if ((eval_cache != "none") && (viz_cache == ".rds")) {
    if (show_eval) {
      eval_results <- get_results(
        filename = eval_fname, 
        filetype = eval_cache, 
        experiment = exp, 
        experiment_save_dir = dir_name,
        field_name = "evaluator"
      )
    }
    if (show_viz) {
      viz_results <- get_results(
        filename = viz_fname, 
        filetype = viz_cache, 
        experiment = exp, 
        experiment_save_dir = dir_name,
        field_name = "visualizer"
      )
    }
  } else {
    if (show_eval) {
      if (eval_cache == "none") {
        fit_results <- suppressMessages(get_cached_results(exp, "fit"))
        if (is.null(fit_results)) {
          stop("Cannot set eval_cache = 'none' since no cached fit results found. Perhaps try setting eval_cache = '.rds' instead.")
        }
        eval_results <- evaluate_experiment(exp, fit_results)
      } else {
        eval_results <- get_results(
          filename = eval_fname, 
          filetype = eval_cache, 
          experiment = exp, 
          experiment_save_dir = dir_name,
          field_name = "evaluator"
        )
      }
    }
    if (show_viz) {
      if (viz_cache == ".rds") {
        viz_results <- get_results(
          filename = viz_fname, 
          filetype = viz_cache, 
          experiment = exp, 
          experiment_save_dir = dir_name,
          field_name = "visualizer"
        )
      } else if (viz_cache == "none") {
        if (is.null(fit_results)) {
          fit_results <- suppressMessages(get_cached_results(exp, "fit"))
          if (is.null(fit_results)) {
            stop("Cannot set viz_cache = 'none' since no cached fit results found. Perhaps try setting viz_cache = '.rds' instead.")
          }
        }
        if (is.null(eval_results)) {
          eval_results <- suppressMessages(get_cached_results(exp, "eval"))
          if (is.null(eval_results)) {
            eval_results <- evaluate_experiment(exp, fit_results)
          }
        }
        viz_results <- visualize_experiment(exp, fit_results, eval_results)
      } else {
        viz_results <- get_image_results(
          file.path(dir_name, names(exp$get_visualizers())), viz_cache
        )
      }
    }
  }
  results <- list(
    exp = exp,
    eval_results = eval_results,
    viz_results = viz_results
  )
  return(results)
}

#' Print renv lockfile info
#' 
#' @param renv_lockfile renv lockfile object from [renv::lockfile_read()]
#' @param what part of renv lockfile to print; must be one of "R" or "packages"
#' @param print_table logical; whether to print as a nice table (TRUE) or 
#'   simply using the `print` method.
#' @return renv lockfile info
print_renv <- function(renv_lockfile, 
                       what = c("R", "Packages"),
                       print_table = FALSE) {
  what <- match.arg(what)
  if (what == "R") {
    if (print_table) {
      tab <- tibble::tibble(
        `R Version` = renv_lockfile$R$Version, 
        Repositories = paste(
          names(renv_lockfile$R$Repositories), 
          " (", renv_lockfile$R$Repositories, ")", 
          sep = "", collapse = "<br>"
        )
      ) |> 
        vthemes::pretty_DT(
          rownames = FALSE, options = list(dom = "t", ordering = FALSE)
        )
      return(tab)
    } else {
      return(renv_lockfile$R)
    }
  } else {
    if (print_table) {
      tab <- tibble::tibble(
        `Package Name` = purrr::map_chr(
          renv_lockfile$Packages, 
          function(.x) if (!is.null(.x$Package)) .x$Package else "N/A"
        ),
        `Version` = purrr::map_chr(
          renv_lockfile$Packages,
          function(.x) if (!is.null(.x$Version)) .x$Version else "N/A"
        ),
        `Source` = purrr::map_chr(
          renv_lockfile$Packages,
          function(.x) if (!is.null(.x$Source)) .x$Source else "N/A"
        ),
        `Repository` = purrr::map_chr(
          renv_lockfile$Packages,
          function(.x) {
            if (!is.null(.x$Repository)) {
              .x$Repository
            } else if (!is.null(.x$RemoteUrl)) {
              .x$RemoteUrl
            } else if (.x$Source == "GitHub") {
              sprintf("https://github.com/%s/%s", .x$RemoteUsername, .x$RemoteRepo)
            } else {
              "N/A"
            }
          }
        ),
        `Requirements` = purrr::map_chr(
          renv_lockfile$Packages, ~ paste(.x$Requirements, collapse = ", ")
        )
      ) |> 
        vthemes::pretty_DT(rownames = FALSE)
      return(tab)
    } else {
      return(renv_lockfile$Packages)
    }
  }
}

#' Show recipe content.
#' 
#' @param objs List of DGP/Method/Evaluator/Visualizer objects. Should be one of
#'   `dgp_objs`, `method_objs`, `eval_objs`, or `viz_objs` defined below.
#' @param name Name of DGP/Method/Evaluator/Visualizer.
#' @param experiment_name Name of the Experiment.
#' @param what Either "function" or "parameters", indicating whether to return 
#'   the function code or parameters.
show_recipe <- function(objs, name, experiment_name,
                        what = c("function", "parameters")) {
  what <- match.arg(what)
  obj <- objs[[experiment_name]][[name]]
  field_name <- tolower(class(obj)[1])
  func_name <- dplyr::case_when(
    field_name == "evaluator" ~ "eval",
    field_name == "visualizer" ~ "viz",
    TRUE ~ field_name
  )
  if (identical(what, "function")) {
    return(obj[[paste0(func_name, "_fun")]])
  } else if (identical(what, "parameters")) {
    return(obj[[paste0(func_name, "_params")]])
  }
}

#' Show results of Evaluator, Visualizer, or Varying Across Parameters.
#' 
#' @param dir_name Name of directory of Experiment.
#' @param name Name of Evaluator/Visualizer.
#' @param field_name One of "evaluator", "visualizer", or "vary_params", 
#'   indicating whether to return the results of an Evaluator, Visualizer, or
#'   the varying across parameters of the Experiment, respectively.
show_results <- function(results, name,
                         field_name = c("evaluator", "visualizer", "vary_params")) {
  field_name <- match.arg(field_name)
  exp <- results$exp
  eval_results <- results$eval_results
  viz_results <- results$viz_results

  if (field_name == "evaluator") {
    doc_nrows <- exp$get_evaluators()[[name]]$doc_nrows
    if (is.null(doc_nrows)) {
      eval_results_show <- eval_results[[name]]
    } else {
      keep_rows <- 1:min(doc_nrows, nrow(eval_results[[name]]))
      eval_results_show <- eval_results[[name]][keep_rows, ]
    }
    do.call(vthemes::pretty_DT,
            c(list(eval_results_show),
              exp$get_evaluators()[[name]]$doc_options))
  } else if (field_name == "visualizer") {
    viz_results[[name]]
  } else if (field_name == "vary_params") {
    exp$get_vary_across()
  }
}
```

# Simulation Experiment Recipe {.tabset .tabset-vmodern}

```{r}
descendants <- get_descendants(dir_name = params$sim_path)
dgp_objs <- purrr::map(descendants, ~.x$get_dgps())
method_objs <- purrr::map(descendants, ~.x$get_methods())
eval_objs <- purrr::map(descendants, ~.x$get_evaluators())
viz_objs <- purrr::map(descendants, ~.x$get_visualizers())
```

## Objectives

<div class='panel panel-default padded-panel'>

[Huang et al. (2025)]() introduces **causal distillation trees** (CDT) to stably estimate interpretable subgroups. CDT allows researchers to fit any machine learning model of their choice to estimate the individual-level treatment effect, and then leverages a simple, second-stage tree-based model to "distill" the estimated treatment effect into meaningful subgroups.In this supplementary document, we report the full simulation results (originally presented in [Huang et al. (2025)]()) alongside code and necessary details to reproduce the results. This document as well as the simulations were run using the [`simChef`](https://yu-group.github.io/simChef/) R package.

</div>

## Data Generation

<div class='panel panel-default padded-panel'>
For all DGPs, we consider a random treatment assignment process, where $Z_i \stackrel{iid}{\sim} \text{Bernoulli}(1/2)$ for $i = 1, \ldots, n$ samples. Furthermore, we generate $n$ samples and $p$ covariates from a standard multivariate normal distribution (i.e., $X_i \stackrel{iid}{\sim} \text{MVN}(0, I)$, where $X_i \in \mathbb{R}^{p}$ and $i = 1, \ldots, n$). Each DGP is generated via a different treatment effect heterogeneity (i.e., subgroup) process (for $\tau_i$) and outcome process (for $Y_i$), described below.
</div>

<div class='panel panel-default padded-panel'>

### Additive DGP {.tabset .tabset-pills .tabset-circle .tabset-recipe}

#### `r fontawesome::fa('readme', fill = 'white')` {.tabset .tabset-pills}

In the Additive DGP, the treatment effect heterogeneity (i.e., subgroup) DGP is generated via\begin{align*}\tau_i = 2 \mathbf{1} \left \{X_{i}^{(1)} > 0 \right\} - \mathbf{1}\left \{X_{i}^{(2)} < -0.5 \right\} + \epsilon_i, \qquad \text{where } \epsilon_i \stackrel{iid}{\sim} N(0, \sigma_{\tau}^2),\end{align*}and the outcome process is generated via:\begin{align*}Y_i = Z_i \cdot \tau_i + \nu_i, \qquad \text{where } \nu_i \stackrel{iid}{\sim} N(0, 0.1^2).\end{align*}

#### `r fontawesome::fa('code', fill = 'white')` {.tabset .tabset-pills}

##### Main Simulations - Additive DGP - Varying n

**Function**

```{r, max.height='200px'}
show_recipe(dgp_objs, 'Additive DGP', 'Main Simulations - Additive DGP - Varying n', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(dgp_objs, 'Additive DGP', 'Main Simulations - Additive DGP - Varying n', what = 'parameters')
```

##### Main Simulations - Additive DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(dgp_objs, 'Additive DGP', 'Main Simulations - Additive DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(dgp_objs, 'Additive DGP', 'Main Simulations - Additive DGP - Varying tau_heritability', what = 'parameters')
```

##### Rulefit Simulations - Additive DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(dgp_objs, 'Additive DGP', 'Rulefit Simulations - Additive DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(dgp_objs, 'Additive DGP', 'Rulefit Simulations - Additive DGP - Varying tau_heritability', what = 'parameters')
```

##### Varying nrep crossfits - Additive DGP-Distilled Rboost (crossfit) - Varying tau_heritability-nreps_crossfit

**Function**

```{r, max.height='200px'}
show_recipe(dgp_objs, 'Additive DGP', 'Varying nrep crossfits - Additive DGP-Distilled Rboost (crossfit) - Varying tau_heritability-nreps_crossfit', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(dgp_objs, 'Additive DGP', 'Varying nrep crossfits - Additive DGP-Distilled Rboost (crossfit) - Varying tau_heritability-nreps_crossfit', what = 'parameters')
```

</div>

<div class='panel panel-default padded-panel'>

### Additive with Linear Covariates DGP {.tabset .tabset-pills .tabset-circle .tabset-recipe}

#### `r fontawesome::fa('readme', fill = 'white')` {.tabset .tabset-pills}

In the Additive with Linear Covariates DGP, the treatment effect heterogeneity (i.e., subgroup) DGP is generated via\begin{align*}\tau_i = 2 \mathbf{1} \left \{X_{i}^{(1)} > 0 \right\} - \mathbf{1}\left \{X_{i}^{(2)} < -0.5 \right\} + \epsilon_i, \qquad \text{where } \epsilon_i \stackrel{iid}{\sim} N(0, \sigma_{\tau}^2),\end{align*}and the outcome process is generated via:\begin{align*}Y_i = Z_i \cdot \tau_i + X_i^{(3)} + X_i^{(4)} + \nu_i, \qquad \text{where } \nu_i \stackrel{iid}{\sim} N(0, 0.1^2).\end{align*}

#### `r fontawesome::fa('code', fill = 'white')` {.tabset .tabset-pills}

##### Main Simulations - Additive with Linear Covariates DGP - Varying n

**Function**

```{r, max.height='200px'}
show_recipe(dgp_objs, 'Additive with Linear Covariates DGP', 'Main Simulations - Additive with Linear Covariates DGP - Varying n', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(dgp_objs, 'Additive with Linear Covariates DGP', 'Main Simulations - Additive with Linear Covariates DGP - Varying n', what = 'parameters')
```

##### Main Simulations - Additive with Linear Covariates DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(dgp_objs, 'Additive with Linear Covariates DGP', 'Main Simulations - Additive with Linear Covariates DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(dgp_objs, 'Additive with Linear Covariates DGP', 'Main Simulations - Additive with Linear Covariates DGP - Varying tau_heritability', what = 'parameters')
```

##### Rulefit Simulations - Additive with Linear Covariates DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(dgp_objs, 'Additive with Linear Covariates DGP', 'Rulefit Simulations - Additive with Linear Covariates DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(dgp_objs, 'Additive with Linear Covariates DGP', 'Rulefit Simulations - Additive with Linear Covariates DGP - Varying tau_heritability', what = 'parameters')
```

</div>

<div class='panel panel-default padded-panel'>

### AND DGP {.tabset .tabset-pills .tabset-circle .tabset-recipe}

#### `r fontawesome::fa('readme', fill = 'white')` {.tabset .tabset-pills}

In the AND DGP, the treatment effect heterogeneity (i.e., subgroup) DGP is generated via\begin{align*}\tau_i = 2 \mathbf{1}\left\{X_{i}^{(1)} > 0 \right\} \cdot \mathbf{1}\left\{X_{i}^{(2)} > 0.5\right\} + \epsilon_i, \qquad \text{where } \epsilon_i \stackrel{iid}{\sim} N(0, \sigma_{\tau}^2),\end{align*}and the outcome process is generated via:\begin{align*}Y_i = Z_i \cdot \tau_i + \nu_i, \qquad \text{where } \nu_i \stackrel{iid}{\sim} N(0, 0.1^2).\end{align*}

#### `r fontawesome::fa('code', fill = 'white')` {.tabset .tabset-pills}

##### Main Simulations - AND DGP - Varying n

**Function**

```{r, max.height='200px'}
show_recipe(dgp_objs, 'AND DGP', 'Main Simulations - AND DGP - Varying n', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(dgp_objs, 'AND DGP', 'Main Simulations - AND DGP - Varying n', what = 'parameters')
```

##### Main Simulations - AND DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(dgp_objs, 'AND DGP', 'Main Simulations - AND DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(dgp_objs, 'AND DGP', 'Main Simulations - AND DGP - Varying tau_heritability', what = 'parameters')
```

##### Rulefit Simulations - AND DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(dgp_objs, 'AND DGP', 'Rulefit Simulations - AND DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(dgp_objs, 'AND DGP', 'Rulefit Simulations - AND DGP - Varying tau_heritability', what = 'parameters')
```

##### Varying nrep crossfits - AND DGP-Distilled Rboost (crossfit) - Varying tau_heritability-nreps_crossfit

**Function**

```{r, max.height='200px'}
show_recipe(dgp_objs, 'AND DGP', 'Varying nrep crossfits - AND DGP-Distilled Rboost (crossfit) - Varying tau_heritability-nreps_crossfit', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(dgp_objs, 'AND DGP', 'Varying nrep crossfits - AND DGP-Distilled Rboost (crossfit) - Varying tau_heritability-nreps_crossfit', what = 'parameters')
```

</div>

<div class='panel panel-default padded-panel'>

### AND with Linear Covariates DGP {.tabset .tabset-pills .tabset-circle .tabset-recipe}

#### `r fontawesome::fa('readme', fill = 'white')` {.tabset .tabset-pills}

In the AND with Linear Covariates DGP, the treatment effect heterogeneity (i.e., subgroup) DGP is generated via\begin{align*}\tau_i = 2 \mathbf{1}\left\{X_{i}^{(1)} > 0 \right\} \cdot \mathbf{1}\left\{X_{i}^{(2)} > 0.5\right\} + \epsilon_i, \qquad \text{where } \epsilon_i \stackrel{iid}{\sim} N(0, \sigma_{\tau}^2),\end{align*}and the outcome process is generated via:\begin{align*}Y_i = Z_i \cdot \tau_i + X_i^{(3)} + X_i^{(4)} + \nu_i, \qquad \text{where } \nu_i \stackrel{iid}{\sim} N(0, 0.1^2).\end{align*}

#### `r fontawesome::fa('code', fill = 'white')` {.tabset .tabset-pills}

##### Main Simulations - AND with Linear Covariates DGP - Varying n

**Function**

```{r, max.height='200px'}
show_recipe(dgp_objs, 'AND with Linear Covariates DGP', 'Main Simulations - AND with Linear Covariates DGP - Varying n', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(dgp_objs, 'AND with Linear Covariates DGP', 'Main Simulations - AND with Linear Covariates DGP - Varying n', what = 'parameters')
```

##### Main Simulations - AND with Linear Covariates DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(dgp_objs, 'AND with Linear Covariates DGP', 'Main Simulations - AND with Linear Covariates DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(dgp_objs, 'AND with Linear Covariates DGP', 'Main Simulations - AND with Linear Covariates DGP - Varying tau_heritability', what = 'parameters')
```

##### Rulefit Simulations - AND with Linear Covariates DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(dgp_objs, 'AND with Linear Covariates DGP', 'Rulefit Simulations - AND with Linear Covariates DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(dgp_objs, 'AND with Linear Covariates DGP', 'Rulefit Simulations - AND with Linear Covariates DGP - Varying tau_heritability', what = 'parameters')
```

</div>

<div class='panel panel-default padded-panel'>

### OR DGP {.tabset .tabset-pills .tabset-circle .tabset-recipe}

#### `r fontawesome::fa('readme', fill = 'white')` {.tabset .tabset-pills}

In the OR DGP, the treatment effect heterogeneity (i.e., subgroup) DGP is generated via\begin{align*}\tau_i = 2 \mathbf{1}\left\{X_{i}^{(1)} > 0 \right\} - \mathbf{1} \left\{(X_{i}^{(2)} > 0.5) \text{ or } (X_i^{(2)} < -0.5)\right\} + \epsilon_i, \qquad \text{where } \epsilon_i \stackrel{iid}{\sim} N(0, \sigma_{\tau}^2),\end{align*}and the outcome process is generated via:\begin{align*}Y_i = Z_i \cdot \tau_i + \nu_i, \qquad \text{where } \nu_i \stackrel{iid}{\sim} N(0, 0.1^2).\end{align*}

#### `r fontawesome::fa('code', fill = 'white')` {.tabset .tabset-pills}

##### Main Simulations - OR DGP - Varying n

**Function**

```{r, max.height='200px'}
show_recipe(dgp_objs, 'OR DGP', 'Main Simulations - OR DGP - Varying n', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(dgp_objs, 'OR DGP', 'Main Simulations - OR DGP - Varying n', what = 'parameters')
```

##### Main Simulations - OR DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(dgp_objs, 'OR DGP', 'Main Simulations - OR DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(dgp_objs, 'OR DGP', 'Main Simulations - OR DGP - Varying tau_heritability', what = 'parameters')
```

##### Rulefit Simulations - OR DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(dgp_objs, 'OR DGP', 'Rulefit Simulations - OR DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(dgp_objs, 'OR DGP', 'Rulefit Simulations - OR DGP - Varying tau_heritability', what = 'parameters')
```

##### Varying nrep crossfits - OR DGP-Distilled Rboost (crossfit) - Varying tau_heritability-nreps_crossfit

**Function**

```{r, max.height='200px'}
show_recipe(dgp_objs, 'OR DGP', 'Varying nrep crossfits - OR DGP-Distilled Rboost (crossfit) - Varying tau_heritability-nreps_crossfit', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(dgp_objs, 'OR DGP', 'Varying nrep crossfits - OR DGP-Distilled Rboost (crossfit) - Varying tau_heritability-nreps_crossfit', what = 'parameters')
```

</div>

<div class='panel panel-default padded-panel'>

### OR with Linear Covariates DGP {.tabset .tabset-pills .tabset-circle .tabset-recipe}

#### `r fontawesome::fa('readme', fill = 'white')` {.tabset .tabset-pills}

In the OR with Linear Covariates DGP, the treatment effect heterogeneity (i.e., subgroup) DGP is generated via\begin{align*}\tau_i = 2 \mathbf{1}\left\{X_{i}^{(1)} > 0 \right\} - \mathbf{1} \left\{(X_{i}^{(2)} > 0.5) \text{ or } (X_i^{(2)} < -0.5)\right\} + \epsilon_i, \qquad \text{where } \epsilon_i \stackrel{iid}{\sim} N(0, \sigma_{\tau}^2),\end{align*}and the outcome process is generated via:\begin{align*}Y_i = Z_i \cdot \tau_i + X_i^{(3)} + X_i^{(4)} + \nu_i, \qquad \text{where } \nu_i \stackrel{iid}{\sim} N(0, 0.1^2).\end{align*}

#### `r fontawesome::fa('code', fill = 'white')` {.tabset .tabset-pills}

##### Main Simulations - OR with Linear Covariates DGP - Varying n

**Function**

```{r, max.height='200px'}
show_recipe(dgp_objs, 'OR with Linear Covariates DGP', 'Main Simulations - OR with Linear Covariates DGP - Varying n', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(dgp_objs, 'OR with Linear Covariates DGP', 'Main Simulations - OR with Linear Covariates DGP - Varying n', what = 'parameters')
```

##### Main Simulations - OR with Linear Covariates DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(dgp_objs, 'OR with Linear Covariates DGP', 'Main Simulations - OR with Linear Covariates DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(dgp_objs, 'OR with Linear Covariates DGP', 'Main Simulations - OR with Linear Covariates DGP - Varying tau_heritability', what = 'parameters')
```

##### Rulefit Simulations - OR with Linear Covariates DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(dgp_objs, 'OR with Linear Covariates DGP', 'Rulefit Simulations - OR with Linear Covariates DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(dgp_objs, 'OR with Linear Covariates DGP', 'Rulefit Simulations - OR with Linear Covariates DGP - Varying tau_heritability', what = 'parameters')
```

</div>

<div class='panel panel-default padded-panel'>
For each DGP, we simulated data with $n = 500$ samples, $p = 10$ features, and varied $\sigma_{\tau}$ such that the proportion of variance explained in $\tau(X)$ by $X$ (i.e., $\text{var}(\mathbb{E}[\tau_i \mid X]) / \text{var}(\tau_i)$) varied across the values $\{0.2, 0.4, 0.6, 0.8, 1\}$. All simulations were run for 100 replicates.
</div>

## Methods

<div class='panel panel-default padded-panel'>

### Distilled Causal Forest {.tabset .tabset-pills .tabset-circle .tabset-recipe}

#### `r fontawesome::fa('readme', fill = 'white')` {.tabset .tabset-pills}

Distilled Causal Forest refers to a Causal Distillation Tree [Huang et al. (2025)]() using causal forests as the first-stage learner and CART as the second stage learner. Both pruned and unpruned versions of Distilled Causal Forest were run for comparison.

#### `r fontawesome::fa('code', fill = 'white')` {.tabset .tabset-pills}

##### Main Simulations - Additive DGP - Varying n

**Function**

```{r, max.height='200px'}
show_recipe(method_objs, 'Distilled Causal Forest', 'Main Simulations - Additive DGP - Varying n', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(method_objs, 'Distilled Causal Forest', 'Main Simulations - Additive DGP - Varying n', what = 'parameters')
```

##### Main Simulations - Additive DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(method_objs, 'Distilled Causal Forest', 'Main Simulations - Additive DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(method_objs, 'Distilled Causal Forest', 'Main Simulations - Additive DGP - Varying tau_heritability', what = 'parameters')
```

##### Main Simulations - Additive with Linear Covariates DGP - Varying n

**Function**

```{r, max.height='200px'}
show_recipe(method_objs, 'Distilled Causal Forest', 'Main Simulations - Additive with Linear Covariates DGP - Varying n', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(method_objs, 'Distilled Causal Forest', 'Main Simulations - Additive with Linear Covariates DGP - Varying n', what = 'parameters')
```

##### Main Simulations - Additive with Linear Covariates DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(method_objs, 'Distilled Causal Forest', 'Main Simulations - Additive with Linear Covariates DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(method_objs, 'Distilled Causal Forest', 'Main Simulations - Additive with Linear Covariates DGP - Varying tau_heritability', what = 'parameters')
```

##### Main Simulations - AND DGP - Varying n

**Function**

```{r, max.height='200px'}
show_recipe(method_objs, 'Distilled Causal Forest', 'Main Simulations - AND DGP - Varying n', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(method_objs, 'Distilled Causal Forest', 'Main Simulations - AND DGP - Varying n', what = 'parameters')
```

##### Main Simulations - AND DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(method_objs, 'Distilled Causal Forest', 'Main Simulations - AND DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(method_objs, 'Distilled Causal Forest', 'Main Simulations - AND DGP - Varying tau_heritability', what = 'parameters')
```

##### Main Simulations - AND with Linear Covariates DGP - Varying n

**Function**

```{r, max.height='200px'}
show_recipe(method_objs, 'Distilled Causal Forest', 'Main Simulations - AND with Linear Covariates DGP - Varying n', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(method_objs, 'Distilled Causal Forest', 'Main Simulations - AND with Linear Covariates DGP - Varying n', what = 'parameters')
```

##### Main Simulations - AND with Linear Covariates DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(method_objs, 'Distilled Causal Forest', 'Main Simulations - AND with Linear Covariates DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(method_objs, 'Distilled Causal Forest', 'Main Simulations - AND with Linear Covariates DGP - Varying tau_heritability', what = 'parameters')
```

##### Main Simulations - OR DGP - Varying n

**Function**

```{r, max.height='200px'}
show_recipe(method_objs, 'Distilled Causal Forest', 'Main Simulations - OR DGP - Varying n', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(method_objs, 'Distilled Causal Forest', 'Main Simulations - OR DGP - Varying n', what = 'parameters')
```

##### Main Simulations - OR DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(method_objs, 'Distilled Causal Forest', 'Main Simulations - OR DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(method_objs, 'Distilled Causal Forest', 'Main Simulations - OR DGP - Varying tau_heritability', what = 'parameters')
```

##### Main Simulations - OR with Linear Covariates DGP - Varying n

**Function**

```{r, max.height='200px'}
show_recipe(method_objs, 'Distilled Causal Forest', 'Main Simulations - OR with Linear Covariates DGP - Varying n', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(method_objs, 'Distilled Causal Forest', 'Main Simulations - OR with Linear Covariates DGP - Varying n', what = 'parameters')
```

##### Main Simulations - OR with Linear Covariates DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(method_objs, 'Distilled Causal Forest', 'Main Simulations - OR with Linear Covariates DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(method_objs, 'Distilled Causal Forest', 'Main Simulations - OR with Linear Covariates DGP - Varying tau_heritability', what = 'parameters')
```

##### Rulefit Simulations - Additive DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(method_objs, 'Distilled Causal Forest', 'Rulefit Simulations - Additive DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(method_objs, 'Distilled Causal Forest', 'Rulefit Simulations - Additive DGP - Varying tau_heritability', what = 'parameters')
```

##### Rulefit Simulations - Additive with Linear Covariates DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(method_objs, 'Distilled Causal Forest', 'Rulefit Simulations - Additive with Linear Covariates DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(method_objs, 'Distilled Causal Forest', 'Rulefit Simulations - Additive with Linear Covariates DGP - Varying tau_heritability', what = 'parameters')
```

##### Rulefit Simulations - AND DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(method_objs, 'Distilled Causal Forest', 'Rulefit Simulations - AND DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(method_objs, 'Distilled Causal Forest', 'Rulefit Simulations - AND DGP - Varying tau_heritability', what = 'parameters')
```

##### Rulefit Simulations - AND with Linear Covariates DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(method_objs, 'Distilled Causal Forest', 'Rulefit Simulations - AND with Linear Covariates DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(method_objs, 'Distilled Causal Forest', 'Rulefit Simulations - AND with Linear Covariates DGP - Varying tau_heritability', what = 'parameters')
```

##### Rulefit Simulations - OR DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(method_objs, 'Distilled Causal Forest', 'Rulefit Simulations - OR DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(method_objs, 'Distilled Causal Forest', 'Rulefit Simulations - OR DGP - Varying tau_heritability', what = 'parameters')
```

##### Rulefit Simulations - OR with Linear Covariates DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(method_objs, 'Distilled Causal Forest', 'Rulefit Simulations - OR with Linear Covariates DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(method_objs, 'Distilled Causal Forest', 'Rulefit Simulations - OR with Linear Covariates DGP - Varying tau_heritability', what = 'parameters')
```

</div>

<div class='panel panel-default padded-panel'>

### Distilled Rboost {.tabset .tabset-pills .tabset-circle .tabset-recipe}

#### `r fontawesome::fa('readme', fill = 'white')` {.tabset .tabset-pills}

Distilled Rboost refers to a Causal Distillation Tree [Huang et al. (2025)]() using an R-learner with boosting (rboost) as the first-stage learner and CART as the second stage learner. Both pruned and unpruned versions of Distilled Rboost were run for comparison.

#### `r fontawesome::fa('code', fill = 'white')` {.tabset .tabset-pills}

##### Main Simulations - Additive DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(method_objs, 'Distilled Rboost', 'Main Simulations - Additive DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(method_objs, 'Distilled Rboost', 'Main Simulations - Additive DGP - Varying tau_heritability', what = 'parameters')
```

##### Main Simulations - Additive with Linear Covariates DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(method_objs, 'Distilled Rboost', 'Main Simulations - Additive with Linear Covariates DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(method_objs, 'Distilled Rboost', 'Main Simulations - Additive with Linear Covariates DGP - Varying tau_heritability', what = 'parameters')
```

##### Main Simulations - AND DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(method_objs, 'Distilled Rboost', 'Main Simulations - AND DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(method_objs, 'Distilled Rboost', 'Main Simulations - AND DGP - Varying tau_heritability', what = 'parameters')
```

##### Main Simulations - AND with Linear Covariates DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(method_objs, 'Distilled Rboost', 'Main Simulations - AND with Linear Covariates DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(method_objs, 'Distilled Rboost', 'Main Simulations - AND with Linear Covariates DGP - Varying tau_heritability', what = 'parameters')
```

##### Main Simulations - OR DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(method_objs, 'Distilled Rboost', 'Main Simulations - OR DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(method_objs, 'Distilled Rboost', 'Main Simulations - OR DGP - Varying tau_heritability', what = 'parameters')
```

##### Main Simulations - OR with Linear Covariates DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(method_objs, 'Distilled Rboost', 'Main Simulations - OR with Linear Covariates DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(method_objs, 'Distilled Rboost', 'Main Simulations - OR with Linear Covariates DGP - Varying tau_heritability', what = 'parameters')
```

</div>

<div class='panel panel-default padded-panel'>

### Causal Tree {.tabset .tabset-pills .tabset-circle .tabset-recipe}

#### `r fontawesome::fa('readme', fill = 'white')` {.tabset .tabset-pills}

Causal trees [Athey and Imbens (2016)](https://www.pnas.org/doi/10.1073/pnas.1510489113) are a type of decision tree, which were specifically designed to estimate heterogeneous treatment effects and subgroups with either reduced or elevated treatment effects. We implement causal trees using the `causalTree` R package with the default parameters shown on [GitHub](https://github.com/susanathey/causalTree).

#### `r fontawesome::fa('code', fill = 'white')` {.tabset .tabset-pills}

##### Main Simulations - Additive DGP - Varying n

**Function**

```{r, max.height='200px'}
show_recipe(method_objs, 'Causal Tree', 'Main Simulations - Additive DGP - Varying n', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(method_objs, 'Causal Tree', 'Main Simulations - Additive DGP - Varying n', what = 'parameters')
```

##### Main Simulations - Additive DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(method_objs, 'Causal Tree', 'Main Simulations - Additive DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(method_objs, 'Causal Tree', 'Main Simulations - Additive DGP - Varying tau_heritability', what = 'parameters')
```

##### Main Simulations - Additive with Linear Covariates DGP - Varying n

**Function**

```{r, max.height='200px'}
show_recipe(method_objs, 'Causal Tree', 'Main Simulations - Additive with Linear Covariates DGP - Varying n', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(method_objs, 'Causal Tree', 'Main Simulations - Additive with Linear Covariates DGP - Varying n', what = 'parameters')
```

##### Main Simulations - Additive with Linear Covariates DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(method_objs, 'Causal Tree', 'Main Simulations - Additive with Linear Covariates DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(method_objs, 'Causal Tree', 'Main Simulations - Additive with Linear Covariates DGP - Varying tau_heritability', what = 'parameters')
```

##### Main Simulations - AND DGP - Varying n

**Function**

```{r, max.height='200px'}
show_recipe(method_objs, 'Causal Tree', 'Main Simulations - AND DGP - Varying n', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(method_objs, 'Causal Tree', 'Main Simulations - AND DGP - Varying n', what = 'parameters')
```

##### Main Simulations - AND DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(method_objs, 'Causal Tree', 'Main Simulations - AND DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(method_objs, 'Causal Tree', 'Main Simulations - AND DGP - Varying tau_heritability', what = 'parameters')
```

##### Main Simulations - AND with Linear Covariates DGP - Varying n

**Function**

```{r, max.height='200px'}
show_recipe(method_objs, 'Causal Tree', 'Main Simulations - AND with Linear Covariates DGP - Varying n', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(method_objs, 'Causal Tree', 'Main Simulations - AND with Linear Covariates DGP - Varying n', what = 'parameters')
```

##### Main Simulations - AND with Linear Covariates DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(method_objs, 'Causal Tree', 'Main Simulations - AND with Linear Covariates DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(method_objs, 'Causal Tree', 'Main Simulations - AND with Linear Covariates DGP - Varying tau_heritability', what = 'parameters')
```

##### Main Simulations - OR DGP - Varying n

**Function**

```{r, max.height='200px'}
show_recipe(method_objs, 'Causal Tree', 'Main Simulations - OR DGP - Varying n', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(method_objs, 'Causal Tree', 'Main Simulations - OR DGP - Varying n', what = 'parameters')
```

##### Main Simulations - OR DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(method_objs, 'Causal Tree', 'Main Simulations - OR DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(method_objs, 'Causal Tree', 'Main Simulations - OR DGP - Varying tau_heritability', what = 'parameters')
```

##### Main Simulations - OR with Linear Covariates DGP - Varying n

**Function**

```{r, max.height='200px'}
show_recipe(method_objs, 'Causal Tree', 'Main Simulations - OR with Linear Covariates DGP - Varying n', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(method_objs, 'Causal Tree', 'Main Simulations - OR with Linear Covariates DGP - Varying n', what = 'parameters')
```

##### Main Simulations - OR with Linear Covariates DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(method_objs, 'Causal Tree', 'Main Simulations - OR with Linear Covariates DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(method_objs, 'Causal Tree', 'Main Simulations - OR with Linear Covariates DGP - Varying tau_heritability', what = 'parameters')
```

</div>

<div class='panel panel-default padded-panel'>

### Virtual Twins {.tabset .tabset-pills .tabset-circle .tabset-recipe}

#### `r fontawesome::fa('readme', fill = 'white')` {.tabset .tabset-pills}

Similar to the proposed distillation approach, virtual twins is a two-stage approach for estimating subgroups with heterogeneous treatment effects, where the first-stage aims to predict the response for each unit's "virtual twin" using random forests. We implement the approach described in [Foster et al. (2011)](https://onlinelibrary.wiley.com/doi/10.1002/sim.4322).

#### `r fontawesome::fa('code', fill = 'white')` {.tabset .tabset-pills}

##### Main Simulations - Additive DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(method_objs, 'Virtual Twins', 'Main Simulations - Additive DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(method_objs, 'Virtual Twins', 'Main Simulations - Additive DGP - Varying tau_heritability', what = 'parameters')
```

##### Main Simulations - Additive with Linear Covariates DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(method_objs, 'Virtual Twins', 'Main Simulations - Additive with Linear Covariates DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(method_objs, 'Virtual Twins', 'Main Simulations - Additive with Linear Covariates DGP - Varying tau_heritability', what = 'parameters')
```

##### Main Simulations - AND DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(method_objs, 'Virtual Twins', 'Main Simulations - AND DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(method_objs, 'Virtual Twins', 'Main Simulations - AND DGP - Varying tau_heritability', what = 'parameters')
```

##### Main Simulations - AND with Linear Covariates DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(method_objs, 'Virtual Twins', 'Main Simulations - AND with Linear Covariates DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(method_objs, 'Virtual Twins', 'Main Simulations - AND with Linear Covariates DGP - Varying tau_heritability', what = 'parameters')
```

##### Main Simulations - OR DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(method_objs, 'Virtual Twins', 'Main Simulations - OR DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(method_objs, 'Virtual Twins', 'Main Simulations - OR DGP - Varying tau_heritability', what = 'parameters')
```

##### Main Simulations - OR with Linear Covariates DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(method_objs, 'Virtual Twins', 'Main Simulations - OR with Linear Covariates DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(method_objs, 'Virtual Twins', 'Main Simulations - OR with Linear Covariates DGP - Varying tau_heritability', what = 'parameters')
```

</div>

<div class='panel panel-default padded-panel'>

### Linear Regression {.tabset .tabset-pills .tabset-circle .tabset-recipe}

#### `r fontawesome::fa('readme', fill = 'white')` {.tabset .tabset-pills}

As a baseline comparison method, we implemented a linear regression model with all main effects $X$ and their interactions with the treatment variable $Z$ in order to predict the response $Y$.In the linear regression, we defined a selected subgroup feature to be any covariate whose interaction with the treatment variable yielded a significant p-value ($p < 0.05$).

#### `r fontawesome::fa('code', fill = 'white')` {.tabset .tabset-pills}

**Function**

```{r, max.height='200px'}
show_recipe(method_objs, 'Linear Regression', 'Main Simulations - Additive DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(method_objs, 'Linear Regression', 'Main Simulations - Additive DGP - Varying tau_heritability', what = 'parameters')
```

</div>

<div class='panel panel-default padded-panel'>

### Lasso {.tabset .tabset-pills .tabset-circle .tabset-recipe}

#### `r fontawesome::fa('readme', fill = 'white')` {.tabset .tabset-pills}

As a baseline comparison method, we implemented a Lasso regression model [(Tibshirani (1996))](https://www.jstor.org/stable/2346178) with all main effects $X$ and their interactions with the treatment variable $Z$ in order to predict the response $Y$.In the Lasso regression, we defined a selected subgroup feature to be any covariate whose interaction with the treatment variable received a non-zero coefficient.

#### `r fontawesome::fa('code', fill = 'white')` {.tabset .tabset-pills}

**Function**

```{r, max.height='200px'}
show_recipe(method_objs, 'Lasso', 'Main Simulations - Additive DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(method_objs, 'Lasso', 'Main Simulations - Additive DGP - Varying tau_heritability', what = 'parameters')
```

</div>

<div class='panel panel-default padded-panel'>

### Distilled Causal Forest (with Rulefit (linear + rules, max depth = 3)) {.tabset .tabset-pills .tabset-circle .tabset-recipe}

#### `r fontawesome::fa('readme', fill = 'white')` {.tabset .tabset-pills}

In this variant of Distilled Causal Forest, we use Rulefit as the second-stage learner (instead of the default CART). The default Rulefit parameters are used except that `type = "both"` and `maxdepth = 3` in `pre::pre()`. That is, both linear structure and rules with maximum depth of 3 are allowed in the fitted Rulefit model.

#### `r fontawesome::fa('code', fill = 'white')` {.tabset .tabset-pills}

##### Rulefit Simulations - Additive DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(method_objs, 'Distilled Causal Forest (rulefit v1)', 'Rulefit Simulations - Additive DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(method_objs, 'Distilled Causal Forest (rulefit v1)', 'Rulefit Simulations - Additive DGP - Varying tau_heritability', what = 'parameters')
```

##### Rulefit Simulations - Additive with Linear Covariates DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(method_objs, 'Distilled Causal Forest (rulefit v1)', 'Rulefit Simulations - Additive with Linear Covariates DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(method_objs, 'Distilled Causal Forest (rulefit v1)', 'Rulefit Simulations - Additive with Linear Covariates DGP - Varying tau_heritability', what = 'parameters')
```

##### Rulefit Simulations - AND DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(method_objs, 'Distilled Causal Forest (rulefit v1)', 'Rulefit Simulations - AND DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(method_objs, 'Distilled Causal Forest (rulefit v1)', 'Rulefit Simulations - AND DGP - Varying tau_heritability', what = 'parameters')
```

##### Rulefit Simulations - AND with Linear Covariates DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(method_objs, 'Distilled Causal Forest (rulefit v1)', 'Rulefit Simulations - AND with Linear Covariates DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(method_objs, 'Distilled Causal Forest (rulefit v1)', 'Rulefit Simulations - AND with Linear Covariates DGP - Varying tau_heritability', what = 'parameters')
```

##### Rulefit Simulations - OR DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(method_objs, 'Distilled Causal Forest (rulefit v1)', 'Rulefit Simulations - OR DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(method_objs, 'Distilled Causal Forest (rulefit v1)', 'Rulefit Simulations - OR DGP - Varying tau_heritability', what = 'parameters')
```

##### Rulefit Simulations - OR with Linear Covariates DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(method_objs, 'Distilled Causal Forest (rulefit v1)', 'Rulefit Simulations - OR with Linear Covariates DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(method_objs, 'Distilled Causal Forest (rulefit v1)', 'Rulefit Simulations - OR with Linear Covariates DGP - Varying tau_heritability', what = 'parameters')
```

</div>

<div class='panel panel-default padded-panel'>

### Distilled Causal Forest (with Rulefit (linear + rules, max depth = 2)) {.tabset .tabset-pills .tabset-circle .tabset-recipe}

#### `r fontawesome::fa('readme', fill = 'white')` {.tabset .tabset-pills}

In this variant of Distilled Causal Forest, we use Rulefit as the second-stage learner (instead of the default CART). The default Rulefit parameters are used except that `type = "both"` and `maxdepth = 2` in `pre::pre()`. That is, both linear structure and rules with maximum depth of 2 are allowed in the fitted Rulefit model.

#### `r fontawesome::fa('code', fill = 'white')` {.tabset .tabset-pills}

##### Rulefit Simulations - Additive DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(method_objs, 'Distilled Causal Forest (rulefit v2)', 'Rulefit Simulations - Additive DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(method_objs, 'Distilled Causal Forest (rulefit v2)', 'Rulefit Simulations - Additive DGP - Varying tau_heritability', what = 'parameters')
```

##### Rulefit Simulations - Additive with Linear Covariates DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(method_objs, 'Distilled Causal Forest (rulefit v2)', 'Rulefit Simulations - Additive with Linear Covariates DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(method_objs, 'Distilled Causal Forest (rulefit v2)', 'Rulefit Simulations - Additive with Linear Covariates DGP - Varying tau_heritability', what = 'parameters')
```

##### Rulefit Simulations - AND DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(method_objs, 'Distilled Causal Forest (rulefit v2)', 'Rulefit Simulations - AND DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(method_objs, 'Distilled Causal Forest (rulefit v2)', 'Rulefit Simulations - AND DGP - Varying tau_heritability', what = 'parameters')
```

##### Rulefit Simulations - AND with Linear Covariates DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(method_objs, 'Distilled Causal Forest (rulefit v2)', 'Rulefit Simulations - AND with Linear Covariates DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(method_objs, 'Distilled Causal Forest (rulefit v2)', 'Rulefit Simulations - AND with Linear Covariates DGP - Varying tau_heritability', what = 'parameters')
```

##### Rulefit Simulations - OR DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(method_objs, 'Distilled Causal Forest (rulefit v2)', 'Rulefit Simulations - OR DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(method_objs, 'Distilled Causal Forest (rulefit v2)', 'Rulefit Simulations - OR DGP - Varying tau_heritability', what = 'parameters')
```

##### Rulefit Simulations - OR with Linear Covariates DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(method_objs, 'Distilled Causal Forest (rulefit v2)', 'Rulefit Simulations - OR with Linear Covariates DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(method_objs, 'Distilled Causal Forest (rulefit v2)', 'Rulefit Simulations - OR with Linear Covariates DGP - Varying tau_heritability', what = 'parameters')
```

</div>

<div class='panel panel-default padded-panel'>

### Distilled Causal Forest (with Rulefit (rules only, max depth = 3)) {.tabset .tabset-pills .tabset-circle .tabset-recipe}

#### `r fontawesome::fa('readme', fill = 'white')` {.tabset .tabset-pills}

In this variant of Distilled Causal Forest, we use Rulefit as the second-stage learner (instead of the default CART). The default Rulefit parameters are used except that `type = "rules"` and `maxdepth = 3` in `pre::pre()`. That is, only rules with maximum depth of 3 are allowed in the fitted Rulefit model.

#### `r fontawesome::fa('code', fill = 'white')` {.tabset .tabset-pills}

##### Rulefit Simulations - Additive DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(method_objs, 'Distilled Causal Forest (rulefit v3)', 'Rulefit Simulations - Additive DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(method_objs, 'Distilled Causal Forest (rulefit v3)', 'Rulefit Simulations - Additive DGP - Varying tau_heritability', what = 'parameters')
```

##### Rulefit Simulations - Additive with Linear Covariates DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(method_objs, 'Distilled Causal Forest (rulefit v3)', 'Rulefit Simulations - Additive with Linear Covariates DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(method_objs, 'Distilled Causal Forest (rulefit v3)', 'Rulefit Simulations - Additive with Linear Covariates DGP - Varying tau_heritability', what = 'parameters')
```

##### Rulefit Simulations - AND DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(method_objs, 'Distilled Causal Forest (rulefit v3)', 'Rulefit Simulations - AND DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(method_objs, 'Distilled Causal Forest (rulefit v3)', 'Rulefit Simulations - AND DGP - Varying tau_heritability', what = 'parameters')
```

##### Rulefit Simulations - AND with Linear Covariates DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(method_objs, 'Distilled Causal Forest (rulefit v3)', 'Rulefit Simulations - AND with Linear Covariates DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(method_objs, 'Distilled Causal Forest (rulefit v3)', 'Rulefit Simulations - AND with Linear Covariates DGP - Varying tau_heritability', what = 'parameters')
```

##### Rulefit Simulations - OR DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(method_objs, 'Distilled Causal Forest (rulefit v3)', 'Rulefit Simulations - OR DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(method_objs, 'Distilled Causal Forest (rulefit v3)', 'Rulefit Simulations - OR DGP - Varying tau_heritability', what = 'parameters')
```

##### Rulefit Simulations - OR with Linear Covariates DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(method_objs, 'Distilled Causal Forest (rulefit v3)', 'Rulefit Simulations - OR with Linear Covariates DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(method_objs, 'Distilled Causal Forest (rulefit v3)', 'Rulefit Simulations - OR with Linear Covariates DGP - Varying tau_heritability', what = 'parameters')
```

</div>

<div class='panel panel-default padded-panel'>

### Distilled Causal Forest (with Rulefit (rules only, max depth = 2)) {.tabset .tabset-pills .tabset-circle .tabset-recipe}

#### `r fontawesome::fa('readme', fill = 'white')` {.tabset .tabset-pills}

In this variant of Distilled Causal Forest, we use Rulefit as the second-stage learner (instead of the default CART). The default Rulefit parameters are used except that `type = "rules"` and `maxdepth = 2` in `pre::pre()`. That is, only rules with maximum depth of 2 are allowed in the fitted Rulefit model.

#### `r fontawesome::fa('code', fill = 'white')` {.tabset .tabset-pills}

##### Rulefit Simulations - Additive DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(method_objs, 'Distilled Causal Forest (rulefit v4)', 'Rulefit Simulations - Additive DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(method_objs, 'Distilled Causal Forest (rulefit v4)', 'Rulefit Simulations - Additive DGP - Varying tau_heritability', what = 'parameters')
```

##### Rulefit Simulations - Additive with Linear Covariates DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(method_objs, 'Distilled Causal Forest (rulefit v4)', 'Rulefit Simulations - Additive with Linear Covariates DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(method_objs, 'Distilled Causal Forest (rulefit v4)', 'Rulefit Simulations - Additive with Linear Covariates DGP - Varying tau_heritability', what = 'parameters')
```

##### Rulefit Simulations - AND DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(method_objs, 'Distilled Causal Forest (rulefit v4)', 'Rulefit Simulations - AND DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(method_objs, 'Distilled Causal Forest (rulefit v4)', 'Rulefit Simulations - AND DGP - Varying tau_heritability', what = 'parameters')
```

##### Rulefit Simulations - AND with Linear Covariates DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(method_objs, 'Distilled Causal Forest (rulefit v4)', 'Rulefit Simulations - AND with Linear Covariates DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(method_objs, 'Distilled Causal Forest (rulefit v4)', 'Rulefit Simulations - AND with Linear Covariates DGP - Varying tau_heritability', what = 'parameters')
```

##### Rulefit Simulations - OR DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(method_objs, 'Distilled Causal Forest (rulefit v4)', 'Rulefit Simulations - OR DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(method_objs, 'Distilled Causal Forest (rulefit v4)', 'Rulefit Simulations - OR DGP - Varying tau_heritability', what = 'parameters')
```

##### Rulefit Simulations - OR with Linear Covariates DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(method_objs, 'Distilled Causal Forest (rulefit v4)', 'Rulefit Simulations - OR with Linear Covariates DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(method_objs, 'Distilled Causal Forest (rulefit v4)', 'Rulefit Simulations - OR with Linear Covariates DGP - Varying tau_heritability', what = 'parameters')
```

</div>

## Evaluation

<div class='panel panel-default padded-panel'>
At a high-level, we evaluate subgroup estimation performance according to three criteria:

1. **Selected subgroup features**: Whether the features used to define the estimated subgroups match the features used to define the true subgroups (i.e., $X^{(1)}$ and $X^{(2)}$)

2. **Estimated subgroup thresholds**: Whether the estimated subgroup thresholds are close to the true subgroup thresholds

3. **Estimated subgroup average treatment effects (ATEs)**: Whether the estimated subgroup average treatment effects are close to the true subgroup average treatment effects

For tree-based methods, we also evaluate the estimated subgroups under the oracle pruning strategy (i.e., pruning the tree to have depth 2).
</div>

<div class='panel panel-default padded-panel'>

### Subgroup Feature Selection Errors Summary {.tabset .tabset-pills .tabset-circle .tabset-recipe}

#### `r fontawesome::fa('readme', fill = 'white')` {.tabset .tabset-pills}

For each subgroup estimation method, we report the features used to characterize the estimated subgroups and compare them to the true subgroup features (i.e., $X^{(1)}$ and $X^{(2)}$). In particular, we compute the $F_1$ score, the number of true positives, and the number of false positives between the estimated subgroup features and the true subgroup features.

#### `r fontawesome::fa('code', fill = 'white')` {.tabset .tabset-pills}

##### Main Simulations - Additive DGP - Varying n

**Function**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Subgroup Feature Selection Errors Summary', 'Main Simulations - Additive DGP - Varying n', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Subgroup Feature Selection Errors Summary', 'Main Simulations - Additive DGP - Varying n', what = 'parameters')
```

##### Main Simulations - Additive DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Subgroup Feature Selection Errors Summary', 'Main Simulations - Additive DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Subgroup Feature Selection Errors Summary', 'Main Simulations - Additive DGP - Varying tau_heritability', what = 'parameters')
```

##### Main Simulations - Additive with Linear Covariates DGP - Varying n

**Function**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Subgroup Feature Selection Errors Summary', 'Main Simulations - Additive with Linear Covariates DGP - Varying n', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Subgroup Feature Selection Errors Summary', 'Main Simulations - Additive with Linear Covariates DGP - Varying n', what = 'parameters')
```

##### Main Simulations - Additive with Linear Covariates DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Subgroup Feature Selection Errors Summary', 'Main Simulations - Additive with Linear Covariates DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Subgroup Feature Selection Errors Summary', 'Main Simulations - Additive with Linear Covariates DGP - Varying tau_heritability', what = 'parameters')
```

##### Main Simulations - AND DGP - Varying n

**Function**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Subgroup Feature Selection Errors Summary', 'Main Simulations - AND DGP - Varying n', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Subgroup Feature Selection Errors Summary', 'Main Simulations - AND DGP - Varying n', what = 'parameters')
```

##### Main Simulations - AND DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Subgroup Feature Selection Errors Summary', 'Main Simulations - AND DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Subgroup Feature Selection Errors Summary', 'Main Simulations - AND DGP - Varying tau_heritability', what = 'parameters')
```

##### Main Simulations - AND with Linear Covariates DGP - Varying n

**Function**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Subgroup Feature Selection Errors Summary', 'Main Simulations - AND with Linear Covariates DGP - Varying n', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Subgroup Feature Selection Errors Summary', 'Main Simulations - AND with Linear Covariates DGP - Varying n', what = 'parameters')
```

##### Main Simulations - AND with Linear Covariates DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Subgroup Feature Selection Errors Summary', 'Main Simulations - AND with Linear Covariates DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Subgroup Feature Selection Errors Summary', 'Main Simulations - AND with Linear Covariates DGP - Varying tau_heritability', what = 'parameters')
```

##### Main Simulations - OR DGP - Varying n

**Function**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Subgroup Feature Selection Errors Summary', 'Main Simulations - OR DGP - Varying n', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Subgroup Feature Selection Errors Summary', 'Main Simulations - OR DGP - Varying n', what = 'parameters')
```

##### Main Simulations - OR DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Subgroup Feature Selection Errors Summary', 'Main Simulations - OR DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Subgroup Feature Selection Errors Summary', 'Main Simulations - OR DGP - Varying tau_heritability', what = 'parameters')
```

##### Main Simulations - OR with Linear Covariates DGP - Varying n

**Function**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Subgroup Feature Selection Errors Summary', 'Main Simulations - OR with Linear Covariates DGP - Varying n', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Subgroup Feature Selection Errors Summary', 'Main Simulations - OR with Linear Covariates DGP - Varying n', what = 'parameters')
```

##### Main Simulations - OR with Linear Covariates DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Subgroup Feature Selection Errors Summary', 'Main Simulations - OR with Linear Covariates DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Subgroup Feature Selection Errors Summary', 'Main Simulations - OR with Linear Covariates DGP - Varying tau_heritability', what = 'parameters')
```

##### Rulefit Simulations - Additive DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Subgroup Feature Selection Errors Summary', 'Rulefit Simulations - Additive DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Subgroup Feature Selection Errors Summary', 'Rulefit Simulations - Additive DGP - Varying tau_heritability', what = 'parameters')
```

##### Rulefit Simulations - Additive with Linear Covariates DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Subgroup Feature Selection Errors Summary', 'Rulefit Simulations - Additive with Linear Covariates DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Subgroup Feature Selection Errors Summary', 'Rulefit Simulations - Additive with Linear Covariates DGP - Varying tau_heritability', what = 'parameters')
```

##### Rulefit Simulations - AND DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Subgroup Feature Selection Errors Summary', 'Rulefit Simulations - AND DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Subgroup Feature Selection Errors Summary', 'Rulefit Simulations - AND DGP - Varying tau_heritability', what = 'parameters')
```

##### Rulefit Simulations - AND with Linear Covariates DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Subgroup Feature Selection Errors Summary', 'Rulefit Simulations - AND with Linear Covariates DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Subgroup Feature Selection Errors Summary', 'Rulefit Simulations - AND with Linear Covariates DGP - Varying tau_heritability', what = 'parameters')
```

##### Rulefit Simulations - OR DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Subgroup Feature Selection Errors Summary', 'Rulefit Simulations - OR DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Subgroup Feature Selection Errors Summary', 'Rulefit Simulations - OR DGP - Varying tau_heritability', what = 'parameters')
```

##### Rulefit Simulations - OR with Linear Covariates DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Subgroup Feature Selection Errors Summary', 'Rulefit Simulations - OR with Linear Covariates DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Subgroup Feature Selection Errors Summary', 'Rulefit Simulations - OR with Linear Covariates DGP - Varying tau_heritability', what = 'parameters')
```

##### Varying nrep crossfits - Additive DGP-Distilled Rboost (crossfit) - Varying tau_heritability-nreps_crossfit

**Function**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Subgroup Feature Selection Errors Summary', 'Varying nrep crossfits - Additive DGP-Distilled Rboost (crossfit) - Varying tau_heritability-nreps_crossfit', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Subgroup Feature Selection Errors Summary', 'Varying nrep crossfits - Additive DGP-Distilled Rboost (crossfit) - Varying tau_heritability-nreps_crossfit', what = 'parameters')
```

##### Varying nrep crossfits - AND DGP-Distilled Rboost (crossfit) - Varying tau_heritability-nreps_crossfit

**Function**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Subgroup Feature Selection Errors Summary', 'Varying nrep crossfits - AND DGP-Distilled Rboost (crossfit) - Varying tau_heritability-nreps_crossfit', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Subgroup Feature Selection Errors Summary', 'Varying nrep crossfits - AND DGP-Distilled Rboost (crossfit) - Varying tau_heritability-nreps_crossfit', what = 'parameters')
```

##### Varying nrep crossfits - OR DGP-Distilled Rboost (crossfit) - Varying tau_heritability-nreps_crossfit

**Function**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Subgroup Feature Selection Errors Summary', 'Varying nrep crossfits - OR DGP-Distilled Rboost (crossfit) - Varying tau_heritability-nreps_crossfit', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Subgroup Feature Selection Errors Summary', 'Varying nrep crossfits - OR DGP-Distilled Rboost (crossfit) - Varying tau_heritability-nreps_crossfit', what = 'parameters')
```

</div>

<div class='panel panel-default padded-panel'>

### Subgroup Feature Selection Errors Summary (max depth = 2) {.tabset .tabset-pills .tabset-circle .tabset-recipe}

#### `r fontawesome::fa('readme', fill = 'white')` {.tabset .tabset-pills}

For each subgroup estimation method, we report the features used to characterize the estimated subgroups, where the tree has been truncated to a maximum (oracle) depth of 2 for tree-based methods, and compare them to the true subgroup features (i.e., $X^{(1)}$ and $X^{(2)}$). In particular, we compute the $F_1$ score, the number of true positives, and the number of false positives between the estimated subgroup features and the true subgroup features.

#### `r fontawesome::fa('code', fill = 'white')` {.tabset .tabset-pills}

##### Main Simulations - Additive DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Subgroup Feature Selection Errors Summary (max depth = 2)', 'Main Simulations - Additive DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Subgroup Feature Selection Errors Summary (max depth = 2)', 'Main Simulations - Additive DGP - Varying tau_heritability', what = 'parameters')
```

##### Main Simulations - Additive with Linear Covariates DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Subgroup Feature Selection Errors Summary (max depth = 2)', 'Main Simulations - Additive with Linear Covariates DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Subgroup Feature Selection Errors Summary (max depth = 2)', 'Main Simulations - Additive with Linear Covariates DGP - Varying tau_heritability', what = 'parameters')
```

##### Main Simulations - AND DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Subgroup Feature Selection Errors Summary (max depth = 2)', 'Main Simulations - AND DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Subgroup Feature Selection Errors Summary (max depth = 2)', 'Main Simulations - AND DGP - Varying tau_heritability', what = 'parameters')
```

##### Main Simulations - AND with Linear Covariates DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Subgroup Feature Selection Errors Summary (max depth = 2)', 'Main Simulations - AND with Linear Covariates DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Subgroup Feature Selection Errors Summary (max depth = 2)', 'Main Simulations - AND with Linear Covariates DGP - Varying tau_heritability', what = 'parameters')
```

##### Main Simulations - OR DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Subgroup Feature Selection Errors Summary (max depth = 2)', 'Main Simulations - OR DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Subgroup Feature Selection Errors Summary (max depth = 2)', 'Main Simulations - OR DGP - Varying tau_heritability', what = 'parameters')
```

##### Main Simulations - OR with Linear Covariates DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Subgroup Feature Selection Errors Summary (max depth = 2)', 'Main Simulations - OR with Linear Covariates DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Subgroup Feature Selection Errors Summary (max depth = 2)', 'Main Simulations - OR with Linear Covariates DGP - Varying tau_heritability', what = 'parameters')
```

</div>

<div class='panel panel-default padded-panel'>

### Thresholds Summary {.tabset .tabset-pills .tabset-circle .tabset-recipe}

#### `r fontawesome::fa('readme', fill = 'white')` {.tabset .tabset-pills}

For each tree-based method, we record all split thresholds for each feature that were used to define the subgroups.

#### `r fontawesome::fa('code', fill = 'white')` {.tabset .tabset-pills}

##### Main Simulations - Additive DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Thresholds Summary', 'Main Simulations - Additive DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Thresholds Summary', 'Main Simulations - Additive DGP - Varying tau_heritability', what = 'parameters')
```

##### Main Simulations - Additive with Linear Covariates DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Thresholds Summary', 'Main Simulations - Additive with Linear Covariates DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Thresholds Summary', 'Main Simulations - Additive with Linear Covariates DGP - Varying tau_heritability', what = 'parameters')
```

##### Main Simulations - AND DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Thresholds Summary', 'Main Simulations - AND DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Thresholds Summary', 'Main Simulations - AND DGP - Varying tau_heritability', what = 'parameters')
```

##### Main Simulations - AND with Linear Covariates DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Thresholds Summary', 'Main Simulations - AND with Linear Covariates DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Thresholds Summary', 'Main Simulations - AND with Linear Covariates DGP - Varying tau_heritability', what = 'parameters')
```

##### Main Simulations - OR DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Thresholds Summary', 'Main Simulations - OR DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Thresholds Summary', 'Main Simulations - OR DGP - Varying tau_heritability', what = 'parameters')
```

##### Main Simulations - OR with Linear Covariates DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Thresholds Summary', 'Main Simulations - OR with Linear Covariates DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Thresholds Summary', 'Main Simulations - OR with Linear Covariates DGP - Varying tau_heritability', what = 'parameters')
```

##### Rulefit Simulations - Additive DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Thresholds Summary', 'Rulefit Simulations - Additive DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Thresholds Summary', 'Rulefit Simulations - Additive DGP - Varying tau_heritability', what = 'parameters')
```

##### Rulefit Simulations - Additive with Linear Covariates DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Thresholds Summary', 'Rulefit Simulations - Additive with Linear Covariates DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Thresholds Summary', 'Rulefit Simulations - Additive with Linear Covariates DGP - Varying tau_heritability', what = 'parameters')
```

##### Rulefit Simulations - AND DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Thresholds Summary', 'Rulefit Simulations - AND DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Thresholds Summary', 'Rulefit Simulations - AND DGP - Varying tau_heritability', what = 'parameters')
```

##### Rulefit Simulations - AND with Linear Covariates DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Thresholds Summary', 'Rulefit Simulations - AND with Linear Covariates DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Thresholds Summary', 'Rulefit Simulations - AND with Linear Covariates DGP - Varying tau_heritability', what = 'parameters')
```

##### Rulefit Simulations - OR DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Thresholds Summary', 'Rulefit Simulations - OR DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Thresholds Summary', 'Rulefit Simulations - OR DGP - Varying tau_heritability', what = 'parameters')
```

##### Rulefit Simulations - OR with Linear Covariates DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Thresholds Summary', 'Rulefit Simulations - OR with Linear Covariates DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Thresholds Summary', 'Rulefit Simulations - OR with Linear Covariates DGP - Varying tau_heritability', what = 'parameters')
```

##### Varying nrep crossfits - Additive DGP-Distilled Rboost (crossfit) - Varying tau_heritability-nreps_crossfit

**Function**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Thresholds Summary', 'Varying nrep crossfits - Additive DGP-Distilled Rboost (crossfit) - Varying tau_heritability-nreps_crossfit', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Thresholds Summary', 'Varying nrep crossfits - Additive DGP-Distilled Rboost (crossfit) - Varying tau_heritability-nreps_crossfit', what = 'parameters')
```

##### Varying nrep crossfits - AND DGP-Distilled Rboost (crossfit) - Varying tau_heritability-nreps_crossfit

**Function**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Thresholds Summary', 'Varying nrep crossfits - AND DGP-Distilled Rboost (crossfit) - Varying tau_heritability-nreps_crossfit', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Thresholds Summary', 'Varying nrep crossfits - AND DGP-Distilled Rboost (crossfit) - Varying tau_heritability-nreps_crossfit', what = 'parameters')
```

##### Varying nrep crossfits - OR DGP-Distilled Rboost (crossfit) - Varying tau_heritability-nreps_crossfit

**Function**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Thresholds Summary', 'Varying nrep crossfits - OR DGP-Distilled Rboost (crossfit) - Varying tau_heritability-nreps_crossfit', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Thresholds Summary', 'Varying nrep crossfits - OR DGP-Distilled Rboost (crossfit) - Varying tau_heritability-nreps_crossfit', what = 'parameters')
```

</div>

<div class='panel panel-default padded-panel'>

### Thresholds Summary (max depth = 2) {.tabset .tabset-pills .tabset-circle .tabset-recipe}

#### `r fontawesome::fa('readme', fill = 'white')` {.tabset .tabset-pills}

For each tree-based method, we record the split thresholds for each feature that were used to define the subgroups, up to a maximum tree depth of 2.

#### `r fontawesome::fa('code', fill = 'white')` {.tabset .tabset-pills}

##### Main Simulations - Additive DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Thresholds Summary (max depth = 2)', 'Main Simulations - Additive DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Thresholds Summary (max depth = 2)', 'Main Simulations - Additive DGP - Varying tau_heritability', what = 'parameters')
```

##### Main Simulations - Additive with Linear Covariates DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Thresholds Summary (max depth = 2)', 'Main Simulations - Additive with Linear Covariates DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Thresholds Summary (max depth = 2)', 'Main Simulations - Additive with Linear Covariates DGP - Varying tau_heritability', what = 'parameters')
```

##### Main Simulations - AND DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Thresholds Summary (max depth = 2)', 'Main Simulations - AND DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Thresholds Summary (max depth = 2)', 'Main Simulations - AND DGP - Varying tau_heritability', what = 'parameters')
```

##### Main Simulations - AND with Linear Covariates DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Thresholds Summary (max depth = 2)', 'Main Simulations - AND with Linear Covariates DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Thresholds Summary (max depth = 2)', 'Main Simulations - AND with Linear Covariates DGP - Varying tau_heritability', what = 'parameters')
```

##### Main Simulations - OR DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Thresholds Summary (max depth = 2)', 'Main Simulations - OR DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Thresholds Summary (max depth = 2)', 'Main Simulations - OR DGP - Varying tau_heritability', what = 'parameters')
```

##### Main Simulations - OR with Linear Covariates DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Thresholds Summary (max depth = 2)', 'Main Simulations - OR with Linear Covariates DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Thresholds Summary (max depth = 2)', 'Main Simulations - OR with Linear Covariates DGP - Varying tau_heritability', what = 'parameters')
```

</div>

<div class='panel panel-default padded-panel'>

### Threshold Distances Summary {.tabset .tabset-pills .tabset-circle .tabset-recipe}

#### `r fontawesome::fa('readme', fill = 'white')` {.tabset .tabset-pills}

For each tree-based method and for each of the true subgroup features (i.e., $X^{(1)}$ and $X^{(2)}$), we compute the root mean squared error between the "best" estimated split threshold for that feature and the true split threshold. In short, we compute the distance between each estimated threshold and each true threshold and report the minimum of these distances for each feature's true threshold.

#### `r fontawesome::fa('code', fill = 'white')` {.tabset .tabset-pills}

##### Main Simulations - Additive DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Threshold Distances Summary', 'Main Simulations - Additive DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Threshold Distances Summary', 'Main Simulations - Additive DGP - Varying tau_heritability', what = 'parameters')
```

##### Main Simulations - Additive with Linear Covariates DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Threshold Distances Summary', 'Main Simulations - Additive with Linear Covariates DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Threshold Distances Summary', 'Main Simulations - Additive with Linear Covariates DGP - Varying tau_heritability', what = 'parameters')
```

##### Main Simulations - AND DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Threshold Distances Summary', 'Main Simulations - AND DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Threshold Distances Summary', 'Main Simulations - AND DGP - Varying tau_heritability', what = 'parameters')
```

##### Main Simulations - AND with Linear Covariates DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Threshold Distances Summary', 'Main Simulations - AND with Linear Covariates DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Threshold Distances Summary', 'Main Simulations - AND with Linear Covariates DGP - Varying tau_heritability', what = 'parameters')
```

##### Main Simulations - OR DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Threshold Distances Summary', 'Main Simulations - OR DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Threshold Distances Summary', 'Main Simulations - OR DGP - Varying tau_heritability', what = 'parameters')
```

##### Main Simulations - OR with Linear Covariates DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Threshold Distances Summary', 'Main Simulations - OR with Linear Covariates DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Threshold Distances Summary', 'Main Simulations - OR with Linear Covariates DGP - Varying tau_heritability', what = 'parameters')
```

##### Rulefit Simulations - Additive DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Threshold Distances Summary', 'Rulefit Simulations - Additive DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Threshold Distances Summary', 'Rulefit Simulations - Additive DGP - Varying tau_heritability', what = 'parameters')
```

##### Rulefit Simulations - Additive with Linear Covariates DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Threshold Distances Summary', 'Rulefit Simulations - Additive with Linear Covariates DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Threshold Distances Summary', 'Rulefit Simulations - Additive with Linear Covariates DGP - Varying tau_heritability', what = 'parameters')
```

##### Rulefit Simulations - AND DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Threshold Distances Summary', 'Rulefit Simulations - AND DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Threshold Distances Summary', 'Rulefit Simulations - AND DGP - Varying tau_heritability', what = 'parameters')
```

##### Rulefit Simulations - AND with Linear Covariates DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Threshold Distances Summary', 'Rulefit Simulations - AND with Linear Covariates DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Threshold Distances Summary', 'Rulefit Simulations - AND with Linear Covariates DGP - Varying tau_heritability', what = 'parameters')
```

##### Rulefit Simulations - OR DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Threshold Distances Summary', 'Rulefit Simulations - OR DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Threshold Distances Summary', 'Rulefit Simulations - OR DGP - Varying tau_heritability', what = 'parameters')
```

##### Rulefit Simulations - OR with Linear Covariates DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Threshold Distances Summary', 'Rulefit Simulations - OR with Linear Covariates DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Threshold Distances Summary', 'Rulefit Simulations - OR with Linear Covariates DGP - Varying tau_heritability', what = 'parameters')
```

##### Varying nrep crossfits - Additive DGP-Distilled Rboost (crossfit) - Varying tau_heritability-nreps_crossfit

**Function**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Threshold Distances Summary', 'Varying nrep crossfits - Additive DGP-Distilled Rboost (crossfit) - Varying tau_heritability-nreps_crossfit', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Threshold Distances Summary', 'Varying nrep crossfits - Additive DGP-Distilled Rboost (crossfit) - Varying tau_heritability-nreps_crossfit', what = 'parameters')
```

##### Varying nrep crossfits - AND DGP-Distilled Rboost (crossfit) - Varying tau_heritability-nreps_crossfit

**Function**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Threshold Distances Summary', 'Varying nrep crossfits - AND DGP-Distilled Rboost (crossfit) - Varying tau_heritability-nreps_crossfit', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Threshold Distances Summary', 'Varying nrep crossfits - AND DGP-Distilled Rboost (crossfit) - Varying tau_heritability-nreps_crossfit', what = 'parameters')
```

##### Varying nrep crossfits - OR DGP-Distilled Rboost (crossfit) - Varying tau_heritability-nreps_crossfit

**Function**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Threshold Distances Summary', 'Varying nrep crossfits - OR DGP-Distilled Rboost (crossfit) - Varying tau_heritability-nreps_crossfit', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Threshold Distances Summary', 'Varying nrep crossfits - OR DGP-Distilled Rboost (crossfit) - Varying tau_heritability-nreps_crossfit', what = 'parameters')
```

</div>

<div class='panel panel-default padded-panel'>

### Threshold Distances Summary (max depth = 2) {.tabset .tabset-pills .tabset-circle .tabset-recipe}

#### `r fontawesome::fa('readme', fill = 'white')` {.tabset .tabset-pills}

For each tree-based method and for each of the true subgroup features (i.e., $X^{(1)}$ and $X^{(2)}$), we compute the root mean squared error between the "best" estimated split threshold for that feature and the true split threshold. In short, we compute the distance between each estimated threshold and each true threshold and report the minimum of these distances for each feature's true threshold. Only estimated splits that occurred in the first two depths of the tree are considered.

#### `r fontawesome::fa('code', fill = 'white')` {.tabset .tabset-pills}

##### Main Simulations - Additive DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Threshold Distances Summary (max depth = 2)', 'Main Simulations - Additive DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Threshold Distances Summary (max depth = 2)', 'Main Simulations - Additive DGP - Varying tau_heritability', what = 'parameters')
```

##### Main Simulations - Additive with Linear Covariates DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Threshold Distances Summary (max depth = 2)', 'Main Simulations - Additive with Linear Covariates DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Threshold Distances Summary (max depth = 2)', 'Main Simulations - Additive with Linear Covariates DGP - Varying tau_heritability', what = 'parameters')
```

##### Main Simulations - AND DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Threshold Distances Summary (max depth = 2)', 'Main Simulations - AND DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Threshold Distances Summary (max depth = 2)', 'Main Simulations - AND DGP - Varying tau_heritability', what = 'parameters')
```

##### Main Simulations - AND with Linear Covariates DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Threshold Distances Summary (max depth = 2)', 'Main Simulations - AND with Linear Covariates DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Threshold Distances Summary (max depth = 2)', 'Main Simulations - AND with Linear Covariates DGP - Varying tau_heritability', what = 'parameters')
```

##### Main Simulations - OR DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Threshold Distances Summary (max depth = 2)', 'Main Simulations - OR DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Threshold Distances Summary (max depth = 2)', 'Main Simulations - OR DGP - Varying tau_heritability', what = 'parameters')
```

##### Main Simulations - OR with Linear Covariates DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Threshold Distances Summary (max depth = 2)', 'Main Simulations - OR with Linear Covariates DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Threshold Distances Summary (max depth = 2)', 'Main Simulations - OR with Linear Covariates DGP - Varying tau_heritability', what = 'parameters')
```

</div>

<div class='panel panel-default padded-panel'>

### Subgroup ATE Errors Summary {.tabset .tabset-pills .tabset-circle .tabset-recipe}

#### `r fontawesome::fa('readme', fill = 'white')` {.tabset .tabset-pills}

For each method, we report the estimation error in subgroup average treatment effect (ATE), measured as the root mean squared error between the true subgroup ATEs and the estimated subgroup ATEs.

#### `r fontawesome::fa('code', fill = 'white')` {.tabset .tabset-pills}

##### Main Simulations - Additive DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Subgroup ATE Errors Summary', 'Main Simulations - Additive DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Subgroup ATE Errors Summary', 'Main Simulations - Additive DGP - Varying tau_heritability', what = 'parameters')
```

##### Main Simulations - Additive with Linear Covariates DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Subgroup ATE Errors Summary', 'Main Simulations - Additive with Linear Covariates DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Subgroup ATE Errors Summary', 'Main Simulations - Additive with Linear Covariates DGP - Varying tau_heritability', what = 'parameters')
```

##### Main Simulations - AND DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Subgroup ATE Errors Summary', 'Main Simulations - AND DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Subgroup ATE Errors Summary', 'Main Simulations - AND DGP - Varying tau_heritability', what = 'parameters')
```

##### Main Simulations - AND with Linear Covariates DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Subgroup ATE Errors Summary', 'Main Simulations - AND with Linear Covariates DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Subgroup ATE Errors Summary', 'Main Simulations - AND with Linear Covariates DGP - Varying tau_heritability', what = 'parameters')
```

##### Main Simulations - OR DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Subgroup ATE Errors Summary', 'Main Simulations - OR DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Subgroup ATE Errors Summary', 'Main Simulations - OR DGP - Varying tau_heritability', what = 'parameters')
```

##### Main Simulations - OR with Linear Covariates DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Subgroup ATE Errors Summary', 'Main Simulations - OR with Linear Covariates DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Subgroup ATE Errors Summary', 'Main Simulations - OR with Linear Covariates DGP - Varying tau_heritability', what = 'parameters')
```

##### Rulefit Simulations - Additive DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Subgroup ATE Errors Summary', 'Rulefit Simulations - Additive DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Subgroup ATE Errors Summary', 'Rulefit Simulations - Additive DGP - Varying tau_heritability', what = 'parameters')
```

##### Rulefit Simulations - Additive with Linear Covariates DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Subgroup ATE Errors Summary', 'Rulefit Simulations - Additive with Linear Covariates DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Subgroup ATE Errors Summary', 'Rulefit Simulations - Additive with Linear Covariates DGP - Varying tau_heritability', what = 'parameters')
```

##### Rulefit Simulations - AND DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Subgroup ATE Errors Summary', 'Rulefit Simulations - AND DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Subgroup ATE Errors Summary', 'Rulefit Simulations - AND DGP - Varying tau_heritability', what = 'parameters')
```

##### Rulefit Simulations - AND with Linear Covariates DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Subgroup ATE Errors Summary', 'Rulefit Simulations - AND with Linear Covariates DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Subgroup ATE Errors Summary', 'Rulefit Simulations - AND with Linear Covariates DGP - Varying tau_heritability', what = 'parameters')
```

##### Rulefit Simulations - OR DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Subgroup ATE Errors Summary', 'Rulefit Simulations - OR DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Subgroup ATE Errors Summary', 'Rulefit Simulations - OR DGP - Varying tau_heritability', what = 'parameters')
```

##### Rulefit Simulations - OR with Linear Covariates DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Subgroup ATE Errors Summary', 'Rulefit Simulations - OR with Linear Covariates DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Subgroup ATE Errors Summary', 'Rulefit Simulations - OR with Linear Covariates DGP - Varying tau_heritability', what = 'parameters')
```

##### Varying nrep crossfits - Additive DGP-Distilled Rboost (crossfit) - Varying tau_heritability-nreps_crossfit

**Function**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Subgroup ATE Errors Summary', 'Varying nrep crossfits - Additive DGP-Distilled Rboost (crossfit) - Varying tau_heritability-nreps_crossfit', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Subgroup ATE Errors Summary', 'Varying nrep crossfits - Additive DGP-Distilled Rboost (crossfit) - Varying tau_heritability-nreps_crossfit', what = 'parameters')
```

##### Varying nrep crossfits - AND DGP-Distilled Rboost (crossfit) - Varying tau_heritability-nreps_crossfit

**Function**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Subgroup ATE Errors Summary', 'Varying nrep crossfits - AND DGP-Distilled Rboost (crossfit) - Varying tau_heritability-nreps_crossfit', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Subgroup ATE Errors Summary', 'Varying nrep crossfits - AND DGP-Distilled Rboost (crossfit) - Varying tau_heritability-nreps_crossfit', what = 'parameters')
```

##### Varying nrep crossfits - OR DGP-Distilled Rboost (crossfit) - Varying tau_heritability-nreps_crossfit

**Function**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Subgroup ATE Errors Summary', 'Varying nrep crossfits - OR DGP-Distilled Rboost (crossfit) - Varying tau_heritability-nreps_crossfit', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Subgroup ATE Errors Summary', 'Varying nrep crossfits - OR DGP-Distilled Rboost (crossfit) - Varying tau_heritability-nreps_crossfit', what = 'parameters')
```

</div>

<div class='panel panel-default padded-panel'>

### Subgroup ATE Errors Summary (max depth = 2) {.tabset .tabset-pills .tabset-circle .tabset-recipe}

#### `r fontawesome::fa('readme', fill = 'white')` {.tabset .tabset-pills}

For each method, we report the estimation error in subgroup average treatment effect (ATE), measured as the root mean squared error between the true subgroup ATEs and the estimated subgroup ATEs. For tree-based methods, the trees have been pruned to a maximum depth of 2.

#### `r fontawesome::fa('code', fill = 'white')` {.tabset .tabset-pills}

##### Main Simulations - Additive DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Subgroup ATE Errors Summary (max depth = 2)', 'Main Simulations - Additive DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Subgroup ATE Errors Summary (max depth = 2)', 'Main Simulations - Additive DGP - Varying tau_heritability', what = 'parameters')
```

##### Main Simulations - Additive with Linear Covariates DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Subgroup ATE Errors Summary (max depth = 2)', 'Main Simulations - Additive with Linear Covariates DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Subgroup ATE Errors Summary (max depth = 2)', 'Main Simulations - Additive with Linear Covariates DGP - Varying tau_heritability', what = 'parameters')
```

##### Main Simulations - AND DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Subgroup ATE Errors Summary (max depth = 2)', 'Main Simulations - AND DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Subgroup ATE Errors Summary (max depth = 2)', 'Main Simulations - AND DGP - Varying tau_heritability', what = 'parameters')
```

##### Main Simulations - AND with Linear Covariates DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Subgroup ATE Errors Summary (max depth = 2)', 'Main Simulations - AND with Linear Covariates DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Subgroup ATE Errors Summary (max depth = 2)', 'Main Simulations - AND with Linear Covariates DGP - Varying tau_heritability', what = 'parameters')
```

##### Main Simulations - OR DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Subgroup ATE Errors Summary (max depth = 2)', 'Main Simulations - OR DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Subgroup ATE Errors Summary (max depth = 2)', 'Main Simulations - OR DGP - Varying tau_heritability', what = 'parameters')
```

##### Main Simulations - OR with Linear Covariates DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Subgroup ATE Errors Summary (max depth = 2)', 'Main Simulations - OR with Linear Covariates DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(eval_objs, 'Subgroup ATE Errors Summary (max depth = 2)', 'Main Simulations - OR with Linear Covariates DGP - Varying tau_heritability', what = 'parameters')
```

</div>

## Visualizations

<div class='panel panel-default padded-panel'>

### Number of Subgroups Plot {.tabset .tabset-pills .tabset-circle .tabset-recipe}

#### `r fontawesome::fa('readme', fill = 'white')` {.tabset .tabset-pills}

We plot the total number of estimated subgroups alongside the number of true positive/false positive subgroup features.

#### `r fontawesome::fa('code', fill = 'white')` {.tabset .tabset-pills}

##### Main Simulations - Additive DGP - Varying n

**Function**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Number of Subgroups Plot', 'Main Simulations - Additive DGP - Varying n', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Number of Subgroups Plot', 'Main Simulations - Additive DGP - Varying n', what = 'parameters')
```

##### Main Simulations - Additive with Linear Covariates DGP - Varying n

**Function**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Number of Subgroups Plot', 'Main Simulations - Additive with Linear Covariates DGP - Varying n', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Number of Subgroups Plot', 'Main Simulations - Additive with Linear Covariates DGP - Varying n', what = 'parameters')
```

##### Main Simulations - AND DGP - Varying n

**Function**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Number of Subgroups Plot', 'Main Simulations - AND DGP - Varying n', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Number of Subgroups Plot', 'Main Simulations - AND DGP - Varying n', what = 'parameters')
```

##### Main Simulations - AND with Linear Covariates DGP - Varying n

**Function**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Number of Subgroups Plot', 'Main Simulations - AND with Linear Covariates DGP - Varying n', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Number of Subgroups Plot', 'Main Simulations - AND with Linear Covariates DGP - Varying n', what = 'parameters')
```

##### Main Simulations - OR DGP - Varying n

**Function**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Number of Subgroups Plot', 'Main Simulations - OR DGP - Varying n', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Number of Subgroups Plot', 'Main Simulations - OR DGP - Varying n', what = 'parameters')
```

##### Main Simulations - OR with Linear Covariates DGP - Varying n

**Function**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Number of Subgroups Plot', 'Main Simulations - OR with Linear Covariates DGP - Varying n', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Number of Subgroups Plot', 'Main Simulations - OR with Linear Covariates DGP - Varying n', what = 'parameters')
```

</div>

<div class='panel panel-default padded-panel'>

### Subgroup Feature Selection Errors Plot {.tabset .tabset-pills .tabset-circle .tabset-recipe}

#### `r fontawesome::fa('readme', fill = 'white')` {.tabset .tabset-pills}

For each subgroup estimation method, we plot the $F_1$ score, the number of true positives, and the number of false positives between the estimated subgroup features and the true subgroup features.

#### `r fontawesome::fa('code', fill = 'white')` {.tabset .tabset-pills}

##### Main Simulations - Additive DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Subgroup Feature Selection Errors Plot', 'Main Simulations - Additive DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Subgroup Feature Selection Errors Plot', 'Main Simulations - Additive DGP - Varying tau_heritability', what = 'parameters')
```

##### Main Simulations - Additive with Linear Covariates DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Subgroup Feature Selection Errors Plot', 'Main Simulations - Additive with Linear Covariates DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Subgroup Feature Selection Errors Plot', 'Main Simulations - Additive with Linear Covariates DGP - Varying tau_heritability', what = 'parameters')
```

##### Main Simulations - AND DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Subgroup Feature Selection Errors Plot', 'Main Simulations - AND DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Subgroup Feature Selection Errors Plot', 'Main Simulations - AND DGP - Varying tau_heritability', what = 'parameters')
```

##### Main Simulations - AND with Linear Covariates DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Subgroup Feature Selection Errors Plot', 'Main Simulations - AND with Linear Covariates DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Subgroup Feature Selection Errors Plot', 'Main Simulations - AND with Linear Covariates DGP - Varying tau_heritability', what = 'parameters')
```

##### Main Simulations - OR DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Subgroup Feature Selection Errors Plot', 'Main Simulations - OR DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Subgroup Feature Selection Errors Plot', 'Main Simulations - OR DGP - Varying tau_heritability', what = 'parameters')
```

##### Main Simulations - OR with Linear Covariates DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Subgroup Feature Selection Errors Plot', 'Main Simulations - OR with Linear Covariates DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Subgroup Feature Selection Errors Plot', 'Main Simulations - OR with Linear Covariates DGP - Varying tau_heritability', what = 'parameters')
```

##### Rulefit Simulations - Additive DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Subgroup Feature Selection Errors Plot', 'Rulefit Simulations - Additive DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Subgroup Feature Selection Errors Plot', 'Rulefit Simulations - Additive DGP - Varying tau_heritability', what = 'parameters')
```

##### Rulefit Simulations - Additive with Linear Covariates DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Subgroup Feature Selection Errors Plot', 'Rulefit Simulations - Additive with Linear Covariates DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Subgroup Feature Selection Errors Plot', 'Rulefit Simulations - Additive with Linear Covariates DGP - Varying tau_heritability', what = 'parameters')
```

##### Rulefit Simulations - AND DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Subgroup Feature Selection Errors Plot', 'Rulefit Simulations - AND DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Subgroup Feature Selection Errors Plot', 'Rulefit Simulations - AND DGP - Varying tau_heritability', what = 'parameters')
```

##### Rulefit Simulations - AND with Linear Covariates DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Subgroup Feature Selection Errors Plot', 'Rulefit Simulations - AND with Linear Covariates DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Subgroup Feature Selection Errors Plot', 'Rulefit Simulations - AND with Linear Covariates DGP - Varying tau_heritability', what = 'parameters')
```

##### Rulefit Simulations - OR DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Subgroup Feature Selection Errors Plot', 'Rulefit Simulations - OR DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Subgroup Feature Selection Errors Plot', 'Rulefit Simulations - OR DGP - Varying tau_heritability', what = 'parameters')
```

##### Rulefit Simulations - OR with Linear Covariates DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Subgroup Feature Selection Errors Plot', 'Rulefit Simulations - OR with Linear Covariates DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Subgroup Feature Selection Errors Plot', 'Rulefit Simulations - OR with Linear Covariates DGP - Varying tau_heritability', what = 'parameters')
```

##### Varying nrep crossfits - Additive DGP-Distilled Rboost (crossfit) - Varying tau_heritability-nreps_crossfit

**Function**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Subgroup Feature Selection Errors Plot', 'Varying nrep crossfits - Additive DGP-Distilled Rboost (crossfit) - Varying tau_heritability-nreps_crossfit', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Subgroup Feature Selection Errors Plot', 'Varying nrep crossfits - Additive DGP-Distilled Rboost (crossfit) - Varying tau_heritability-nreps_crossfit', what = 'parameters')
```

##### Varying nrep crossfits - AND DGP-Distilled Rboost (crossfit) - Varying tau_heritability-nreps_crossfit

**Function**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Subgroup Feature Selection Errors Plot', 'Varying nrep crossfits - AND DGP-Distilled Rboost (crossfit) - Varying tau_heritability-nreps_crossfit', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Subgroup Feature Selection Errors Plot', 'Varying nrep crossfits - AND DGP-Distilled Rboost (crossfit) - Varying tau_heritability-nreps_crossfit', what = 'parameters')
```

##### Varying nrep crossfits - OR DGP-Distilled Rboost (crossfit) - Varying tau_heritability-nreps_crossfit

**Function**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Subgroup Feature Selection Errors Plot', 'Varying nrep crossfits - OR DGP-Distilled Rboost (crossfit) - Varying tau_heritability-nreps_crossfit', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Subgroup Feature Selection Errors Plot', 'Varying nrep crossfits - OR DGP-Distilled Rboost (crossfit) - Varying tau_heritability-nreps_crossfit', what = 'parameters')
```

</div>

<div class='panel panel-default padded-panel'>

### Subgroup Feature Selection Errors Plot (max depth = 2) {.tabset .tabset-pills .tabset-circle .tabset-recipe}

#### `r fontawesome::fa('readme', fill = 'white')` {.tabset .tabset-pills}

For each subgroup estimation method, we plot the $F_1$ score, the number of true positives, and the number of false positives between the estimated subgroup features and the true subgroup features. Here, the tree-based methods have been pruned to a maximum depth of 2.

#### `r fontawesome::fa('code', fill = 'white')` {.tabset .tabset-pills}

##### Main Simulations - Additive DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Subgroup Feature Selection Errors Plot (max depth = 2)', 'Main Simulations - Additive DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Subgroup Feature Selection Errors Plot (max depth = 2)', 'Main Simulations - Additive DGP - Varying tau_heritability', what = 'parameters')
```

##### Main Simulations - Additive with Linear Covariates DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Subgroup Feature Selection Errors Plot (max depth = 2)', 'Main Simulations - Additive with Linear Covariates DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Subgroup Feature Selection Errors Plot (max depth = 2)', 'Main Simulations - Additive with Linear Covariates DGP - Varying tau_heritability', what = 'parameters')
```

##### Main Simulations - AND DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Subgroup Feature Selection Errors Plot (max depth = 2)', 'Main Simulations - AND DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Subgroup Feature Selection Errors Plot (max depth = 2)', 'Main Simulations - AND DGP - Varying tau_heritability', what = 'parameters')
```

##### Main Simulations - AND with Linear Covariates DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Subgroup Feature Selection Errors Plot (max depth = 2)', 'Main Simulations - AND with Linear Covariates DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Subgroup Feature Selection Errors Plot (max depth = 2)', 'Main Simulations - AND with Linear Covariates DGP - Varying tau_heritability', what = 'parameters')
```

##### Main Simulations - OR DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Subgroup Feature Selection Errors Plot (max depth = 2)', 'Main Simulations - OR DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Subgroup Feature Selection Errors Plot (max depth = 2)', 'Main Simulations - OR DGP - Varying tau_heritability', what = 'parameters')
```

##### Main Simulations - OR with Linear Covariates DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Subgroup Feature Selection Errors Plot (max depth = 2)', 'Main Simulations - OR with Linear Covariates DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Subgroup Feature Selection Errors Plot (max depth = 2)', 'Main Simulations - OR with Linear Covariates DGP - Varying tau_heritability', what = 'parameters')
```

</div>

<div class='panel panel-default padded-panel'>

### Subgroup Thresholds Distribution Plot {.tabset .tabset-pills .tabset-circle .tabset-recipe}

#### `r fontawesome::fa('readme', fill = 'white')` {.tabset .tabset-pills}

For each tree-based method and for each of the true subgroup features (i.e., $X^{(1)}$ and $X^{(2)}$), we plot the distribution of estimated split thresholds. The true split thresholds are shown as dashed horizontal lines.

#### `r fontawesome::fa('code', fill = 'white')` {.tabset .tabset-pills}

##### Main Simulations - Additive DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Subgroup Thresholds Distribution Plot', 'Main Simulations - Additive DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Subgroup Thresholds Distribution Plot', 'Main Simulations - Additive DGP - Varying tau_heritability', what = 'parameters')
```

##### Main Simulations - Additive with Linear Covariates DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Subgroup Thresholds Distribution Plot', 'Main Simulations - Additive with Linear Covariates DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Subgroup Thresholds Distribution Plot', 'Main Simulations - Additive with Linear Covariates DGP - Varying tau_heritability', what = 'parameters')
```

##### Main Simulations - AND DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Subgroup Thresholds Distribution Plot', 'Main Simulations - AND DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Subgroup Thresholds Distribution Plot', 'Main Simulations - AND DGP - Varying tau_heritability', what = 'parameters')
```

##### Main Simulations - AND with Linear Covariates DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Subgroup Thresholds Distribution Plot', 'Main Simulations - AND with Linear Covariates DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Subgroup Thresholds Distribution Plot', 'Main Simulations - AND with Linear Covariates DGP - Varying tau_heritability', what = 'parameters')
```

##### Main Simulations - OR DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Subgroup Thresholds Distribution Plot', 'Main Simulations - OR DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Subgroup Thresholds Distribution Plot', 'Main Simulations - OR DGP - Varying tau_heritability', what = 'parameters')
```

##### Main Simulations - OR with Linear Covariates DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Subgroup Thresholds Distribution Plot', 'Main Simulations - OR with Linear Covariates DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Subgroup Thresholds Distribution Plot', 'Main Simulations - OR with Linear Covariates DGP - Varying tau_heritability', what = 'parameters')
```

##### Rulefit Simulations - Additive DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Subgroup Thresholds Distribution Plot', 'Rulefit Simulations - Additive DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Subgroup Thresholds Distribution Plot', 'Rulefit Simulations - Additive DGP - Varying tau_heritability', what = 'parameters')
```

##### Rulefit Simulations - Additive with Linear Covariates DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Subgroup Thresholds Distribution Plot', 'Rulefit Simulations - Additive with Linear Covariates DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Subgroup Thresholds Distribution Plot', 'Rulefit Simulations - Additive with Linear Covariates DGP - Varying tau_heritability', what = 'parameters')
```

##### Rulefit Simulations - AND DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Subgroup Thresholds Distribution Plot', 'Rulefit Simulations - AND DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Subgroup Thresholds Distribution Plot', 'Rulefit Simulations - AND DGP - Varying tau_heritability', what = 'parameters')
```

##### Rulefit Simulations - AND with Linear Covariates DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Subgroup Thresholds Distribution Plot', 'Rulefit Simulations - AND with Linear Covariates DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Subgroup Thresholds Distribution Plot', 'Rulefit Simulations - AND with Linear Covariates DGP - Varying tau_heritability', what = 'parameters')
```

##### Rulefit Simulations - OR DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Subgroup Thresholds Distribution Plot', 'Rulefit Simulations - OR DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Subgroup Thresholds Distribution Plot', 'Rulefit Simulations - OR DGP - Varying tau_heritability', what = 'parameters')
```

##### Rulefit Simulations - OR with Linear Covariates DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Subgroup Thresholds Distribution Plot', 'Rulefit Simulations - OR with Linear Covariates DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Subgroup Thresholds Distribution Plot', 'Rulefit Simulations - OR with Linear Covariates DGP - Varying tau_heritability', what = 'parameters')
```

</div>

<div class='panel panel-default padded-panel'>

### Subgroup Thresholds Distribution Plot (max depth = 2) {.tabset .tabset-pills .tabset-circle .tabset-recipe}

#### `r fontawesome::fa('readme', fill = 'white')` {.tabset .tabset-pills}

For each tree-based method abd for each of the true subgroup features (i.e., $X^{(1)}$ and $X^{(2)}$), we plot the distribution of estimated split thresholds. The true split thresholds are shown as dashed horizontal lines. Only splits that occurred in the first two depths of the tree are considered.

#### `r fontawesome::fa('code', fill = 'white')` {.tabset .tabset-pills}

##### Main Simulations - Additive DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Subgroup Thresholds Distribution Plot (max depth = 2)', 'Main Simulations - Additive DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Subgroup Thresholds Distribution Plot (max depth = 2)', 'Main Simulations - Additive DGP - Varying tau_heritability', what = 'parameters')
```

##### Main Simulations - Additive with Linear Covariates DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Subgroup Thresholds Distribution Plot (max depth = 2)', 'Main Simulations - Additive with Linear Covariates DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Subgroup Thresholds Distribution Plot (max depth = 2)', 'Main Simulations - Additive with Linear Covariates DGP - Varying tau_heritability', what = 'parameters')
```

##### Main Simulations - AND DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Subgroup Thresholds Distribution Plot (max depth = 2)', 'Main Simulations - AND DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Subgroup Thresholds Distribution Plot (max depth = 2)', 'Main Simulations - AND DGP - Varying tau_heritability', what = 'parameters')
```

##### Main Simulations - AND with Linear Covariates DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Subgroup Thresholds Distribution Plot (max depth = 2)', 'Main Simulations - AND with Linear Covariates DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Subgroup Thresholds Distribution Plot (max depth = 2)', 'Main Simulations - AND with Linear Covariates DGP - Varying tau_heritability', what = 'parameters')
```

##### Main Simulations - OR DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Subgroup Thresholds Distribution Plot (max depth = 2)', 'Main Simulations - OR DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Subgroup Thresholds Distribution Plot (max depth = 2)', 'Main Simulations - OR DGP - Varying tau_heritability', what = 'parameters')
```

##### Main Simulations - OR with Linear Covariates DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Subgroup Thresholds Distribution Plot (max depth = 2)', 'Main Simulations - OR with Linear Covariates DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Subgroup Thresholds Distribution Plot (max depth = 2)', 'Main Simulations - OR with Linear Covariates DGP - Varying tau_heritability', what = 'parameters')
```

</div>

<div class='panel panel-default padded-panel'>

### Threshold Distances Plot {.tabset .tabset-pills .tabset-circle .tabset-recipe}

#### `r fontawesome::fa('readme', fill = 'white')` {.tabset .tabset-pills}

For each tree-based method and for each of the true subgroup features (i.e., $X^{(1)}$ and $X^{(2)}$), we plot the root mean squared error between the "best" estimated split threshold for that feature and the true split threshold. In short, we compute the distance between each estimated threshold and each true threshold and plot the minimum of these distances for each feature's true threshold.

#### `r fontawesome::fa('code', fill = 'white')` {.tabset .tabset-pills}

##### Main Simulations - Additive DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Threshold Distances Plot', 'Main Simulations - Additive DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Threshold Distances Plot', 'Main Simulations - Additive DGP - Varying tau_heritability', what = 'parameters')
```

##### Main Simulations - Additive with Linear Covariates DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Threshold Distances Plot', 'Main Simulations - Additive with Linear Covariates DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Threshold Distances Plot', 'Main Simulations - Additive with Linear Covariates DGP - Varying tau_heritability', what = 'parameters')
```

##### Main Simulations - AND DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Threshold Distances Plot', 'Main Simulations - AND DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Threshold Distances Plot', 'Main Simulations - AND DGP - Varying tau_heritability', what = 'parameters')
```

##### Main Simulations - AND with Linear Covariates DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Threshold Distances Plot', 'Main Simulations - AND with Linear Covariates DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Threshold Distances Plot', 'Main Simulations - AND with Linear Covariates DGP - Varying tau_heritability', what = 'parameters')
```

##### Main Simulations - OR DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Threshold Distances Plot', 'Main Simulations - OR DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Threshold Distances Plot', 'Main Simulations - OR DGP - Varying tau_heritability', what = 'parameters')
```

##### Main Simulations - OR with Linear Covariates DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Threshold Distances Plot', 'Main Simulations - OR with Linear Covariates DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Threshold Distances Plot', 'Main Simulations - OR with Linear Covariates DGP - Varying tau_heritability', what = 'parameters')
```

##### Rulefit Simulations - Additive DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Threshold Distances Plot', 'Rulefit Simulations - Additive DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Threshold Distances Plot', 'Rulefit Simulations - Additive DGP - Varying tau_heritability', what = 'parameters')
```

##### Rulefit Simulations - Additive with Linear Covariates DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Threshold Distances Plot', 'Rulefit Simulations - Additive with Linear Covariates DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Threshold Distances Plot', 'Rulefit Simulations - Additive with Linear Covariates DGP - Varying tau_heritability', what = 'parameters')
```

##### Rulefit Simulations - AND DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Threshold Distances Plot', 'Rulefit Simulations - AND DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Threshold Distances Plot', 'Rulefit Simulations - AND DGP - Varying tau_heritability', what = 'parameters')
```

##### Rulefit Simulations - AND with Linear Covariates DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Threshold Distances Plot', 'Rulefit Simulations - AND with Linear Covariates DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Threshold Distances Plot', 'Rulefit Simulations - AND with Linear Covariates DGP - Varying tau_heritability', what = 'parameters')
```

##### Rulefit Simulations - OR DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Threshold Distances Plot', 'Rulefit Simulations - OR DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Threshold Distances Plot', 'Rulefit Simulations - OR DGP - Varying tau_heritability', what = 'parameters')
```

##### Rulefit Simulations - OR with Linear Covariates DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Threshold Distances Plot', 'Rulefit Simulations - OR with Linear Covariates DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Threshold Distances Plot', 'Rulefit Simulations - OR with Linear Covariates DGP - Varying tau_heritability', what = 'parameters')
```

##### Varying nrep crossfits - Additive DGP-Distilled Rboost (crossfit) - Varying tau_heritability-nreps_crossfit

**Function**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Threshold Distances Plot', 'Varying nrep crossfits - Additive DGP-Distilled Rboost (crossfit) - Varying tau_heritability-nreps_crossfit', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Threshold Distances Plot', 'Varying nrep crossfits - Additive DGP-Distilled Rboost (crossfit) - Varying tau_heritability-nreps_crossfit', what = 'parameters')
```

##### Varying nrep crossfits - AND DGP-Distilled Rboost (crossfit) - Varying tau_heritability-nreps_crossfit

**Function**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Threshold Distances Plot', 'Varying nrep crossfits - AND DGP-Distilled Rboost (crossfit) - Varying tau_heritability-nreps_crossfit', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Threshold Distances Plot', 'Varying nrep crossfits - AND DGP-Distilled Rboost (crossfit) - Varying tau_heritability-nreps_crossfit', what = 'parameters')
```

##### Varying nrep crossfits - OR DGP-Distilled Rboost (crossfit) - Varying tau_heritability-nreps_crossfit

**Function**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Threshold Distances Plot', 'Varying nrep crossfits - OR DGP-Distilled Rboost (crossfit) - Varying tau_heritability-nreps_crossfit', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Threshold Distances Plot', 'Varying nrep crossfits - OR DGP-Distilled Rboost (crossfit) - Varying tau_heritability-nreps_crossfit', what = 'parameters')
```

</div>

<div class='panel panel-default padded-panel'>

### Threshold Distances Plot (max depth = 2) {.tabset .tabset-pills .tabset-circle .tabset-recipe}

#### `r fontawesome::fa('readme', fill = 'white')` {.tabset .tabset-pills}

For each tree-based method and for each of the true subgroup features (i.e., $X^{(1)}$ and $X^{(2)}$), we plot the root mean squared error between the "best" estimated split threshold for that feature and the true split threshold. In short, we compute the distance between each estimated threshold and each true threshold and plot the minimum of these distances for each feature's true threshold. Only estimated splits that occurred in the first two depths of the tree are considered.

#### `r fontawesome::fa('code', fill = 'white')` {.tabset .tabset-pills}

##### Main Simulations - Additive DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Threshold Distances Plot (max depth = 2)', 'Main Simulations - Additive DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Threshold Distances Plot (max depth = 2)', 'Main Simulations - Additive DGP - Varying tau_heritability', what = 'parameters')
```

##### Main Simulations - Additive with Linear Covariates DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Threshold Distances Plot (max depth = 2)', 'Main Simulations - Additive with Linear Covariates DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Threshold Distances Plot (max depth = 2)', 'Main Simulations - Additive with Linear Covariates DGP - Varying tau_heritability', what = 'parameters')
```

##### Main Simulations - AND DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Threshold Distances Plot (max depth = 2)', 'Main Simulations - AND DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Threshold Distances Plot (max depth = 2)', 'Main Simulations - AND DGP - Varying tau_heritability', what = 'parameters')
```

##### Main Simulations - AND with Linear Covariates DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Threshold Distances Plot (max depth = 2)', 'Main Simulations - AND with Linear Covariates DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Threshold Distances Plot (max depth = 2)', 'Main Simulations - AND with Linear Covariates DGP - Varying tau_heritability', what = 'parameters')
```

##### Main Simulations - OR DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Threshold Distances Plot (max depth = 2)', 'Main Simulations - OR DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Threshold Distances Plot (max depth = 2)', 'Main Simulations - OR DGP - Varying tau_heritability', what = 'parameters')
```

##### Main Simulations - OR with Linear Covariates DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Threshold Distances Plot (max depth = 2)', 'Main Simulations - OR with Linear Covariates DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Threshold Distances Plot (max depth = 2)', 'Main Simulations - OR with Linear Covariates DGP - Varying tau_heritability', what = 'parameters')
```

</div>

<div class='panel panel-default padded-panel'>

### Subgroup ATE Errors Plot {.tabset .tabset-pills .tabset-circle .tabset-recipe}

#### `r fontawesome::fa('readme', fill = 'white')` {.tabset .tabset-pills}

We plot each method's subgroup average treatment effect (ATE) estimation error, measured as the root mean squared error between the true subgroup ATEs and the estimated subgroup ATEs.

#### `r fontawesome::fa('code', fill = 'white')` {.tabset .tabset-pills}

##### Main Simulations - Additive DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Subgroup ATE Errors Plot', 'Main Simulations - Additive DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Subgroup ATE Errors Plot', 'Main Simulations - Additive DGP - Varying tau_heritability', what = 'parameters')
```

##### Main Simulations - Additive with Linear Covariates DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Subgroup ATE Errors Plot', 'Main Simulations - Additive with Linear Covariates DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Subgroup ATE Errors Plot', 'Main Simulations - Additive with Linear Covariates DGP - Varying tau_heritability', what = 'parameters')
```

##### Main Simulations - AND DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Subgroup ATE Errors Plot', 'Main Simulations - AND DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Subgroup ATE Errors Plot', 'Main Simulations - AND DGP - Varying tau_heritability', what = 'parameters')
```

##### Main Simulations - AND with Linear Covariates DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Subgroup ATE Errors Plot', 'Main Simulations - AND with Linear Covariates DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Subgroup ATE Errors Plot', 'Main Simulations - AND with Linear Covariates DGP - Varying tau_heritability', what = 'parameters')
```

##### Main Simulations - OR DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Subgroup ATE Errors Plot', 'Main Simulations - OR DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Subgroup ATE Errors Plot', 'Main Simulations - OR DGP - Varying tau_heritability', what = 'parameters')
```

##### Main Simulations - OR with Linear Covariates DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Subgroup ATE Errors Plot', 'Main Simulations - OR with Linear Covariates DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Subgroup ATE Errors Plot', 'Main Simulations - OR with Linear Covariates DGP - Varying tau_heritability', what = 'parameters')
```

##### Rulefit Simulations - Additive DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Subgroup ATE Errors Plot', 'Rulefit Simulations - Additive DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Subgroup ATE Errors Plot', 'Rulefit Simulations - Additive DGP - Varying tau_heritability', what = 'parameters')
```

##### Rulefit Simulations - Additive with Linear Covariates DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Subgroup ATE Errors Plot', 'Rulefit Simulations - Additive with Linear Covariates DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Subgroup ATE Errors Plot', 'Rulefit Simulations - Additive with Linear Covariates DGP - Varying tau_heritability', what = 'parameters')
```

##### Rulefit Simulations - AND DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Subgroup ATE Errors Plot', 'Rulefit Simulations - AND DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Subgroup ATE Errors Plot', 'Rulefit Simulations - AND DGP - Varying tau_heritability', what = 'parameters')
```

##### Rulefit Simulations - AND with Linear Covariates DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Subgroup ATE Errors Plot', 'Rulefit Simulations - AND with Linear Covariates DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Subgroup ATE Errors Plot', 'Rulefit Simulations - AND with Linear Covariates DGP - Varying tau_heritability', what = 'parameters')
```

##### Rulefit Simulations - OR DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Subgroup ATE Errors Plot', 'Rulefit Simulations - OR DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Subgroup ATE Errors Plot', 'Rulefit Simulations - OR DGP - Varying tau_heritability', what = 'parameters')
```

##### Rulefit Simulations - OR with Linear Covariates DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Subgroup ATE Errors Plot', 'Rulefit Simulations - OR with Linear Covariates DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Subgroup ATE Errors Plot', 'Rulefit Simulations - OR with Linear Covariates DGP - Varying tau_heritability', what = 'parameters')
```

##### Varying nrep crossfits - Additive DGP-Distilled Rboost (crossfit) - Varying tau_heritability-nreps_crossfit

**Function**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Subgroup ATE Errors Plot', 'Varying nrep crossfits - Additive DGP-Distilled Rboost (crossfit) - Varying tau_heritability-nreps_crossfit', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Subgroup ATE Errors Plot', 'Varying nrep crossfits - Additive DGP-Distilled Rboost (crossfit) - Varying tau_heritability-nreps_crossfit', what = 'parameters')
```

##### Varying nrep crossfits - AND DGP-Distilled Rboost (crossfit) - Varying tau_heritability-nreps_crossfit

**Function**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Subgroup ATE Errors Plot', 'Varying nrep crossfits - AND DGP-Distilled Rboost (crossfit) - Varying tau_heritability-nreps_crossfit', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Subgroup ATE Errors Plot', 'Varying nrep crossfits - AND DGP-Distilled Rboost (crossfit) - Varying tau_heritability-nreps_crossfit', what = 'parameters')
```

##### Varying nrep crossfits - OR DGP-Distilled Rboost (crossfit) - Varying tau_heritability-nreps_crossfit

**Function**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Subgroup ATE Errors Plot', 'Varying nrep crossfits - OR DGP-Distilled Rboost (crossfit) - Varying tau_heritability-nreps_crossfit', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Subgroup ATE Errors Plot', 'Varying nrep crossfits - OR DGP-Distilled Rboost (crossfit) - Varying tau_heritability-nreps_crossfit', what = 'parameters')
```

</div>

<div class='panel panel-default padded-panel'>

### Subgroup ATE Errors Plot (max depth = 2) {.tabset .tabset-pills .tabset-circle .tabset-recipe}

#### `r fontawesome::fa('readme', fill = 'white')` {.tabset .tabset-pills}

We plot each method's subgroup average treatment effect (ATE) estimation error, measured as the root mean squared error between the true subgroup ATEs and the estimated subgroup ATEs. For tree-based methods, the trees have been pruned to a maximum depth of 2.

#### `r fontawesome::fa('code', fill = 'white')` {.tabset .tabset-pills}

##### Main Simulations - Additive DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Subgroup ATE Errors Plot (max depth = 2)', 'Main Simulations - Additive DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Subgroup ATE Errors Plot (max depth = 2)', 'Main Simulations - Additive DGP - Varying tau_heritability', what = 'parameters')
```

##### Main Simulations - Additive with Linear Covariates DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Subgroup ATE Errors Plot (max depth = 2)', 'Main Simulations - Additive with Linear Covariates DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Subgroup ATE Errors Plot (max depth = 2)', 'Main Simulations - Additive with Linear Covariates DGP - Varying tau_heritability', what = 'parameters')
```

##### Main Simulations - AND DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Subgroup ATE Errors Plot (max depth = 2)', 'Main Simulations - AND DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Subgroup ATE Errors Plot (max depth = 2)', 'Main Simulations - AND DGP - Varying tau_heritability', what = 'parameters')
```

##### Main Simulations - AND with Linear Covariates DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Subgroup ATE Errors Plot (max depth = 2)', 'Main Simulations - AND with Linear Covariates DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Subgroup ATE Errors Plot (max depth = 2)', 'Main Simulations - AND with Linear Covariates DGP - Varying tau_heritability', what = 'parameters')
```

##### Main Simulations - OR DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Subgroup ATE Errors Plot (max depth = 2)', 'Main Simulations - OR DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Subgroup ATE Errors Plot (max depth = 2)', 'Main Simulations - OR DGP - Varying tau_heritability', what = 'parameters')
```

##### Main Simulations - OR with Linear Covariates DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Subgroup ATE Errors Plot (max depth = 2)', 'Main Simulations - OR with Linear Covariates DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Subgroup ATE Errors Plot (max depth = 2)', 'Main Simulations - OR with Linear Covariates DGP - Varying tau_heritability', what = 'parameters')
```

</div>

<div class='panel panel-default padded-panel'>

### Feature and Split Frequency Plots {.tabset .tabset-pills .tabset-circle .tabset-recipe}

#### `r fontawesome::fa('readme', fill = 'white')` {.tabset .tabset-pills}

We plot (A) the number of simulation replicates (out of 100), for which each variable was selected (at least once) in the estimated subgroups and (B) the average number of tree splits that used each feature per tree.

#### `r fontawesome::fa('code', fill = 'white')` {.tabset .tabset-pills}

##### Main Simulations - Additive DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Feature and Split Frequency Plots', 'Main Simulations - Additive DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Feature and Split Frequency Plots', 'Main Simulations - Additive DGP - Varying tau_heritability', what = 'parameters')
```

##### Main Simulations - Additive with Linear Covariates DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Feature and Split Frequency Plots', 'Main Simulations - Additive with Linear Covariates DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Feature and Split Frequency Plots', 'Main Simulations - Additive with Linear Covariates DGP - Varying tau_heritability', what = 'parameters')
```

##### Main Simulations - AND DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Feature and Split Frequency Plots', 'Main Simulations - AND DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Feature and Split Frequency Plots', 'Main Simulations - AND DGP - Varying tau_heritability', what = 'parameters')
```

##### Main Simulations - AND with Linear Covariates DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Feature and Split Frequency Plots', 'Main Simulations - AND with Linear Covariates DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Feature and Split Frequency Plots', 'Main Simulations - AND with Linear Covariates DGP - Varying tau_heritability', what = 'parameters')
```

##### Main Simulations - OR with Linear Covariates DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Feature and Split Frequency Plots', 'Main Simulations - OR with Linear Covariates DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Feature and Split Frequency Plots', 'Main Simulations - OR with Linear Covariates DGP - Varying tau_heritability', what = 'parameters')
```

##### Rulefit Simulations - Additive DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Feature and Split Frequency Plots', 'Rulefit Simulations - Additive DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Feature and Split Frequency Plots', 'Rulefit Simulations - Additive DGP - Varying tau_heritability', what = 'parameters')
```

##### Rulefit Simulations - Additive with Linear Covariates DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Feature and Split Frequency Plots', 'Rulefit Simulations - Additive with Linear Covariates DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Feature and Split Frequency Plots', 'Rulefit Simulations - Additive with Linear Covariates DGP - Varying tau_heritability', what = 'parameters')
```

##### Rulefit Simulations - AND DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Feature and Split Frequency Plots', 'Rulefit Simulations - AND DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Feature and Split Frequency Plots', 'Rulefit Simulations - AND DGP - Varying tau_heritability', what = 'parameters')
```

##### Rulefit Simulations - AND with Linear Covariates DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Feature and Split Frequency Plots', 'Rulefit Simulations - AND with Linear Covariates DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Feature and Split Frequency Plots', 'Rulefit Simulations - AND with Linear Covariates DGP - Varying tau_heritability', what = 'parameters')
```

##### Rulefit Simulations - OR DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Feature and Split Frequency Plots', 'Rulefit Simulations - OR DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Feature and Split Frequency Plots', 'Rulefit Simulations - OR DGP - Varying tau_heritability', what = 'parameters')
```

##### Rulefit Simulations - OR with Linear Covariates DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Feature and Split Frequency Plots', 'Rulefit Simulations - OR with Linear Covariates DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Feature and Split Frequency Plots', 'Rulefit Simulations - OR with Linear Covariates DGP - Varying tau_heritability', what = 'parameters')
```

</div>

<div class='panel panel-default padded-panel'>

### Feature and Split Frequency Plots (max depth = 2) {.tabset .tabset-pills .tabset-circle .tabset-recipe}

#### `r fontawesome::fa('readme', fill = 'white')` {.tabset .tabset-pills}

We plot (A) the number of simulation replicates (out of 100), for which each variable was selected (at least once) in the estimated subgroups and (B) the average number of tree splits that used each feature per tree. Only splits made in the first two depths of the tree are considered.

#### `r fontawesome::fa('code', fill = 'white')` {.tabset .tabset-pills}

##### Main Simulations - Additive DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Feature and Split Frequency Plots (max depth = 2)', 'Main Simulations - Additive DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Feature and Split Frequency Plots (max depth = 2)', 'Main Simulations - Additive DGP - Varying tau_heritability', what = 'parameters')
```

##### Main Simulations - Additive with Linear Covariates DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Feature and Split Frequency Plots (max depth = 2)', 'Main Simulations - Additive with Linear Covariates DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Feature and Split Frequency Plots (max depth = 2)', 'Main Simulations - Additive with Linear Covariates DGP - Varying tau_heritability', what = 'parameters')
```

##### Main Simulations - AND DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Feature and Split Frequency Plots (max depth = 2)', 'Main Simulations - AND DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Feature and Split Frequency Plots (max depth = 2)', 'Main Simulations - AND DGP - Varying tau_heritability', what = 'parameters')
```

##### Main Simulations - AND with Linear Covariates DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Feature and Split Frequency Plots (max depth = 2)', 'Main Simulations - AND with Linear Covariates DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Feature and Split Frequency Plots (max depth = 2)', 'Main Simulations - AND with Linear Covariates DGP - Varying tau_heritability', what = 'parameters')
```

##### Main Simulations - OR with Linear Covariates DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Feature and Split Frequency Plots (max depth = 2)', 'Main Simulations - OR with Linear Covariates DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Feature and Split Frequency Plots (max depth = 2)', 'Main Simulations - OR with Linear Covariates DGP - Varying tau_heritability', what = 'parameters')
```

</div>

<div class='panel panel-default padded-panel'>

### Subgroup Stability Diagnostics Plot {.tabset .tabset-pills .tabset-circle .tabset-recipe}

#### `r fontawesome::fa('readme', fill = 'white')` {.tabset .tabset-pills}

[Huang et al. (2025)]() develop a stability-driven model selection procedure to choose an appropriate teacher model in the CDT framework. This teacher model selection procedure leverages the Jaccard subgroup stability index (SSI) to identify the most stable teacher model. We thus perform this teacher model selection procedure and plot the Jaccard SSI for a range of tree depths (1-4). We also plot these teacher model selection results alongside the distribution of features selected at each tree depth across 100 bootstrap samples. Higher Jaccard SSI indicates a more stable teacher model.

#### `r fontawesome::fa('code', fill = 'white')` {.tabset .tabset-pills}

##### Main Simulations - Additive DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Subgroup Stability Diagnostics Plot', 'Main Simulations - Additive DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Subgroup Stability Diagnostics Plot', 'Main Simulations - Additive DGP - Varying tau_heritability', what = 'parameters')
```

##### Main Simulations - Additive with Linear Covariates DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Subgroup Stability Diagnostics Plot', 'Main Simulations - Additive with Linear Covariates DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Subgroup Stability Diagnostics Plot', 'Main Simulations - Additive with Linear Covariates DGP - Varying tau_heritability', what = 'parameters')
```

##### Main Simulations - AND DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Subgroup Stability Diagnostics Plot', 'Main Simulations - AND DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Subgroup Stability Diagnostics Plot', 'Main Simulations - AND DGP - Varying tau_heritability', what = 'parameters')
```

##### Main Simulations - AND with Linear Covariates DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Subgroup Stability Diagnostics Plot', 'Main Simulations - AND with Linear Covariates DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Subgroup Stability Diagnostics Plot', 'Main Simulations - AND with Linear Covariates DGP - Varying tau_heritability', what = 'parameters')
```

##### Main Simulations - OR DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Subgroup Stability Diagnostics Plot', 'Main Simulations - OR DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Subgroup Stability Diagnostics Plot', 'Main Simulations - OR DGP - Varying tau_heritability', what = 'parameters')
```

##### Main Simulations - OR with Linear Covariates DGP - Varying tau_heritability

**Function**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Subgroup Stability Diagnostics Plot', 'Main Simulations - OR with Linear Covariates DGP - Varying tau_heritability', what = 'function')
```

**Input Parameters**

```{r, max.height='200px'}
show_recipe(viz_objs, 'Subgroup Stability Diagnostics Plot', 'Main Simulations - OR with Linear Covariates DGP - Varying tau_heritability', what = 'parameters')
```

</div>

## R Environment {.tabset .tabset-pills .tabset-square}

### R/Packages

#### R {.tabset .tabset-pills .tabset-circle .tabset-recipe}

<div class='panel panel-default padded-panel'>

##### `r fontawesome::fa('table', fill = 'white')`

```{r, results = 'asis'}
print_renv(renv::lockfile_read(project = here::here()), 'R', print_table = TRUE)
```

##### `r fontawesome::fa('code', fill = 'white')`

```{r, max.height='700px'}
print_renv(renv::lockfile_read(project = here::here()), 'R', print_table = FALSE)
```

</div>

#### Packages {.tabset .tabset-pills .tabset-circle .tabset-recipe}

<div class='panel panel-default padded-panel'>

##### `r fontawesome::fa('table', fill = 'white')`

```{r, results = 'asis'}
print_renv(renv::lockfile_read(project = here::here()), 'Packages', print_table = TRUE)
```

##### `r fontawesome::fa('code', fill = 'white')`

```{r, max.height='700px'}
print_renv(renv::lockfile_read(project = here::here()), 'Packages', print_table = FALSE)
```

</div>

### Full renv Diagnostics

<div class='panel panel-default padded-panel'>

```{r, max.height='700px'}
renv::diagnostics()
```

</div>

# Main Simulations {.tabset .tabset-pills .tabset-vmodern}

```{r, results = 'asis'}
results <- get_exp_results(
  '/Users/ttang4/Documents/Research/Causal Subgroups/causalDT/causalDT-manuscript/results/Main Simulations', show_eval = FALSE, show_viz = TRUE, eval_cache = '.rds', viz_cache = '.png'
)
```

## Additive DGP {.tabset .tabset-pills}

```{r, results = 'asis'}
results <- get_exp_results(
  '/Users/ttang4/Documents/Research/Causal Subgroups/causalDT/causalDT-manuscript/results/Main Simulations/Additive DGP', show_eval = FALSE, show_viz = TRUE, eval_cache = '.rds', viz_cache = '.png'
)
```

### Varying n {.tabset .tabset-pills .tabset-circle}

```{r, results = 'asis'}
results <- get_exp_results(
  '/Users/ttang4/Documents/Research/Causal Subgroups/causalDT/causalDT-manuscript/results/Main Simulations/Additive DGP/Varying n', show_eval = FALSE, show_viz = TRUE, eval_cache = '.rds', viz_cache = '.png'
)
```

#### `r fontawesome::fa('chart-bar', fill = 'white')`

##### Number of Subgroups Plot

###### {.tabset .tabset-pills}

```{r, fig.height = 5, fig.width = 12, out.width = '100%', add.panel = TRUE}
show_results(results, 'Number of Subgroups Plot', 'visualizer')
```

#### `r fontawesome::fa('code', fill = 'white')`

**Parameter Values**

```{r, max.height='200px'}
show_results(results, NULL, 'vary_params')
```

### Varying tau_heritability {.tabset .tabset-pills .tabset-circle}

```{r, results = 'asis'}
results <- get_exp_results(
  '/Users/ttang4/Documents/Research/Causal Subgroups/causalDT/causalDT-manuscript/results/Main Simulations/Additive DGP/Varying tau_heritability', show_eval = FALSE, show_viz = TRUE, eval_cache = '.rds', viz_cache = '.png'
)
```

#### `r fontawesome::fa('chart-bar', fill = 'white')`

##### Subgroup Feature Selection Errors Plot

###### {.tabset .tabset-pills}

```{r, fig.height = 5, fig.width = 12, out.width = '100%', add.panel = TRUE}
show_results(results, 'Subgroup Feature Selection Errors Plot', 'visualizer')
```

##### Subgroup Thresholds Distribution Plot

###### {.tabset .tabset-pills}

```{r, fig.height = 5, fig.width = 12, out.width = '100%', add.panel = TRUE}
show_results(results, 'Subgroup Thresholds Distribution Plot', 'visualizer')
```

##### Threshold Distances Plot

###### {.tabset .tabset-pills}

```{r, fig.height = 4.5, fig.width = 9, out.width = '100%', add.panel = TRUE}
show_results(results, 'Threshold Distances Plot', 'visualizer')
```

##### Subgroup ATE Errors Plot

###### {.tabset .tabset-pills}

```{r, fig.height = 4.5, fig.width = 8, out.width = '100%', add.panel = TRUE}
show_results(results, 'Subgroup ATE Errors Plot', 'visualizer')
```

##### Feature and Split Frequency Plots

###### {.tabset .tabset-pills}

```{r, fig.height = 7.5, fig.width = 18, out.width = '100%', add.panel = TRUE}
show_results(results, 'Feature and Split Frequency Plots', 'visualizer')
```

##### Subgroup Stability Diagnostics Plot

###### {.tabset .tabset-pills}

```{r, fig.height = 7, fig.width = 11, out.width = '100%', add.panel = TRUE}
show_results(results, 'Subgroup Stability Diagnostics Plot', 'visualizer')
```

##### Subgroup Feature Selection Errors Plot (max depth = 2)

###### {.tabset .tabset-pills}

```{r, fig.height = 5, fig.width = 12, out.width = '100%', add.panel = TRUE}
show_results(results, 'Subgroup Feature Selection Errors Plot (max depth = 2)', 'visualizer')
```

##### Subgroup Thresholds Distribution Plot (max depth = 2)

###### {.tabset .tabset-pills}

```{r, fig.height = 5, fig.width = 12, out.width = '100%', add.panel = TRUE}
show_results(results, 'Subgroup Thresholds Distribution Plot (max depth = 2)', 'visualizer')
```

##### Threshold Distances Plot (max depth = 2)

###### {.tabset .tabset-pills}

```{r, fig.height = 4.5, fig.width = 9, out.width = '100%', add.panel = TRUE}
show_results(results, 'Threshold Distances Plot (max depth = 2)', 'visualizer')
```

##### Subgroup ATE Errors Plot (max depth = 2)

###### {.tabset .tabset-pills}

```{r, fig.height = 4.5, fig.width = 8, out.width = '100%', add.panel = TRUE}
show_results(results, 'Subgroup ATE Errors Plot (max depth = 2)', 'visualizer')
```

##### Feature and Split Frequency Plots (max depth = 2)

###### {.tabset .tabset-pills}

```{r, fig.height = 7.5, fig.width = 18, out.width = '100%', add.panel = TRUE}
show_results(results, 'Feature and Split Frequency Plots (max depth = 2)', 'visualizer')
```

#### `r fontawesome::fa('code', fill = 'white')`

**Parameter Values**

```{r, max.height='200px'}
show_results(results, NULL, 'vary_params')
```

## Additive with Linear Covariates DGP {.tabset .tabset-pills}

```{r, results = 'asis'}
results <- get_exp_results(
  '/Users/ttang4/Documents/Research/Causal Subgroups/causalDT/causalDT-manuscript/results/Main Simulations/Additive with Linear Covariates DGP', show_eval = FALSE, show_viz = TRUE, eval_cache = '.rds', viz_cache = '.png'
)
```

### Varying n {.tabset .tabset-pills .tabset-circle}

```{r, results = 'asis'}
results <- get_exp_results(
  '/Users/ttang4/Documents/Research/Causal Subgroups/causalDT/causalDT-manuscript/results/Main Simulations/Additive with Linear Covariates DGP/Varying n', show_eval = FALSE, show_viz = TRUE, eval_cache = '.rds', viz_cache = '.png'
)
```

#### `r fontawesome::fa('chart-bar', fill = 'white')`

##### Number of Subgroups Plot

###### {.tabset .tabset-pills}

```{r, fig.height = 5, fig.width = 12, out.width = '100%', add.panel = TRUE}
show_results(results, 'Number of Subgroups Plot', 'visualizer')
```

#### `r fontawesome::fa('code', fill = 'white')`

**Parameter Values**

```{r, max.height='200px'}
show_results(results, NULL, 'vary_params')
```

### Varying tau_heritability {.tabset .tabset-pills .tabset-circle}

```{r, results = 'asis'}
results <- get_exp_results(
  '/Users/ttang4/Documents/Research/Causal Subgroups/causalDT/causalDT-manuscript/results/Main Simulations/Additive with Linear Covariates DGP/Varying tau_heritability', show_eval = FALSE, show_viz = TRUE, eval_cache = '.rds', viz_cache = '.png'
)
```

#### `r fontawesome::fa('chart-bar', fill = 'white')`

##### Subgroup Feature Selection Errors Plot

###### {.tabset .tabset-pills}

```{r, fig.height = 5, fig.width = 12, out.width = '100%', add.panel = TRUE}
show_results(results, 'Subgroup Feature Selection Errors Plot', 'visualizer')
```

##### Subgroup Thresholds Distribution Plot

###### {.tabset .tabset-pills}

```{r, fig.height = 5, fig.width = 12, out.width = '100%', add.panel = TRUE}
show_results(results, 'Subgroup Thresholds Distribution Plot', 'visualizer')
```

##### Threshold Distances Plot

###### {.tabset .tabset-pills}

```{r, fig.height = 4.5, fig.width = 9, out.width = '100%', add.panel = TRUE}
show_results(results, 'Threshold Distances Plot', 'visualizer')
```

##### Subgroup ATE Errors Plot

###### {.tabset .tabset-pills}

```{r, fig.height = 4.5, fig.width = 8, out.width = '100%', add.panel = TRUE}
show_results(results, 'Subgroup ATE Errors Plot', 'visualizer')
```

##### Feature and Split Frequency Plots

###### {.tabset .tabset-pills}

```{r, fig.height = 7.5, fig.width = 18, out.width = '100%', add.panel = TRUE}
show_results(results, 'Feature and Split Frequency Plots', 'visualizer')
```

##### Subgroup Stability Diagnostics Plot

###### {.tabset .tabset-pills}

```{r, fig.height = 7, fig.width = 11, out.width = '100%', add.panel = TRUE}
show_results(results, 'Subgroup Stability Diagnostics Plot', 'visualizer')
```

##### Subgroup Feature Selection Errors Plot (max depth = 2)

###### {.tabset .tabset-pills}

```{r, fig.height = 5, fig.width = 12, out.width = '100%', add.panel = TRUE}
show_results(results, 'Subgroup Feature Selection Errors Plot (max depth = 2)', 'visualizer')
```

##### Subgroup Thresholds Distribution Plot (max depth = 2)

###### {.tabset .tabset-pills}

```{r, fig.height = 5, fig.width = 12, out.width = '100%', add.panel = TRUE}
show_results(results, 'Subgroup Thresholds Distribution Plot (max depth = 2)', 'visualizer')
```

##### Threshold Distances Plot (max depth = 2)

###### {.tabset .tabset-pills}

```{r, fig.height = 4.5, fig.width = 9, out.width = '100%', add.panel = TRUE}
show_results(results, 'Threshold Distances Plot (max depth = 2)', 'visualizer')
```

##### Subgroup ATE Errors Plot (max depth = 2)

###### {.tabset .tabset-pills}

```{r, fig.height = 4.5, fig.width = 8, out.width = '100%', add.panel = TRUE}
show_results(results, 'Subgroup ATE Errors Plot (max depth = 2)', 'visualizer')
```

##### Feature and Split Frequency Plots (max depth = 2)

###### {.tabset .tabset-pills}

```{r, fig.height = 7.5, fig.width = 18, out.width = '100%', add.panel = TRUE}
show_results(results, 'Feature and Split Frequency Plots (max depth = 2)', 'visualizer')
```

#### `r fontawesome::fa('code', fill = 'white')`

**Parameter Values**

```{r, max.height='200px'}
show_results(results, NULL, 'vary_params')
```

## AND DGP {.tabset .tabset-pills}

```{r, results = 'asis'}
results <- get_exp_results(
  '/Users/ttang4/Documents/Research/Causal Subgroups/causalDT/causalDT-manuscript/results/Main Simulations/AND DGP', show_eval = FALSE, show_viz = TRUE, eval_cache = '.rds', viz_cache = '.png'
)
```

### Varying n {.tabset .tabset-pills .tabset-circle}

```{r, results = 'asis'}
results <- get_exp_results(
  '/Users/ttang4/Documents/Research/Causal Subgroups/causalDT/causalDT-manuscript/results/Main Simulations/AND DGP/Varying n', show_eval = FALSE, show_viz = TRUE, eval_cache = '.rds', viz_cache = '.png'
)
```

#### `r fontawesome::fa('chart-bar', fill = 'white')`

##### Number of Subgroups Plot

###### {.tabset .tabset-pills}

```{r, fig.height = 5, fig.width = 12, out.width = '100%', add.panel = TRUE}
show_results(results, 'Number of Subgroups Plot', 'visualizer')
```

#### `r fontawesome::fa('code', fill = 'white')`

**Parameter Values**

```{r, max.height='200px'}
show_results(results, NULL, 'vary_params')
```

### Varying tau_heritability {.tabset .tabset-pills .tabset-circle}

```{r, results = 'asis'}
results <- get_exp_results(
  '/Users/ttang4/Documents/Research/Causal Subgroups/causalDT/causalDT-manuscript/results/Main Simulations/AND DGP/Varying tau_heritability', show_eval = FALSE, show_viz = TRUE, eval_cache = '.rds', viz_cache = '.png'
)
```

#### `r fontawesome::fa('chart-bar', fill = 'white')`

##### Subgroup Feature Selection Errors Plot

###### {.tabset .tabset-pills}

```{r, fig.height = 5, fig.width = 12, out.width = '100%', add.panel = TRUE}
show_results(results, 'Subgroup Feature Selection Errors Plot', 'visualizer')
```

##### Subgroup Thresholds Distribution Plot

###### {.tabset .tabset-pills}

```{r, fig.height = 5, fig.width = 12, out.width = '100%', add.panel = TRUE}
show_results(results, 'Subgroup Thresholds Distribution Plot', 'visualizer')
```

##### Threshold Distances Plot

###### {.tabset .tabset-pills}

```{r, fig.height = 4.5, fig.width = 9, out.width = '100%', add.panel = TRUE}
show_results(results, 'Threshold Distances Plot', 'visualizer')
```

##### Subgroup ATE Errors Plot

###### {.tabset .tabset-pills}

```{r, fig.height = 4.5, fig.width = 8, out.width = '100%', add.panel = TRUE}
show_results(results, 'Subgroup ATE Errors Plot', 'visualizer')
```

##### Feature and Split Frequency Plots

###### {.tabset .tabset-pills}

```{r, fig.height = 7.5, fig.width = 18, out.width = '100%', add.panel = TRUE}
show_results(results, 'Feature and Split Frequency Plots', 'visualizer')
```

##### Subgroup Stability Diagnostics Plot

###### {.tabset .tabset-pills}

```{r, fig.height = 7, fig.width = 11, out.width = '100%', add.panel = TRUE}
show_results(results, 'Subgroup Stability Diagnostics Plot', 'visualizer')
```

##### Subgroup Feature Selection Errors Plot (max depth = 2)

###### {.tabset .tabset-pills}

```{r, fig.height = 5, fig.width = 12, out.width = '100%', add.panel = TRUE}
show_results(results, 'Subgroup Feature Selection Errors Plot (max depth = 2)', 'visualizer')
```

##### Subgroup Thresholds Distribution Plot (max depth = 2)

###### {.tabset .tabset-pills}

```{r, fig.height = 5, fig.width = 12, out.width = '100%', add.panel = TRUE}
show_results(results, 'Subgroup Thresholds Distribution Plot (max depth = 2)', 'visualizer')
```

##### Threshold Distances Plot (max depth = 2)

###### {.tabset .tabset-pills}

```{r, fig.height = 4.5, fig.width = 9, out.width = '100%', add.panel = TRUE}
show_results(results, 'Threshold Distances Plot (max depth = 2)', 'visualizer')
```

##### Subgroup ATE Errors Plot (max depth = 2)

###### {.tabset .tabset-pills}

```{r, fig.height = 4.5, fig.width = 8, out.width = '100%', add.panel = TRUE}
show_results(results, 'Subgroup ATE Errors Plot (max depth = 2)', 'visualizer')
```

##### Feature and Split Frequency Plots (max depth = 2)

###### {.tabset .tabset-pills}

```{r, fig.height = 7.5, fig.width = 18, out.width = '100%', add.panel = TRUE}
show_results(results, 'Feature and Split Frequency Plots (max depth = 2)', 'visualizer')
```

#### `r fontawesome::fa('code', fill = 'white')`

**Parameter Values**

```{r, max.height='200px'}
show_results(results, NULL, 'vary_params')
```

## AND with Linear Covariates DGP {.tabset .tabset-pills}

```{r, results = 'asis'}
results <- get_exp_results(
  '/Users/ttang4/Documents/Research/Causal Subgroups/causalDT/causalDT-manuscript/results/Main Simulations/AND with Linear Covariates DGP', show_eval = FALSE, show_viz = TRUE, eval_cache = '.rds', viz_cache = '.png'
)
```

### Varying n {.tabset .tabset-pills .tabset-circle}

```{r, results = 'asis'}
results <- get_exp_results(
  '/Users/ttang4/Documents/Research/Causal Subgroups/causalDT/causalDT-manuscript/results/Main Simulations/AND with Linear Covariates DGP/Varying n', show_eval = FALSE, show_viz = TRUE, eval_cache = '.rds', viz_cache = '.png'
)
```

#### `r fontawesome::fa('chart-bar', fill = 'white')`

##### Number of Subgroups Plot

###### {.tabset .tabset-pills}

```{r, fig.height = 5, fig.width = 12, out.width = '100%', add.panel = TRUE}
show_results(results, 'Number of Subgroups Plot', 'visualizer')
```

#### `r fontawesome::fa('code', fill = 'white')`

**Parameter Values**

```{r, max.height='200px'}
show_results(results, NULL, 'vary_params')
```

### Varying tau_heritability {.tabset .tabset-pills .tabset-circle}

```{r, results = 'asis'}
results <- get_exp_results(
  '/Users/ttang4/Documents/Research/Causal Subgroups/causalDT/causalDT-manuscript/results/Main Simulations/AND with Linear Covariates DGP/Varying tau_heritability', show_eval = FALSE, show_viz = TRUE, eval_cache = '.rds', viz_cache = '.png'
)
```

#### `r fontawesome::fa('chart-bar', fill = 'white')`

##### Subgroup Feature Selection Errors Plot

###### {.tabset .tabset-pills}

```{r, fig.height = 5, fig.width = 12, out.width = '100%', add.panel = TRUE}
show_results(results, 'Subgroup Feature Selection Errors Plot', 'visualizer')
```

##### Subgroup Thresholds Distribution Plot

###### {.tabset .tabset-pills}

```{r, fig.height = 5, fig.width = 12, out.width = '100%', add.panel = TRUE}
show_results(results, 'Subgroup Thresholds Distribution Plot', 'visualizer')
```

##### Threshold Distances Plot

###### {.tabset .tabset-pills}

```{r, fig.height = 4.5, fig.width = 9, out.width = '100%', add.panel = TRUE}
show_results(results, 'Threshold Distances Plot', 'visualizer')
```

##### Subgroup ATE Errors Plot

###### {.tabset .tabset-pills}

```{r, fig.height = 4.5, fig.width = 8, out.width = '100%', add.panel = TRUE}
show_results(results, 'Subgroup ATE Errors Plot', 'visualizer')
```

##### Feature and Split Frequency Plots

###### {.tabset .tabset-pills}

```{r, fig.height = 7.5, fig.width = 18, out.width = '100%', add.panel = TRUE}
show_results(results, 'Feature and Split Frequency Plots', 'visualizer')
```

##### Subgroup Stability Diagnostics Plot

###### {.tabset .tabset-pills}

```{r, fig.height = 7, fig.width = 11, out.width = '100%', add.panel = TRUE}
show_results(results, 'Subgroup Stability Diagnostics Plot', 'visualizer')
```

##### Subgroup Feature Selection Errors Plot (max depth = 2)

###### {.tabset .tabset-pills}

```{r, fig.height = 5, fig.width = 12, out.width = '100%', add.panel = TRUE}
show_results(results, 'Subgroup Feature Selection Errors Plot (max depth = 2)', 'visualizer')
```

##### Subgroup Thresholds Distribution Plot (max depth = 2)

###### {.tabset .tabset-pills}

```{r, fig.height = 5, fig.width = 12, out.width = '100%', add.panel = TRUE}
show_results(results, 'Subgroup Thresholds Distribution Plot (max depth = 2)', 'visualizer')
```

##### Threshold Distances Plot (max depth = 2)

###### {.tabset .tabset-pills}

```{r, fig.height = 4.5, fig.width = 9, out.width = '100%', add.panel = TRUE}
show_results(results, 'Threshold Distances Plot (max depth = 2)', 'visualizer')
```

##### Subgroup ATE Errors Plot (max depth = 2)

###### {.tabset .tabset-pills}

```{r, fig.height = 4.5, fig.width = 8, out.width = '100%', add.panel = TRUE}
show_results(results, 'Subgroup ATE Errors Plot (max depth = 2)', 'visualizer')
```

##### Feature and Split Frequency Plots (max depth = 2)

###### {.tabset .tabset-pills}

```{r, fig.height = 7.5, fig.width = 18, out.width = '100%', add.panel = TRUE}
show_results(results, 'Feature and Split Frequency Plots (max depth = 2)', 'visualizer')
```

#### `r fontawesome::fa('code', fill = 'white')`

**Parameter Values**

```{r, max.height='200px'}
show_results(results, NULL, 'vary_params')
```

## OR DGP {.tabset .tabset-pills}

```{r, results = 'asis'}
results <- get_exp_results(
  '/Users/ttang4/Documents/Research/Causal Subgroups/causalDT/causalDT-manuscript/results/Main Simulations/OR DGP', show_eval = FALSE, show_viz = TRUE, eval_cache = '.rds', viz_cache = '.png'
)
```

### Varying n {.tabset .tabset-pills .tabset-circle}

```{r, results = 'asis'}
results <- get_exp_results(
  '/Users/ttang4/Documents/Research/Causal Subgroups/causalDT/causalDT-manuscript/results/Main Simulations/OR DGP/Varying n', show_eval = FALSE, show_viz = TRUE, eval_cache = '.rds', viz_cache = '.png'
)
```

#### `r fontawesome::fa('chart-bar', fill = 'white')`

##### Number of Subgroups Plot

###### {.tabset .tabset-pills}

```{r, fig.height = 5, fig.width = 12, out.width = '100%', add.panel = TRUE}
show_results(results, 'Number of Subgroups Plot', 'visualizer')
```

#### `r fontawesome::fa('code', fill = 'white')`

**Parameter Values**

```{r, max.height='200px'}
show_results(results, NULL, 'vary_params')
```

### Varying tau_heritability {.tabset .tabset-pills .tabset-circle}

```{r, results = 'asis'}
results <- get_exp_results(
  '/Users/ttang4/Documents/Research/Causal Subgroups/causalDT/causalDT-manuscript/results/Main Simulations/OR DGP/Varying tau_heritability', show_eval = FALSE, show_viz = TRUE, eval_cache = '.rds', viz_cache = '.png'
)
```

#### `r fontawesome::fa('chart-bar', fill = 'white')`

##### Subgroup Feature Selection Errors Plot

###### {.tabset .tabset-pills}

```{r, fig.height = 5, fig.width = 12, out.width = '100%', add.panel = TRUE}
show_results(results, 'Subgroup Feature Selection Errors Plot', 'visualizer')
```

##### Subgroup Thresholds Distribution Plot

###### {.tabset .tabset-pills}

```{r, fig.height = 5, fig.width = 12, out.width = '100%', add.panel = TRUE}
show_results(results, 'Subgroup Thresholds Distribution Plot', 'visualizer')
```

##### Threshold Distances Plot

###### {.tabset .tabset-pills}

```{r, fig.height = 4.5, fig.width = 9, out.width = '100%', add.panel = TRUE}
show_results(results, 'Threshold Distances Plot', 'visualizer')
```

##### Subgroup ATE Errors Plot

###### {.tabset .tabset-pills}

```{r, fig.height = 4.5, fig.width = 8, out.width = '100%', add.panel = TRUE}
show_results(results, 'Subgroup ATE Errors Plot', 'visualizer')
```

##### Subgroup Stability Diagnostics Plot

###### {.tabset .tabset-pills}

```{r, fig.height = 7, fig.width = 11, out.width = '100%', add.panel = TRUE}
show_results(results, 'Subgroup Stability Diagnostics Plot', 'visualizer')
```

##### Subgroup Feature Selection Errors Plot (max depth = 2)

###### {.tabset .tabset-pills}

```{r, fig.height = 5, fig.width = 12, out.width = '100%', add.panel = TRUE}
show_results(results, 'Subgroup Feature Selection Errors Plot (max depth = 2)', 'visualizer')
```

##### Subgroup Thresholds Distribution Plot (max depth = 2)

###### {.tabset .tabset-pills}

```{r, fig.height = 5, fig.width = 12, out.width = '100%', add.panel = TRUE}
show_results(results, 'Subgroup Thresholds Distribution Plot (max depth = 2)', 'visualizer')
```

##### Threshold Distances Plot (max depth = 2)

###### {.tabset .tabset-pills}

```{r, fig.height = 4.5, fig.width = 9, out.width = '100%', add.panel = TRUE}
show_results(results, 'Threshold Distances Plot (max depth = 2)', 'visualizer')
```

##### Subgroup ATE Errors Plot (max depth = 2)

###### {.tabset .tabset-pills}

```{r, fig.height = 4.5, fig.width = 8, out.width = '100%', add.panel = TRUE}
show_results(results, 'Subgroup ATE Errors Plot (max depth = 2)', 'visualizer')
```

#### `r fontawesome::fa('code', fill = 'white')`

**Parameter Values**

```{r, max.height='200px'}
show_results(results, NULL, 'vary_params')
```

## OR with Linear Covariates DGP {.tabset .tabset-pills}

```{r, results = 'asis'}
results <- get_exp_results(
  '/Users/ttang4/Documents/Research/Causal Subgroups/causalDT/causalDT-manuscript/results/Main Simulations/OR with Linear Covariates DGP', show_eval = FALSE, show_viz = TRUE, eval_cache = '.rds', viz_cache = '.png'
)
```

### Varying n {.tabset .tabset-pills .tabset-circle}

```{r, results = 'asis'}
results <- get_exp_results(
  '/Users/ttang4/Documents/Research/Causal Subgroups/causalDT/causalDT-manuscript/results/Main Simulations/OR with Linear Covariates DGP/Varying n', show_eval = FALSE, show_viz = TRUE, eval_cache = '.rds', viz_cache = '.png'
)
```

#### `r fontawesome::fa('chart-bar', fill = 'white')`

##### Number of Subgroups Plot

###### {.tabset .tabset-pills}

```{r, fig.height = 5, fig.width = 12, out.width = '100%', add.panel = TRUE}
show_results(results, 'Number of Subgroups Plot', 'visualizer')
```

#### `r fontawesome::fa('code', fill = 'white')`

**Parameter Values**

```{r, max.height='200px'}
show_results(results, NULL, 'vary_params')
```

### Varying tau_heritability {.tabset .tabset-pills .tabset-circle}

```{r, results = 'asis'}
results <- get_exp_results(
  '/Users/ttang4/Documents/Research/Causal Subgroups/causalDT/causalDT-manuscript/results/Main Simulations/OR with Linear Covariates DGP/Varying tau_heritability', show_eval = FALSE, show_viz = TRUE, eval_cache = '.rds', viz_cache = '.png'
)
```

#### `r fontawesome::fa('chart-bar', fill = 'white')`

##### Subgroup Feature Selection Errors Plot

###### {.tabset .tabset-pills}

```{r, fig.height = 5, fig.width = 12, out.width = '100%', add.panel = TRUE}
show_results(results, 'Subgroup Feature Selection Errors Plot', 'visualizer')
```

##### Subgroup Thresholds Distribution Plot

###### {.tabset .tabset-pills}

```{r, fig.height = 5, fig.width = 12, out.width = '100%', add.panel = TRUE}
show_results(results, 'Subgroup Thresholds Distribution Plot', 'visualizer')
```

##### Threshold Distances Plot

###### {.tabset .tabset-pills}

```{r, fig.height = 4.5, fig.width = 9, out.width = '100%', add.panel = TRUE}
show_results(results, 'Threshold Distances Plot', 'visualizer')
```

##### Subgroup ATE Errors Plot

###### {.tabset .tabset-pills}

```{r, fig.height = 4.5, fig.width = 8, out.width = '100%', add.panel = TRUE}
show_results(results, 'Subgroup ATE Errors Plot', 'visualizer')
```

##### Feature and Split Frequency Plots

###### {.tabset .tabset-pills}

```{r, fig.height = 7.5, fig.width = 18, out.width = '100%', add.panel = TRUE}
show_results(results, 'Feature and Split Frequency Plots', 'visualizer')
```

##### Subgroup Stability Diagnostics Plot

###### {.tabset .tabset-pills}

```{r, fig.height = 7, fig.width = 11, out.width = '100%', add.panel = TRUE}
show_results(results, 'Subgroup Stability Diagnostics Plot', 'visualizer')
```

##### Subgroup Feature Selection Errors Plot (max depth = 2)

###### {.tabset .tabset-pills}

```{r, fig.height = 5, fig.width = 12, out.width = '100%', add.panel = TRUE}
show_results(results, 'Subgroup Feature Selection Errors Plot (max depth = 2)', 'visualizer')
```

##### Subgroup Thresholds Distribution Plot (max depth = 2)

###### {.tabset .tabset-pills}

```{r, fig.height = 5, fig.width = 12, out.width = '100%', add.panel = TRUE}
show_results(results, 'Subgroup Thresholds Distribution Plot (max depth = 2)', 'visualizer')
```

##### Threshold Distances Plot (max depth = 2)

###### {.tabset .tabset-pills}

```{r, fig.height = 4.5, fig.width = 9, out.width = '100%', add.panel = TRUE}
show_results(results, 'Threshold Distances Plot (max depth = 2)', 'visualizer')
```

##### Subgroup ATE Errors Plot (max depth = 2)

###### {.tabset .tabset-pills}

```{r, fig.height = 4.5, fig.width = 8, out.width = '100%', add.panel = TRUE}
show_results(results, 'Subgroup ATE Errors Plot (max depth = 2)', 'visualizer')
```

##### Feature and Split Frequency Plots (max depth = 2)

###### {.tabset .tabset-pills}

```{r, fig.height = 7.5, fig.width = 18, out.width = '100%', add.panel = TRUE}
show_results(results, 'Feature and Split Frequency Plots (max depth = 2)', 'visualizer')
```

#### `r fontawesome::fa('code', fill = 'white')`

**Parameter Values**

```{r, max.height='200px'}
show_results(results, NULL, 'vary_params')
```

# Rulefit Simulations {.tabset .tabset-pills .tabset-vmodern}

```{r, results = 'asis'}
results <- get_exp_results(
  '/Users/ttang4/Documents/Research/Causal Subgroups/causalDT/causalDT-manuscript/results/Rulefit Simulations', show_eval = FALSE, show_viz = TRUE, eval_cache = '.rds', viz_cache = '.png'
)
```

## Additive DGP {.tabset .tabset-pills}

```{r, results = 'asis'}
results <- get_exp_results(
  '/Users/ttang4/Documents/Research/Causal Subgroups/causalDT/causalDT-manuscript/results/Rulefit Simulations/Additive DGP', show_eval = FALSE, show_viz = TRUE, eval_cache = '.rds', viz_cache = '.png'
)
```

### Varying tau_heritability {.tabset .tabset-pills .tabset-circle}

```{r, results = 'asis'}
results <- get_exp_results(
  '/Users/ttang4/Documents/Research/Causal Subgroups/causalDT/causalDT-manuscript/results/Rulefit Simulations/Additive DGP/Varying tau_heritability', show_eval = FALSE, show_viz = TRUE, eval_cache = '.rds', viz_cache = '.png'
)
```

#### `r fontawesome::fa('chart-bar', fill = 'white')`

##### Subgroup Feature Selection Errors Plot

###### {.tabset .tabset-pills}

```{r, fig.height = 5, fig.width = 12, out.width = '100%', add.panel = TRUE}
show_results(results, 'Subgroup Feature Selection Errors Plot', 'visualizer')
```

##### Subgroup Thresholds Distribution Plot

###### {.tabset .tabset-pills}

```{r, fig.height = 5, fig.width = 12, out.width = '100%', add.panel = TRUE}
show_results(results, 'Subgroup Thresholds Distribution Plot', 'visualizer')
```

##### Threshold Distances Plot

###### {.tabset .tabset-pills}

```{r, fig.height = 4.5, fig.width = 9, out.width = '100%', add.panel = TRUE}
show_results(results, 'Threshold Distances Plot', 'visualizer')
```

##### Subgroup ATE Errors Plot

###### {.tabset .tabset-pills}

```{r, fig.height = 4.5, fig.width = 8, out.width = '100%', add.panel = TRUE}
show_results(results, 'Subgroup ATE Errors Plot', 'visualizer')
```

##### Feature and Split Frequency Plots

###### {.tabset .tabset-pills}

```{r, fig.height = 7.5, fig.width = 18, out.width = '100%', add.panel = TRUE}
show_results(results, 'Feature and Split Frequency Plots', 'visualizer')
```

#### `r fontawesome::fa('code', fill = 'white')`

**Parameter Values**

```{r, max.height='200px'}
show_results(results, NULL, 'vary_params')
```

## Additive with Linear Covariates DGP {.tabset .tabset-pills}

```{r, results = 'asis'}
results <- get_exp_results(
  '/Users/ttang4/Documents/Research/Causal Subgroups/causalDT/causalDT-manuscript/results/Rulefit Simulations/Additive with Linear Covariates DGP', show_eval = FALSE, show_viz = TRUE, eval_cache = '.rds', viz_cache = '.png'
)
```

### Varying tau_heritability {.tabset .tabset-pills .tabset-circle}

```{r, results = 'asis'}
results <- get_exp_results(
  '/Users/ttang4/Documents/Research/Causal Subgroups/causalDT/causalDT-manuscript/results/Rulefit Simulations/Additive with Linear Covariates DGP/Varying tau_heritability', show_eval = FALSE, show_viz = TRUE, eval_cache = '.rds', viz_cache = '.png'
)
```

#### `r fontawesome::fa('chart-bar', fill = 'white')`

##### Subgroup Feature Selection Errors Plot

###### {.tabset .tabset-pills}

```{r, fig.height = 5, fig.width = 12, out.width = '100%', add.panel = TRUE}
show_results(results, 'Subgroup Feature Selection Errors Plot', 'visualizer')
```

##### Subgroup Thresholds Distribution Plot

###### {.tabset .tabset-pills}

```{r, fig.height = 5, fig.width = 12, out.width = '100%', add.panel = TRUE}
show_results(results, 'Subgroup Thresholds Distribution Plot', 'visualizer')
```

##### Threshold Distances Plot

###### {.tabset .tabset-pills}

```{r, fig.height = 4.5, fig.width = 9, out.width = '100%', add.panel = TRUE}
show_results(results, 'Threshold Distances Plot', 'visualizer')
```

##### Subgroup ATE Errors Plot

###### {.tabset .tabset-pills}

```{r, fig.height = 4.5, fig.width = 8, out.width = '100%', add.panel = TRUE}
show_results(results, 'Subgroup ATE Errors Plot', 'visualizer')
```

##### Feature and Split Frequency Plots

###### {.tabset .tabset-pills}

```{r, fig.height = 7.5, fig.width = 18, out.width = '100%', add.panel = TRUE}
show_results(results, 'Feature and Split Frequency Plots', 'visualizer')
```

#### `r fontawesome::fa('code', fill = 'white')`

**Parameter Values**

```{r, max.height='200px'}
show_results(results, NULL, 'vary_params')
```

## AND DGP {.tabset .tabset-pills}

```{r, results = 'asis'}
results <- get_exp_results(
  '/Users/ttang4/Documents/Research/Causal Subgroups/causalDT/causalDT-manuscript/results/Rulefit Simulations/AND DGP', show_eval = FALSE, show_viz = TRUE, eval_cache = '.rds', viz_cache = '.png'
)
```

### Varying tau_heritability {.tabset .tabset-pills .tabset-circle}

```{r, results = 'asis'}
results <- get_exp_results(
  '/Users/ttang4/Documents/Research/Causal Subgroups/causalDT/causalDT-manuscript/results/Rulefit Simulations/AND DGP/Varying tau_heritability', show_eval = FALSE, show_viz = TRUE, eval_cache = '.rds', viz_cache = '.png'
)
```

#### `r fontawesome::fa('chart-bar', fill = 'white')`

##### Subgroup Feature Selection Errors Plot

###### {.tabset .tabset-pills}

```{r, fig.height = 5, fig.width = 12, out.width = '100%', add.panel = TRUE}
show_results(results, 'Subgroup Feature Selection Errors Plot', 'visualizer')
```

##### Subgroup Thresholds Distribution Plot

###### {.tabset .tabset-pills}

```{r, fig.height = 5, fig.width = 12, out.width = '100%', add.panel = TRUE}
show_results(results, 'Subgroup Thresholds Distribution Plot', 'visualizer')
```

##### Threshold Distances Plot

###### {.tabset .tabset-pills}

```{r, fig.height = 4.5, fig.width = 9, out.width = '100%', add.panel = TRUE}
show_results(results, 'Threshold Distances Plot', 'visualizer')
```

##### Subgroup ATE Errors Plot

###### {.tabset .tabset-pills}

```{r, fig.height = 4.5, fig.width = 8, out.width = '100%', add.panel = TRUE}
show_results(results, 'Subgroup ATE Errors Plot', 'visualizer')
```

##### Feature and Split Frequency Plots

###### {.tabset .tabset-pills}

```{r, fig.height = 7.5, fig.width = 18, out.width = '100%', add.panel = TRUE}
show_results(results, 'Feature and Split Frequency Plots', 'visualizer')
```

#### `r fontawesome::fa('code', fill = 'white')`

**Parameter Values**

```{r, max.height='200px'}
show_results(results, NULL, 'vary_params')
```

## AND with Linear Covariates DGP {.tabset .tabset-pills}

```{r, results = 'asis'}
results <- get_exp_results(
  '/Users/ttang4/Documents/Research/Causal Subgroups/causalDT/causalDT-manuscript/results/Rulefit Simulations/AND with Linear Covariates DGP', show_eval = FALSE, show_viz = TRUE, eval_cache = '.rds', viz_cache = '.png'
)
```

### Varying tau_heritability {.tabset .tabset-pills .tabset-circle}

```{r, results = 'asis'}
results <- get_exp_results(
  '/Users/ttang4/Documents/Research/Causal Subgroups/causalDT/causalDT-manuscript/results/Rulefit Simulations/AND with Linear Covariates DGP/Varying tau_heritability', show_eval = FALSE, show_viz = TRUE, eval_cache = '.rds', viz_cache = '.png'
)
```

#### `r fontawesome::fa('chart-bar', fill = 'white')`

##### Subgroup Feature Selection Errors Plot

###### {.tabset .tabset-pills}

```{r, fig.height = 5, fig.width = 12, out.width = '100%', add.panel = TRUE}
show_results(results, 'Subgroup Feature Selection Errors Plot', 'visualizer')
```

##### Subgroup Thresholds Distribution Plot

###### {.tabset .tabset-pills}

```{r, fig.height = 5, fig.width = 12, out.width = '100%', add.panel = TRUE}
show_results(results, 'Subgroup Thresholds Distribution Plot', 'visualizer')
```

##### Threshold Distances Plot

###### {.tabset .tabset-pills}

```{r, fig.height = 4.5, fig.width = 9, out.width = '100%', add.panel = TRUE}
show_results(results, 'Threshold Distances Plot', 'visualizer')
```

##### Subgroup ATE Errors Plot

###### {.tabset .tabset-pills}

```{r, fig.height = 4.5, fig.width = 8, out.width = '100%', add.panel = TRUE}
show_results(results, 'Subgroup ATE Errors Plot', 'visualizer')
```

##### Feature and Split Frequency Plots

###### {.tabset .tabset-pills}

```{r, fig.height = 7.5, fig.width = 18, out.width = '100%', add.panel = TRUE}
show_results(results, 'Feature and Split Frequency Plots', 'visualizer')
```

#### `r fontawesome::fa('code', fill = 'white')`

**Parameter Values**

```{r, max.height='200px'}
show_results(results, NULL, 'vary_params')
```

## OR DGP {.tabset .tabset-pills}

```{r, results = 'asis'}
results <- get_exp_results(
  '/Users/ttang4/Documents/Research/Causal Subgroups/causalDT/causalDT-manuscript/results/Rulefit Simulations/OR DGP', show_eval = FALSE, show_viz = TRUE, eval_cache = '.rds', viz_cache = '.png'
)
```

### Varying tau_heritability {.tabset .tabset-pills .tabset-circle}

```{r, results = 'asis'}
results <- get_exp_results(
  '/Users/ttang4/Documents/Research/Causal Subgroups/causalDT/causalDT-manuscript/results/Rulefit Simulations/OR DGP/Varying tau_heritability', show_eval = FALSE, show_viz = TRUE, eval_cache = '.rds', viz_cache = '.png'
)
```

#### `r fontawesome::fa('chart-bar', fill = 'white')`

##### Subgroup Feature Selection Errors Plot

###### {.tabset .tabset-pills}

```{r, fig.height = 5, fig.width = 12, out.width = '100%', add.panel = TRUE}
show_results(results, 'Subgroup Feature Selection Errors Plot', 'visualizer')
```

##### Subgroup Thresholds Distribution Plot

###### {.tabset .tabset-pills}

```{r, fig.height = 5, fig.width = 12, out.width = '100%', add.panel = TRUE}
show_results(results, 'Subgroup Thresholds Distribution Plot', 'visualizer')
```

##### Threshold Distances Plot

###### {.tabset .tabset-pills}

```{r, fig.height = 4.5, fig.width = 9, out.width = '100%', add.panel = TRUE}
show_results(results, 'Threshold Distances Plot', 'visualizer')
```

##### Subgroup ATE Errors Plot

###### {.tabset .tabset-pills}

```{r, fig.height = 4.5, fig.width = 8, out.width = '100%', add.panel = TRUE}
show_results(results, 'Subgroup ATE Errors Plot', 'visualizer')
```

##### Feature and Split Frequency Plots

###### {.tabset .tabset-pills}

```{r, fig.height = 7.5, fig.width = 18, out.width = '100%', add.panel = TRUE}
show_results(results, 'Feature and Split Frequency Plots', 'visualizer')
```

#### `r fontawesome::fa('code', fill = 'white')`

**Parameter Values**

```{r, max.height='200px'}
show_results(results, NULL, 'vary_params')
```

## OR with Linear Covariates DGP {.tabset .tabset-pills}

```{r, results = 'asis'}
results <- get_exp_results(
  '/Users/ttang4/Documents/Research/Causal Subgroups/causalDT/causalDT-manuscript/results/Rulefit Simulations/OR with Linear Covariates DGP', show_eval = FALSE, show_viz = TRUE, eval_cache = '.rds', viz_cache = '.png'
)
```

### Varying tau_heritability {.tabset .tabset-pills .tabset-circle}

```{r, results = 'asis'}
results <- get_exp_results(
  '/Users/ttang4/Documents/Research/Causal Subgroups/causalDT/causalDT-manuscript/results/Rulefit Simulations/OR with Linear Covariates DGP/Varying tau_heritability', show_eval = FALSE, show_viz = TRUE, eval_cache = '.rds', viz_cache = '.png'
)
```

#### `r fontawesome::fa('chart-bar', fill = 'white')`

##### Subgroup Feature Selection Errors Plot

###### {.tabset .tabset-pills}

```{r, fig.height = 5, fig.width = 12, out.width = '100%', add.panel = TRUE}
show_results(results, 'Subgroup Feature Selection Errors Plot', 'visualizer')
```

##### Subgroup Thresholds Distribution Plot

###### {.tabset .tabset-pills}

```{r, fig.height = 5, fig.width = 12, out.width = '100%', add.panel = TRUE}
show_results(results, 'Subgroup Thresholds Distribution Plot', 'visualizer')
```

##### Threshold Distances Plot

###### {.tabset .tabset-pills}

```{r, fig.height = 4.5, fig.width = 9, out.width = '100%', add.panel = TRUE}
show_results(results, 'Threshold Distances Plot', 'visualizer')
```

##### Subgroup ATE Errors Plot

###### {.tabset .tabset-pills}

```{r, fig.height = 4.5, fig.width = 8, out.width = '100%', add.panel = TRUE}
show_results(results, 'Subgroup ATE Errors Plot', 'visualizer')
```

##### Feature and Split Frequency Plots

###### {.tabset .tabset-pills}

```{r, fig.height = 7.5, fig.width = 18, out.width = '100%', add.panel = TRUE}
show_results(results, 'Feature and Split Frequency Plots', 'visualizer')
```

#### `r fontawesome::fa('code', fill = 'white')`

**Parameter Values**

```{r, max.height='200px'}
show_results(results, NULL, 'vary_params')
```

# Varying nrep crossfits {.tabset .tabset-pills .tabset-vmodern}

```{r, results = 'asis'}
results <- get_exp_results(
  '/Users/ttang4/Documents/Research/Causal Subgroups/causalDT/causalDT-manuscript/results/Varying nrep crossfits', show_eval = FALSE, show_viz = TRUE, eval_cache = '.rds', viz_cache = '.png'
)
```

## Additive DGP-Distilled Rboost (crossfit) {.tabset .tabset-pills}

```{r, results = 'asis'}
results <- get_exp_results(
  '/Users/ttang4/Documents/Research/Causal Subgroups/causalDT/causalDT-manuscript/results/Varying nrep crossfits/Additive DGP-Distilled Rboost (crossfit)', show_eval = FALSE, show_viz = TRUE, eval_cache = '.rds', viz_cache = '.png'
)
```

### Varying tau_heritability-nreps_crossfit {.tabset .tabset-pills .tabset-circle}

```{r, results = 'asis'}
results <- get_exp_results(
  '/Users/ttang4/Documents/Research/Causal Subgroups/causalDT/causalDT-manuscript/results/Varying nrep crossfits/Additive DGP-Distilled Rboost (crossfit)/Varying tau_heritability-nreps_crossfit', show_eval = FALSE, show_viz = TRUE, eval_cache = '.rds', viz_cache = '.png'
)
```

#### `r fontawesome::fa('chart-bar', fill = 'white')`

##### Subgroup Feature Selection Errors Plot

###### {.tabset .tabset-pills}

```{r, fig.height = 5, fig.width = 12, out.width = '100%', add.panel = TRUE}
show_results(results, 'Subgroup Feature Selection Errors Plot', 'visualizer')
```

##### Threshold Distances Plot

###### {.tabset .tabset-pills}

```{r, fig.height = 4.5, fig.width = 9, out.width = '100%', add.panel = TRUE}
show_results(results, 'Threshold Distances Plot', 'visualizer')
```

##### Subgroup ATE Errors Plot

###### {.tabset .tabset-pills}

```{r, fig.height = 4.5, fig.width = 8, out.width = '100%', add.panel = TRUE}
show_results(results, 'Subgroup ATE Errors Plot', 'visualizer')
```

#### `r fontawesome::fa('code', fill = 'white')`

**Parameter Values**

```{r, max.height='200px'}
show_results(results, NULL, 'vary_params')
```

## AND DGP-Distilled Rboost (crossfit) {.tabset .tabset-pills}

```{r, results = 'asis'}
results <- get_exp_results(
  '/Users/ttang4/Documents/Research/Causal Subgroups/causalDT/causalDT-manuscript/results/Varying nrep crossfits/AND DGP-Distilled Rboost (crossfit)', show_eval = FALSE, show_viz = TRUE, eval_cache = '.rds', viz_cache = '.png'
)
```

### Varying tau_heritability-nreps_crossfit {.tabset .tabset-pills .tabset-circle}

```{r, results = 'asis'}
results <- get_exp_results(
  '/Users/ttang4/Documents/Research/Causal Subgroups/causalDT/causalDT-manuscript/results/Varying nrep crossfits/AND DGP-Distilled Rboost (crossfit)/Varying tau_heritability-nreps_crossfit', show_eval = FALSE, show_viz = TRUE, eval_cache = '.rds', viz_cache = '.png'
)
```

#### `r fontawesome::fa('chart-bar', fill = 'white')`

##### Subgroup Feature Selection Errors Plot

###### {.tabset .tabset-pills}

```{r, fig.height = 5, fig.width = 12, out.width = '100%', add.panel = TRUE}
show_results(results, 'Subgroup Feature Selection Errors Plot', 'visualizer')
```

##### Threshold Distances Plot

###### {.tabset .tabset-pills}

```{r, fig.height = 4.5, fig.width = 9, out.width = '100%', add.panel = TRUE}
show_results(results, 'Threshold Distances Plot', 'visualizer')
```

##### Subgroup ATE Errors Plot

###### {.tabset .tabset-pills}

```{r, fig.height = 4.5, fig.width = 8, out.width = '100%', add.panel = TRUE}
show_results(results, 'Subgroup ATE Errors Plot', 'visualizer')
```

#### `r fontawesome::fa('code', fill = 'white')`

**Parameter Values**

```{r, max.height='200px'}
show_results(results, NULL, 'vary_params')
```

## OR DGP-Distilled Rboost (crossfit) {.tabset .tabset-pills}

```{r, results = 'asis'}
results <- get_exp_results(
  '/Users/ttang4/Documents/Research/Causal Subgroups/causalDT/causalDT-manuscript/results/Varying nrep crossfits/OR DGP-Distilled Rboost (crossfit)', show_eval = FALSE, show_viz = TRUE, eval_cache = '.rds', viz_cache = '.png'
)
```

### Varying tau_heritability-nreps_crossfit {.tabset .tabset-pills .tabset-circle}

```{r, results = 'asis'}
results <- get_exp_results(
  '/Users/ttang4/Documents/Research/Causal Subgroups/causalDT/causalDT-manuscript/results/Varying nrep crossfits/OR DGP-Distilled Rboost (crossfit)/Varying tau_heritability-nreps_crossfit', show_eval = FALSE, show_viz = TRUE, eval_cache = '.rds', viz_cache = '.png'
)
```

#### `r fontawesome::fa('chart-bar', fill = 'white')`

##### Subgroup Feature Selection Errors Plot

###### {.tabset .tabset-pills}

```{r, fig.height = 5, fig.width = 12, out.width = '100%', add.panel = TRUE}
show_results(results, 'Subgroup Feature Selection Errors Plot', 'visualizer')
```

##### Threshold Distances Plot

###### {.tabset .tabset-pills}

```{r, fig.height = 4.5, fig.width = 9, out.width = '100%', add.panel = TRUE}
show_results(results, 'Threshold Distances Plot', 'visualizer')
```

##### Subgroup ATE Errors Plot

###### {.tabset .tabset-pills}

```{r, fig.height = 4.5, fig.width = 8, out.width = '100%', add.panel = TRUE}
show_results(results, 'Subgroup ATE Errors Plot', 'visualizer')
```

#### `r fontawesome::fa('code', fill = 'white')`

**Parameter Values**

```{r, max.height='200px'}
show_results(results, NULL, 'vary_params')
```