### 2025-02-13 causalDT/README.Rmd
fix readme links

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = file.path("man", "figures", "README-")
)
```

# Causal Distillation Trees

[Causal Distillation Trees](https://arxiv.org/abs/2502.07275) (CDT) is a novel machine learning method for estimating interpretable subgroups in causal inference. CDT allows researchers to fit *any* machine learning model of their choice to estimate the individual-level treatment effect, and then leverages a simple, second-stage tree-based model to "distill" the estimated treatment effect into meaningful subgroups. As a result, CDT inherits the improvements in predictive performance from black-box machine learning models while preserving the interpretability of a simple decision tree.

![](../causalDT/man/figures/cdt_diagram.png)

Briefly, CDT is a two-stage learner that first fits a teacher model (e.g., a black-box metalearner) to estimate individual-level treatment effects and secondly fits a student model (e.g., a decision tree) to predict the estimated individual-level treatment effects, in effect distilling the estimated individual-level treatment effects and producing interpretable subgroups. This two-stage learner is learned using the training data. Finally, using the estimated subgroups, the subgroup average treatment effects are honestly estimated with a held-out estimation set.

For more details, check out [Huang, M., Tang, T. M., Kenney, A. M. "Distilling heterogeneous treatment effects: Stable subgroup estimation in causal inference." (2025).](https://arxiv.org/abs/2502.07275)

## Installation of the R package

You can install the `causalDT` R package via:

``` r
# install.packages("devtools")
devtools::install_github("tiffanymtang/causalDT", subdir = "causalDT")
```

## Example Usage

To illustrate an example usage of `causalDT`, we will use the AIDS Clinical Trials Group Study 175 (ACTG 175), a randomized controlled trial to determine the effectiveness of monotherapy compared to combination therapy on HIV-1-infected patients. This data can be found in the `speff2trial` R package.

```{r load-data}
# install.packages("speff2trial")
library(speff2trial)
library(dplyr)

data <- speff2trial::ACTG175 |>
  dplyr::filter(arms %in% c(0, 2))

# pre-treatment covariates data
X <- data |> 
  dplyr::select(
    age, wtkg, hemo, homo, drugs, karnof, race, 
    gender, symptom, preanti, strat, cd80
  ) |> 
  as.matrix()
# treatment indicator variable
Z <- data |>
  dplyr::pull(treat)
# response variable
Y <- data |>
  dplyr::pull(cens)
```

Given the pre-treatment covariates data $X$, the treatment variable $Z$, and the response variable $Y$, we can run CDT as follows:

```{r causalDT, fig.width=12}
library(causalDT)

set.seed(331)
causal_forest_cdt <- causalDT(
  X = X, Y = Y, Z = Z,
  teacher_model = "causal_forest"
)

plot_cdt(causal_forest_cdt)
```

Note that when using CDT, a teacher model must be chosen (the default is a causal forest). To help researchers select an appropriate teacher model, the Jaccard subgroup stability index (SSI) was introduced in [Huang et al. (2025)](https://arxiv.org/abs/2502.07275). Generally, a higher Jaccard SSI indicates a better teacher model. This teacher model selection procedure can be run as follows:

```{r jaccard}
# selecting between causal forest versus rboost
rboost_cdt <- causalDT(
  X = as.matrix(X), Y = Y, Z = Z,
  teacher_model = rboost
)
plot_jaccard(`Causal Forest` = causal_forest_cdt, `Rboost` = rboost_cdt)
```

## Citation

```r
@article{huang2025distilling,
  title={Distilling heterogeneous treatment effects: Stable subgroup estimation in causal inference}, 
  author={Melody Huang and Tiffany M. Tang and Ana M. Kenney},
  year={2025},
  eprint={2502.07275},
  archivePrefix={arXiv},
  primaryClass={stat.ME},
  url={https://arxiv.org/abs/2502.07275}, 
}
```